### 2013-10-27 paper/paper.tex
added working draft of quicksched paper.

% v2-acmsmall-sample.tex, dated March 6 2012
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.3 - March 2012

\documentclass[prodmode,acmtecs]{acmsmall} % Aptara syntax

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information
\acmVolume{0}
\acmNumber{0}
\acmArticle{0}
\acmYear{2014}
\acmMonth{0}

% My own macros
\usepackage{epsfig}
\usepackage{listings}
\lstset{%
    language=C,
    basicstyle=\small\tt,
    numbers=left,
    numberstyle=\tiny
    }
\newcommand{\oh}[1]
    {\mbox{$ {\mathcal O}( #1 ) $}}
\newcommand{\eqn}[1]
    {(\ref{eqn:#1})}
\newcommand{\fig}[1]
    {Figure~\ref{fig:#1}}
\newcommand{\tabl}[1]
    {Table~\ref{tab:#1}}
\newcommand{\figs}[2]
    {Figures~\ref{fig:#1} and~\ref{fig:#2}}
\newcommand{\sect}[1]
    {Section~\ref{sec:#1}}
\newcommand{\cuda}
    {{\sc cuda}\xspace}
\newcommand{\namd}
    {{\sc namd}\xspace}
\newcommand{\acemd}
    {{\sc acemd}\xspace}
\newcommand{\fenzi}
    {{\sc fenzi}\xspace}
\newcommand{\mdcore}
    {{\tt mdcore}\xspace}
\newcommand{\swift}
    {{\sc swift}\xspace}
\newcommand{\bsf}[1]
    {\textbf{\textsf{#1}}}

% Document starts
\begin{document}

% Page heads
\markboth{P.~Gonnet}{QuickSched: Task-based parallelism with dependencies and conflicts}

% Title portion
\title{QuickSched: Task-based parallelism with dependencies and conflicts}
\author{PEDRO GONNET
\affil{Durham University}}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}
Bla.
\end{abstract}

\category{???}{Computing methodologies}{Shared memory algorithms \and Concurrent algorithms}

\terms{Algorithms, Design, Performance}

\keywords{Task-based parallelism}

\acmformat{Pedro Gonnet, 2014.
QuickSched: Task-based parallelism with dependencies and conflicts.}
% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

\begin{bottomstuff}
This work was supported by a Durham Universiy Seedcorn Grant.

Author's address: P. Gonnet, School of Engineering and Computing Sciences,
Durham University, South Road, Durham, DH1 3LE, United Kingdom.
\end{bottomstuff}

\maketitle

\section{Introduction}

Task-based parallelism is a conceptually simple paradigm for
shared-memory paralelism in which a computation is broken-down
into a set of inter-dependent tasks which are executed
concurrently.
Task dependencies are used to model the flow of data between
tasks, e.g.~if task $B$ requires some data generated by task $A$,
then task $B$ {\em depends} on task $A$ and cannot be executed
before task $A$ has completed.
The tasks and their dependencies can be seen as the nodes and edges
respectively of a Directed Acyclic Graph (DAG) which can be
traversed in topological order, executing the tasks at the nodes
on the way down.

This computational model is trival to parallelize.
Given a set of inter-dependent tasks and a set of computational
threads, each thread repeatedly selects a task with no
unsatisfied dependencies from the DAG and executes it.
If no tasks are available, the thread waits until any other
thread finishes executing a task, thus potentially releasing
new tasks, or until all tasks in the DAG have been executed.

\fig{Tasks} shows the DAG for a set of tasks.
The arrows indicate the direction of the dependency, i.e.~an
arrow from task $A$ to task $B$ indicates that task $B$ depends
on task $A$.
In a parallel setting, tasks $A$, $G$, and $J$ can be
executed concurrently.
Once task $G$ has completed, tasks $F$ and $H$ become available
and can be executed by any other computational thread.

\begin{figure}
    \centerline{\epsfig{file=figures/Tasks.pdf,width=0.5\textwidth}}
    \caption{A set of tasks (circles) and their dependencies (arrows).
        The arrows indicate the direction of the dependency, i.e.~an
        arrow from task $A$ to task $B$ indicates that task $B$ depends
        on task $A$.
        Tasks $A$, $G$, and $J$ have no unsatisifed dependencies and
        can therefore be executed.
        Once task $G$ has completed, tasks $F$ and $H$ become available.}
    \label{fig:Tasks}
\end{figure}

One of the first implementations of a task-based parallel programming
systems is Cilk \cite{ref:Blumofe1995}, an extension to the C
programming language which allows function calls to be ``spawned''
as new tasks.
Dependencies are enforced by the {\tt sync} keyword, which
forces a thread to wait for all the tasks that it spawned
to complete.
Although simple to use, this implicit dependency management
limits the types of DAGs that can be represented, i.e.~for
the example in \fig{Tasks}, using such a spawning model
would create implicit dependencies between the lowest-level
tasks $C$, $E$, and $K$.

In SMP superscalar \cite{ref:Perez2008}, StarPU \cite{ref:Augonnet2011},
QUARK \cite{ref:Yarkhan2011}, and KAAPI \cite{ref:Gautier2007}
the programmer spcifies
what shared data each task will access, and how that data will
be accessed, e.g.~read, write, or read-write access.
The dependencies between tasks are then generated
automatically by the runtime system, assuming that the
data must be accessed and updated in the order in which
the tasks are generated.
StarPU also provides an interface for specifying additional
dependencies explicitly.
Intel's Threding Building Blocks (TBB)
\cite{ref:Reinders2010}
provide task-based parallelism using C++ templates.
Dependencies are handled either by explicitly waiting
for spawned tasks, or by explicitly manipulating
task reference counters.

Finally, the very popular OpenMP standard provides some basic support
for spawning tasks, similar to Cilk, as of version 3.0
\cite{ref:OpenMP2008}.
OmpSs \cite{ref:Duran2011} extends this scheme with automatic
dependency generation as in SMP superscalar, along with
the ability to explicitly wait on certain tasks.

In all of these systems, the tasks are only aware of a single
type of relationship between each other, i.e. dependencies, which
specify a strict ordering between two tasks.
In many cases, however, the task ordering need not necessarily
be this strict.
Consider the case of two tasks that update some shared resource
in an order-independent way, e.g. when accumulating a result in
a shared variable, or exclusively writing to an output file.
In order to avoid concurrent access to that resource, it is
imperative that the execution of both tasks does not overlap,
yet the order in which the tasks are exectued does not matter.
In the following, such a relationship will be refered to
as a ``conflict'' between two tasks.
\fig{TaskConflicts} shows a task graph with conflicting tasks
joined by thick dashed lines.
None of tasks $F$, $H$, and $I$ cannot be executed concurrently,
i.e. they must be serialized, yet in no particular order.

Conflicts can be modeled as exclusive locks on a shared resource
which have to be obtained by a task before it can execute.
Thus, in \fig{TaskConflicts}, before executing, task $F$ has
to obtain an exclusive lock on the resource associated with
the conflict between tasks $F$, $H$, and $I$.
While task $F$ is being executed, neither $H$ nor $I$ can 
lock the same resource, and therefore will not execute until
task $F$ is done and the lock has been released.

In dependency-only systems, such conflicts can be modelled
with dependencies, which enforce a pre-determined arbitrary
ordering on conflicting tasks.
This, however, imposes unnecessary restriction on the order
in which tasks can be scheduled, especially in the presence
of multiple conflicts per task.
Both \citeN{ref:Ltaief2012} and \citeN{ref:Agullo2013} note
this problem in their respective implementations of the Fast Multipole
Method, in which forces computed in different tasks are
accumulated on a set of particles.

This paper presents QuickSched, a framework for task-based
parallel programming with constraints.
Section~2 describes the underlying algorithms and data structures,
and Section~3 describes their specific implementation in QuickSched.
Section~4 presents two test-cases:
\begin{itemize}
    \item The tiled QR
    decomposition described in \cite{ref:Buttari2009} and for
    which the QUARK scheduler was originally developed,
    \item A task-based Barnes-Hut tree-code to compute the
    gravitational N-body problem,
\end{itemize}
These real-world examples show how QuickSched can be used in practice,
and can be used to assess its efficiency.
Section~5 concludes with some general observations and future work.

\begin{figure}
    \centerline{\epsfig{file=figures/TaskConflicts.pdf,width=0.5\textwidth}}
    \caption{Task graph with conflicts (thick dashed lines).
        If two or more tasks are joined by a conflict, they cannot be
        executed concurrently, i.e. tasks $B$ and $D$ cannot be run an
        the same time.
        Tasks belonging to different conflicting sets, e.g. tasks $B$
        and $F$, or tasks $F$ and $J$, however, can be executed
        concurrently.}
    \label{fig:TaskConflicts}
\end{figure}

\section{Data Structures and Algorithms}

The QuickSched task scheduler consits of four main
objects types: {\em task}, {\em resource}, {\em scheduler},
and {\em queue}.

The task and resource objects are used
to model the computation, i.e. the work that is to be done
and the data on which it will be done, respectively.
The scheduler and queue objects manage
how the work is done, i.e. which tasks get scheduled
where and when, respectively.

\begin{figure}
    \centerline{\epsfig{file=figures/QSched.pdf,width=0.7\textwidth}}
    \caption{Schematic of the QuickSched task scheduler.
        The tasks (circles) are stored in the scheduler (left).
        Once a task's dependencies have been resolved, the task
        is moved to one of the task queues.
        Tasks that are not involved in any active conflicts
        are then taken from the queues by the different
        computational threads and executed.
        After execution, their dependent tasks are unlocked
        in the scheduler (dashed arrow).}
    \label{fig:QSched}
\end{figure}

The division of labour regarding {\em correctness}
between the scheduler and
the queue objects is illustrated in \fig{QSched}.
The scheduler holds the tasks and is in charge
of managing {\em dependencies}.
Once a task has no unresolved dependencies, it is passed
on to a queue object.
The queue object, on the other hand, is in charge
of managing {\em conflicts}.
Computational threads can query a queue and will
receive only tasks for which all conflicts have been
resolved, i.e. for which all necessary resources could be 
exclusively locked.

There is also a division of responsibilities regarding
{\em efficiency} between the scheduler and the queue
objects.
The tasks in each queue are grouped according to the resources
they use, i.e. all the tasks in the same queue use a
similar set of resources.
The underlying assumption is that each computational
thread will preferentially access the same queue for tasks.
If the tasks in the queue share the same set of resources,
in increases the probability of said resources already
being present in the thread's cache, thus increasing
{\em memory efficiency}.
The scheduler is in charge of selecting the most appropriate
queue for each task, based on information stored in each task
on which resources are used.
Given a set of tasks with similar resources for which all
dependencies are resolved, the queue then decides which
tasks to prioritize.
This decision is made based on the length of the critical
path of the dependencies of each task.

The following subsections describe these four object types
in detail, as well as their operations.

\subsection{Tasks}

A task consists of the following data structure, in C-like pseudo-code:

\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
struct task {
    int type, wait;
    void *data;
    struct task **unlocks;
    struct resource **locks, **uses;
    int size_data, nr_unlocks, nr_locks, nr_uses;
    int cost, weight;
    };
    \end{lstlisting}
\end{minipage}\end{center}

What the task does is determined by the {\tt type}
field, e.g.~which can be mapped to any particular function,
and the {\tt data} pointer which points to an array of
{\tt size\_data} bytes containing data specific to the task,
e.g.~the parameters for a specific function call.
Both fields are application-specific and therefore not
important for the scheduler itself.

The {\tt unlocks} field points to the first element of
an array of {\tt nr\_unlocks} pointers to other tasks.
These pointers represent the dependencies in reverse:
if task $B$ depends on task $A$, then task $A$ {\em unlocks}
task $B$.
The unlocks therefore follow the direction of the arrows
in \figs{Tasks}{TaskConflicts}.

Conversely, {\tt wait} is the number of unresolved dependencies
associated with this task, i.e.~the number of unexecuted tasks
that unlock this task.
Given an array of {\tt N} tasks, the wait-counters can
be set as follows:
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
int j, k;
for ( k = 0 ; k < N ; k++ )
    tasks[k].wait = 0;
for ( k = 0 ; k < N ; k++ )
    for ( j = 0 ; j < tasks[k].nr_unlocks ; j++ )
        tasks[k].unlocks[j]->wait += 1;
    \end{lstlisting}
\end{minipage}\end{center}

The {\tt locks} field points to the first element of
an array of {\tt nr\_locks} pointers to {\em resources}
for which exclusive locks must be obtained for the task
to execute.
Each locked resource represents a task conflict.
Similarly, {\tt uses} points to the first element of
an array of {\tt nr\_uses} pointers to resources which
will be used, but need not be locked.

Finally, {\tt cost} and {\tt weight} are measures
for the relative computational cost of this task, and the
relative cost of the critical path following the
dependencies of this task, respectively.
The task weights can be computed by first sorting
the tasks topologically according to their dependencies, i.e.
as per \citeN{ref:Kahn1962}:
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
int top[N], j, k, sorted = 0, maxweight;
set the task waits as described above.
for ( k = 0 ; k < N ; k++ )
    if ( tasks[k].wait == 0 )
        top[ sorted++ ] = k;
for ( k = 0 ; k < sorted ; k++ )
    for ( j = 0 ; j < tasks[ top[k] ].nr_unlocks ; j++ )
        if ( ( tasks[ top[k] ].unlock[j]->wait -= 1 ) == 0 )
            top[ sorted++ ] = k;
if ( k < N )
    error( "Circular dependencies detected." );
    \end{lstlisting}
\end{minipage}\end{center}
\noindent where the array {\tt top} contains the task indices
in reverse topological order. 
The test in line~10 is a convenient check if the tasks and teir
dependencies actually do form an acyclic graph.
The weights themselves are then computed as follows
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
for ( k = N-1 ; k >= 0 ; k-- ) {
    for ( w = 0 , j = 0 ; j < tasks[ top[k] ].nr_unlocks ; j++ ) {
        tasks[ top[k] ].unlock[j]->wait += 1;
        w = max( tasks[ top[k] ].unlock[j]->weight , w );
        }
    tasks[ top[k] ].weight = tasks[ top[k] ].cost + w;
    }
    \end{lstlisting}
\end{minipage}\end{center}
\noindent where the tasks are traversed in reverse
topological order, computing the recursive weight as the sum of the
task cost and the maximum weight of the tasks it unlocks,
and recomputing the task waits at the same time.

\subsection{Resources}

The data structure for the resources is as follows:
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
struct resource {
    struct resource *parent;
    volatile int lock, hold;
    int owner;
    };
    \end{lstlisting}
\end{minipage}\end{center}

The {\tt parent} field, which points to another resource, is
used to create hierarchical resources, i.e.~resources
that are themselves subsets of larger resources.
This can be useful, e.g.~in the context of particle simulations
described in the next section, where particles are sorted
into hierarchical cells which are used at different levels.

The {\tt lock} field is either {\tt 0} or {\tt 1} and indicates
whether this resource is currently in use, i.e.~{\em locked}.
To avoid race conditions, this value should only be tested
and set using atomic instructions.
The {\tt hold} field is a counter indicating how many
sub-resources of the current resouce are locked.
If a resource's hold counter is not zero, then it is
{\em held} and cannot be locked.
Likewise, if a resource is locked, it cannot be held
(see \fig{Resources}).

\begin{figure}
    \centerline{\epsfig{file=figures/Resources.pdf,width=0.6\textwidth}}
    \caption{A hierarchicy of cells (left) and the hierarchy of
        corresponding hierarchical resources at each level.
        Each square on the right represents a single resource, and
        arrows indicate the resource's parent.
        Resources coloured red are locked, resources coloured orange
        are held, where the number in the square indicates the
        value of the hold counter.}
    \label{fig:Resources}
\end{figure}

Incrementing the hold counter of the resource can be implemented
as follows:
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
void resource_hold ( struct resource *r ) {
    if ( atomic_cas( &r->lock , 0 , 1 ) != 0 )
        return 0;
    atomic_inc( &r->hold );
    r->lock = 0;
    return 1;
    }
    \end{lstlisting}
\end{minipage}\end{center}
\noindent where {\tt atomic\_cas(val,old,new)} is an atomic
compare-and-swap operation that sets {\tt val} to {\tt new}
if it is currently equal to {\tt old}.
Similarly, {\tt atomic\_inc(val)} increments {\tt val} by one
atomically.
The resource's {\tt lock} is used to check if the resource
is already locked (line 2), and it is held while {\tt hold}
is incremented, to avoid overlapping hold/lock operations.
If the resource can be locked, the hold counter is incremented
atomically (line~4), and the lock is released (line~5),
returning {\tt 1} or {\tt 0} if the resource could be held
or not, respectively.

The locking procedure itself is implemented as follows:
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
void resource_lock ( struct resource *r ) {
    struct resource *up, *top;
    if ( r->hold && atomic_cas( &r->lock , 0 , 1 ) != 0 )
        return 0;
    if ( r->hold ) {
        r->lock = 0;
        return 0;
        }
    for ( up = r->parent ; up != NULL ; up = up->parent ) {
        if ( !resource_hold( up ) )
            break;
    if ( ( top = up ) != NULL ) {
        for ( up = r->parent ; up != top , up = up->parent )
            atomic_dec( &up->hold );
        r->lock = 0;
        return 0;
        }
    else
        return 1;
    }
    \end{lstlisting}
\end{minipage}\end{center}
\noindent where in line~3 the resource is first locked if it
is not held.
Due to a possible race condition when holding the resource,
the {\tt hold} counter must checked again once the resource
has been locked (line~5).
In lines~9--11 the hold counters of the hierarchical parents
are incremented using the procedure described earlier.
If this process fails at any point (line~12), the
previously set hold counters are decremented (lines~13--14)
and the lock is released (line~15).
The procedure then returns {\tt 1} or {\tt 0} if the resource
could be locked or not, respectively.

Finally, unlocking a resource is relatively straight-forward:
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
void resource_unlock ( struct resource *r ) {
    struct resource *up;
    r->lock = 0;
    for ( up = r->parent ; up != NULL ; up = up->parent ) {
        atomic_dec( &up->hold );
    }
    \end{lstlisting}
\end{minipage}\end{center}
\noindent where the resource itself is unlocked (line~3)
and the hold counter of its parents is decremented (lines~4--5).

\subsection{Queues}

The main job of the task queues is, given a set of ready tasks,
to find the task with maximum weight whose resources can all
be locked, and to do so as efficiently as possible.

One possible strategy would be to maintain an array of tasks
sorted by their weights, and to trverse that list in descending
order, trying to lock the resources of each task, until
a lockable task is found, or returning a failure otherwise.
Although this would return the best possible task, it
requires maintaining a sorted list in which inserting
or removing an entry is in \oh{n} for $n$ elements.

Using an unsorted array requires only \oh{1} operations for
insertion and deletion, but is undesireable as it completely
ignores the task weights.

As a compromise, the queue stores the tasks in an array
organized as a max-heap, with the task with maximum weight
in the first position.
Maintainig this heap structure thus requires \oh{\log n}
operations for both insertion and deletion, i.e. for the
bubble-up and trickle-down operations respectively.

The array of tasks is then traversed as if it were sorted,
returning the first task that can be locked.
Although the first task in the array will be the task with
maximum weight, the following tasks are only losely ordered,
where the $k$th of $n$ tasks has a larger weight than at least
$\lfloor n/k\rfloor -1$ other tasks.

The data structure for the queue is thus defined as follows:
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
struct queue {
    struct task **tasks;
    int count, lock;
    };
    \end{lstlisting}
\end{minipage}\end{center}
\noindent where {\tt tasks} is an array of {\tt count} pointers
to the tasks in max-heap order, and {\tt lock} is used to
guarantee exclusive access to the queue.

Inserting a task in the queue is relatively straight-forward:
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
void queue_put ( struct queue *q , struct task *t ) {
    while ( atomic_cas( q->lock , 0 , 1 ) != 0 );
    q->tasks[ q->count++ ] = t;
    bubble-up the q->count-1st entry of q->tasks.
    q->lock = 0;
    }
    \end{lstlisting}
\end{minipage}\end{center}
\noindent where the loop in line~2 spins until an exclusive
lock on the queue can be obtained.
The task is added to the end of the heap array (line~3)
and the heap order is fixed (line~4).
Before exiting, the lock on the queue is released (line~5).

Obtaining a task from the queue can be implemented as follows:
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
struct task *queue_get ( struct queue *q ) {
    struct task *res = NULL;
    int j, k;
    while ( atomic_cas( q->lock , 0 , 1 ) != 0 );
    for ( k = 0 ; k < q->count ; k++ ) {
        for ( j = 0 ; j < q->tasks[k]->nr_locks ; j++ )
            if ( !resource_lock( q->tasks[k]->lock[j] )
                break;
        if ( j < q->tasks[k]->nr_locks )
            for ( j = j-1 ; j >= 0 ; j-- )
                resource_unlock( q->tasks[k]->lock[j] );
        else
            break;
        }
    if ( k < q->count ) {
        res = q->tasks[k];
        q->tasks[k] = q->tasks[ --q->count ];
        trickle-down the kth entry of q->tasks.
        }
    q->lock = 0;
    return res;
    }
    \end{lstlisting}
\end{minipage}\end{center}
\noindent where as with the queue insertion, the queue is first
locked for exclusive access (line~4).
The array of task pointers is then traversed in heap order (line~5),
locking the resources of each task (lines~6--8).
If any of these locks fail (line~9), the locks that were obtained
are released (lines~10--11), otherwise, the traversal is aborted
(line~13).
If all the locks on a task could be obtained (line~14), the
task pointer is replaced by the last pointer in the heap (line~17)
and the heap order is restored (line~18).
Finally, the queue lock is released (line~19) and the locked task
or, if no lockable task could be found, {\tt NULL} is returned.

\subsection{Scheduler}

The scheduler object is used as the main interface to the
QuickSched task scheduler, and as such contains the instances
other three object types:
\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
struct qsched {
    struct task *tasks;
    struct queue *queues;
    struct resource *res;
    int nr_tasks, nr_queues, nr_resources;
    volatile int waiting;
    };
    \end{lstlisting}
\end{minipage}\end{center}
\noindent where\dots
Note that for brevity, and to avoid conflicts with the naming
schemes of other standard libraries, the type name {\tt qsched}
is used.

\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
void qsched_start ( qsched *s ) {
    set the task waits and weights.
    s->waiting = s->nr_tasks;
    for ( k = 0 ; k < s->nr_tasks && s->tasks[tid[k]].wait == 0 ; k++ )
        qsched_enqueue( s , tid[k] );
    }
    \end{lstlisting}
\end{minipage}\end{center}

\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
void qsched_enqueue ( qsched *s , struct task *t ) {
    int k, qid, score[ s->nr_queues ];
    for ( k = 0 ; k < s->nr_queues ; k++ )
        score[k] = 0;
    for ( j = 0 ; j < t->nr_locks ; j++ )
        score[ t->locks[j]->owner ] += 1;
    for ( j = 0 ; j < t->nr_uses ; j++ )
        score[ t->uses[j]->owner ] += 1;
    for ( qid = 0 , k = 1 ; k < s->nr_queues ; k++ )
        if ( score[k] > score[qid] )
            qid = k;
    queue_put( &s->queues[qid] , t );
    }
    \end{lstlisting}
\end{minipage}\end{center}

\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
struct task *qsched_gettask ( qsched *s , int qid ) {
    struct task *res;
    int k;
    while ( s->waiting ) {
        if ( ( res = queue_get( s->queues[qid] ) ) == NULL ) {
            k = id of queue with largest count.
            if ( ( res = queue_get( s->queues[k] ) ) != NULL )
                break;
            }
        else
            break;
        }
    if ( res != NULL ) {
        for ( k = 0 ; k < res->nr_locks ; k++ )
            res->locks[k]->owner = qid;
        for ( k = 0 ; k < res->nr_uses ; k++ )
            res->uses[k]->owner = qid;
        }
    return res;
    }
    \end{lstlisting}
\end{minipage}\end{center}

\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
struct task *qsched_done ( qsched *s , struct task *t ) {
    int k;
    for ( k = 0 ; k < t->nr_unlocks ; k++ )
        if ( atomic_dec( &t->unlocks[k]->wait ) == 1 )
            qsched_enqueue( s , t->unlocks[k] );
    atomic_dec( &s->waiting );
    }
    \end{lstlisting}
\end{minipage}\end{center}

\begin{center}\begin{minipage}{0.9\textwidth}
    \begin{lstlisting}
struct task *qsched_run ( qsched *s , void (*fun)( int , void * ) ) {
    #pragma omp parallel
    {
        int qid = omp_get_thread_num() % s->nr_queues;
        struct task *t;
        while ( ( t = qsched_gettask( s , qid ) ) != NULL ) {
            fun( t->type , t->data );
            qsched_done( s , t );
            }
        }
    }
    \end{lstlisting}
\end{minipage}\end{center}

\section{User Interface}

\section{Validation}

\subsection{Task-Based QR Decomposition}

\subsection{Task-Based Barnes-Hut N-Body Solver}

\section{Conclusions}

Main points:
\begin{itemize}
    \item Conflicts are better than dependencies.
    \item Specifying the entire task tree before execution
        gives better scheduling.
    \item Combination of locality, weight, and availability.
\end{itemize}

% Acknowledgments
\begin{acks}
The author would like to thank Sam Townsend for implementing the
kernels in the task-based QR decomposition as part of his MSc Thesis
at Durham University, as well as Lydia Heck of the Institute for
Computational Cosmology at Durham University for providing access
to, and expertise on, the COSMA cluster.
\end{acks}

% Bibliography
\bibliographystyle{hacm}
\bibliography{quicksched}
                             % Sample .bib file with references that match those in
                             % the 'Specifications Document (V1.5)' as well containing
                             % 'legacy' bibs and bibs with 'alternate codings'.
                             % Gerry Murray - March 2012

% History dates
\received{February 2007}{March 2009}{June 2009}

\end{document}
% End of v2-acmsmall-sample.tex (March 2012) - Gerry Murray, ACM

### 2013-10-28 src/qsched.c
a bit of renaming.

@@ -29,7 +29,7 @@
 #include "lock.h"
 #include "task.h"
 #include "res.h"
-#include "sched.h"
+#include "qsched.h"
 #include "queue.h"
 
 /* Max macro. */
@@ -43,7 +43,7 @@
  * @param t Pointer to the #task.
  */
  
-void *sched_getdata( struct sched *s , struct task *t ) {
+void *qsched_getdata( struct qsched *s , struct task *t ) {
 
     return &s->data[ t->data ];
     
@@ -57,7 +57,7 @@ void *sched_getdata( struct sched *s , struct task *t ) {
  * @param t Pointer to the #task.
  */
  
-void sched_enqueue ( struct sched *s , struct task *t ) {
+void qsched_enqueue ( struct qsched *s , struct task *t ) {
 
     int j, qid, scores[ s->nr_queues ];
     
@@ -66,7 +66,7 @@ void sched_enqueue ( struct sched *s , struct task *t ) {
         
         /* This task is done before it started. */
         t->tic = getticks();
-        sched_done( s , t );
+        qsched_done( s , t );
         
         }
         
@@ -105,14 +105,14 @@ void sched_enqueue ( struct sched *s , struct task *t ) {
  * @param t Pointer to the completed #task.
  */
  
-void sched_done ( struct sched *s , struct task *t ) {
+void qsched_done ( struct qsched *s , struct task *t ) {
 
     int k;
     struct task *t2;
     
     /* Release this task's locks. */
     for ( k = 0 ; k < t->nr_locks ; k++ )
-        sched_unlockres( s , t->locks[k] );
+        qsched_unlockres( s , t->locks[k] );
     
     /* Loop over the task's unlocks... */
     for ( k = 0 ; k < t->nr_unlocks ; k++ ) {
@@ -122,7 +122,7 @@ void sched_done ( struct sched *s , struct task *t ) {
 
         /* Is the unlocked task ready to run? */
         if ( atomic_dec( &t2->wait ) == 1 && !( t2->flags & task_flag_skip ) )
-            sched_enqueue( s , t2 );
+            qsched_enqueue( s , t2 );
             
         }
         
@@ -142,7 +142,7 @@ void sched_done ( struct sched *s , struct task *t ) {
  * @return @c 1 if the resource could be locked, @c 0 otherwise.
  */
  
-int sched_lockres ( struct sched *s , int rid ) {
+int qsched_lockres ( struct qsched *s , int rid ) {
 
     int finger, finger2;
     
[…]

### 2013-10-28 src/qsched.h

@@ -18,24 +18,24 @@
  ******************************************************************************/
 
 /* Scheduler flags. */
-#define sched_flag_none                 0
-#define sched_flag_dirty                1
-#define sched_flag_ready                2
+#define qsched_flag_none                 0
+#define qsched_flag_dirty                1
+#define qsched_flag_ready                2
 
 /* Some sched-specific constants. */
-#define sched_stretch                   1.41
-#define sched_inc                       10
-#define sched_init_depspertask          2
-#define sched_init_lockspertask         2
-#define sched_init_usespertask          2
-#define sched_init_respertask           2
-#define sched_init_datapertask          2
-#define sched_data_round                16
-#define sched_res_none                  (-1)
+#define qsched_stretch                   1.41
+#define qsched_inc                       10
+#define qsched_init_depspertask          2
+#define qsched_init_lockspertask         2
+#define qsched_init_usespertask          2
+#define qsched_init_respertask           2
+#define qsched_init_datapertask          2
+#define qsched_data_round                16
+#define qsched_res_none                  (-1)
 
 
 /* The sched data structre. */
-struct sched {
+struct qsched {
 
     /* Flags for this scheduler. */
     unsigned int flags;
@@ -109,21 +109,21 @@ struct sched {
 
 
 /* Function prototypes. */
-void sched_init ( struct sched *s , int nr_queues , int size );
-void sched_sort ( int *restrict data , int *restrict ind , int N , int min , int max );
-void sched_sort_rec ( int *restrict data , int *restrict ind , int N , int min , int max );
-void sched_prepare ( struct sched *s );
-int sched_addres ( struct sched *s , int parent );
-void sched_addlock ( struct sched *s , int t , int res );
-void sched_addunlock ( struct sched *s , int ta , int tb );
-int sched_newtask ( struct sched *s , int type , int subtype , unsigned int flags , void *data , int data_size , int cost );
-struct task *sched_gettask ( struct sched *s , int qid );
-void sched_adduse ( struct sched *s , int t , int res );
-void sched_done ( struct sched *s , struct task *t );
-void *sched_getdata( struct sched *s , struct task *t );
-void sched_free ( struct sched *s );
-int sched_lockres ( struct sched *s , int rid );
-void sched_unlockres ( struct sched *s , int rid );
-int sched_locktask ( struct sched *s , int tid );
[…]

### 2013-10-28 src/queue.c

@@ -28,7 +28,7 @@
 #include "atomic.h"
 #include "lock.h"
 #include "task.h"
-#include "sched.h"
+#include "qsched.h"
 #include "queue.h"
 
 
@@ -41,7 +41,7 @@
  * @return The task ID or -1 if no available task could be found.
  */
  
-int queue_get ( struct queue *q , struct sched *s ) {
+int queue_get ( struct queue *q , struct qsched *s ) {
 
     int k, j, temp, tid, *inds, w, count;
     struct task *tasks = s->tasks;
@@ -65,7 +65,7 @@ int queue_get ( struct queue *q , struct sched *s ) {
         tid = inds[k];
         
         /* If the task can be locked, break. */
-        if ( sched_locktask( s , tid ) )
+        if ( qsched_locktask( s , tid ) )
             break;
     
         }
@@ -117,7 +117,7 @@ int queue_get ( struct queue *q , struct sched *s ) {
  * @param tid The task index.
  */
  
-void queue_put ( struct queue *q , struct sched *s , int tid ) {
+void queue_put ( struct queue *q , struct qsched *s , int tid ) {
 
     int ind, j, temp;
     struct task *tasks = s->tasks;

### 2013-10-28 src/queue.h

@@ -40,7 +40,7 @@ struct queue {
 
 
 /* Function prototypes. */
-int queue_get ( struct queue *q , struct sched *s );
-void queue_put ( struct queue *q , struct sched *s , int tid );
+int queue_get ( struct queue *q , struct qsched *s );
+void queue_put ( struct queue *q , struct qsched *s , int tid );
 void queue_init ( struct queue *q , int size );
 void queue_free ( struct queue *q );

### 2013-10-28 src/quicksched.h

@@ -24,5 +24,5 @@
 #include "atomic.h"
 #include "lock.h"
 #include "task.h"
-#include "sched.h"
+#include "qsched.h"
 #include "queue.h"

### 2013-10-28 src/task.h

@@ -26,8 +26,10 @@
 /* The task data structure. */
 struct task {
 
-    /* Task type, subtype, and flags. */
-    int type, subtype;
+    /* Task type. */
+    int type;
+    
+    /* Task flags. */
     unsigned int flags;
     
     /* Task payload offset. */

### 2013-10-28 examples/mkplots_tasks.m
a bit of renaming.

@@ -84,7 +84,7 @@ print -depsc2 tasks_qr_dynamic.eps
 
 
 
-%% Plot the tiled Barnes-Hutt tasks
+%% Plot the tiled Barnes-Hut tasks
 
 %% Plot the task timelines for tasks allocation
 % Load the data
@@ -112,7 +112,7 @@ hold off;
 xlabel('time (ms)');
 ylabel('core ID');
 set(gca,'YTick',1:(max(tasks(:,1))+1))
-title('Barnes-Hutt tasks');
+title('Barnes-Hut tasks');
 axis([ 0 , max( tasks(:,3) + tasks(:,4) ) , -0.5 , nr_cores-0.5 ]);
 
 % Print this plot

### 2013-10-28 examples/test.c

@@ -62,7 +62,7 @@ void matmul ( int m , int n , int k , double *a , int lda , double *b , int ldb 
 void test2 ( int m , int n , int k , int nr_threads ) { 
 
     int i, j, kk, qid, data[3], *d, tid, rid;
-    struct sched s;
+    struct qsched s;
     struct task *t;
     double *a, *b, *c, *res, err = 0.0, irm = 1.0/RAND_MAX;
     ticks tic_task, toc_task, tic_ref, toc_ref;
@@ -72,8 +72,8 @@ void test2 ( int m , int n , int k , int nr_threads ) {
              "C_ij = A_i: * B_:j, with tasks for each k where C_ij += A_ik*B_kj." );
     
     /* Init the sched. */
-    bzero( &s , sizeof(struct sched) );
-    sched_init( &s , nr_threads , m * n );
+    bzero( &s , sizeof(struct qsched) );
+    qsched_init( &s , nr_threads , m * n );
 
     /* Allocate the matrices. */
     if ( ( a = (double *)malloc( sizeof(double) * m * k * 32 * 32 ) ) == NULL ||
@@ -93,17 +93,17 @@ void test2 ( int m , int n , int k , int nr_threads ) {
     /* Build a task for each tile of the matrix c. */
     for ( i = 0 ; i < m ; i++ )
         for ( j = 0 ; j < n ; j++ ) {
-            rid = sched_addres( &s , -1 );
+            rid = qsched_addres( &s , -1 );
             data[0] = i; data[1] = j;
             for ( kk = 0 ; kk < k ; kk++ ) {
                 data[2] = kk;
-                tid = sched_newtask( &s , 1 , 0 , 0 , data , 3*sizeof(int) , 1 );
-                sched_addlock( &s , tid , rid );
+                tid = qsched_newtask( &s , 1 , task_flag_none , data , 3*sizeof(int) , 1 );
+                qsched_addlock( &s , tid , rid );
                 }
             }
             
     /* Prepare the sched for execution. */
-    sched_prepare( &s );
+    qsched_prepare( &s );
             
     /* Parallel loop. */
     tic_task = getticks();
@@ -117,13 +117,13 @@ void test2 ( int m , int n , int k , int nr_threads ) {
             while ( 1 ) {
 
                 /* Get a task, break if unsucessful. */
-                if ( ( t = sched_gettask( &s , qid ) ) == NULL )
+                if ( ( t = qsched_gettask( &s , qid ) ) == NULL )
                     break;
 
[…]

### 2013-10-28 examples/test_bh.c

@@ -104,7 +104,7 @@ struct cell *cell_get ( ) {
     cell_pool = cell_pool->progeny[0];
     
     /* Clean up a few things. */
-    res->res = sched_res_none;
+    res->res = qsched_res_none;
     
     /* Return the cell. */
     return res;
@@ -121,7 +121,7 @@ struct cell *cell_get ( ) {
  * @param s The #sched to store the resources.
  */
  
-void cell_split ( struct cell *c , struct sched *s ) {
+void cell_split ( struct cell *c , struct qsched *s ) {
 
     int i, j, k, count = c->count;
     struct part temp, *parts = c->parts;
@@ -130,11 +130,11 @@ void cell_split ( struct cell *c , struct sched *s ) {
     double pivot[3];
     
     /* Add a resource for this cell if it doesn't have one yet. */
-    if ( c->res == sched_res_none ) 
-        c->res = sched_addres( s , sched_res_none );
+    if ( c->res == qsched_res_none ) 
+        c->res = qsched_addres( s , qsched_res_none );
     
     /* Attach a center-of-mass task to the cell. */
-    c->com_tid = sched_newtask( s , task_type_com , 0 , 0 , &c , sizeof(struct cell *) , 1 );
+    c->com_tid = qsched_newtask( s , task_type_com , task_flag_none , &c , sizeof(struct cell *) , 1 );
     
     /* Does this cell need to be split? */
     if ( count > cell_maxparts ) {
@@ -152,7 +152,7 @@ void cell_split ( struct cell *c , struct sched *s ) {
             cp->h[0] = c->h[0]/2;
             cp->h[1] = c->h[1]/2;
             cp->h[2] = c->h[2]/2;
-            cp->res = sched_addres( s , c->res );
+            cp->res = qsched_addres( s , c->res );
             if ( k & 4 )
                 cp->loc[0] += cp->h[0];
             if ( k & 2 )
@@ -223,7 +223,7 @@ void cell_split ( struct cell *c , struct sched *s ) {
             
         /* Link the COM tasks. */
         for ( k = 0 ; k < 8 ; k++ )
-            sched_addunlock( s , c->progeny[k]->com_tid , c->com_tid );
+            qsched_addunlock( s , c->progeny[k]->com_tid , c->com_tid );
             
         } /* does the cell need to be split? */
         
@@ -483,7 +483,7 @@ void iact_self ( struct cell *c ) {
  * @param cj The second #cell.
  */
  
[…]

### 2013-10-28 examples/test_qr.c

@@ -396,7 +396,7 @@ void test_qr ( int m , int n , int nr_threads ) {
 
     int k, j, i;
     double *A, *A_orig, *tau;
-    struct sched s;
+    struct qsched s;
     int *tid, *rid, tid_new;
     int data[3];
     
@@ -422,7 +422,7 @@ void test_qr ( int m , int n , int nr_threads ) {
     printf( "];\n" ); */
     
     /* Initialize the scheduler. */
-    sched_init( &s , nr_threads , m*n );
+    qsched_init( &s , nr_threads , m*n );
     
     /* Allocate and init the task ID and resource ID matrix. */
     if ( ( tid = (int *)malloc( sizeof(int) * m * n ) ) == NULL ||
@@ -430,7 +430,7 @@ void test_qr ( int m , int n , int nr_threads ) {
         error( "Failed to allocate tid/rid matrix." );
     for ( k = 0 ; k < m * n ; k++ ) {
         tid[k] = -1;
-        rid[k] = sched_addres( &s , -1 );
+        rid[k] = qsched_addres( &s , -1 );
         }
     
     /* Build the tasks. */
@@ -438,21 +438,21 @@ void test_qr ( int m , int n , int nr_threads ) {
     
         /* Add kth corner task. */
         data[0] = k; data[1] = k; data[2] = k;
-        tid_new = sched_newtask( &s , task_DGEQRF , 0 , 0 , data , sizeof(int)*3 , 2 );
-        sched_addlock( &s , tid_new , rid[ k*m + k ] );
+        tid_new = qsched_newtask( &s , task_DGEQRF , task_flag_none , data , sizeof(int)*3 , 2 );
+        qsched_addlock( &s , tid_new , rid[ k*m + k ] );
         if ( tid[ k*m + k ] != -1 )
-            sched_addunlock( &s , tid[ k*m + k ] , tid_new );
+            qsched_addunlock( &s , tid[ k*m + k ] , tid_new );
         tid[ k*m + k ] = tid_new;
             
         /* Add column tasks on kth row. */
         for ( j = k+1 ; j < n ; j++ ) {
             data[0] = k; data[1] = j; data[2] = k;
-            tid_new = sched_newtask( &s , task_DLARFT , 0 , 0 , data , sizeof(int)*3 , 3 );
-            sched_addlock( &s , tid_new , rid[ j*m + k ] );
-            sched_adduse( &s , tid_new , rid[ k*m + k ] );
-            sched_addunlock( &s , tid[ k*m + k ] , tid_new );
[…]

### 2013-10-28 src/qsched.c
more specific type names.

@@ -22,6 +22,16 @@
 #include <stdlib.h>
 #include <string.h>
 
+/* OpenMP headers, only if available. */
+#ifdef HAVE_OPENMP
+    #include <omp.h>
+#endif
+
+/* Pthread headers, only if available. */
+#ifdef HAVE_PTHREAD
+    #include <pthread.h>
+#endif
+
 /* Local includes. */
 #include "cycle.h"
 #include "atomic.h"
@@ -37,9 +47,91 @@
 
 
 /**
+ * @brief Execute all the tasks in the current scheduler.
+ *
+ * @param s Pointer to the #qsched.
+ * @param nr_threads Number of threads to use.
+ * @param fun User-supplied function that will be called with the
+ *        task type and a pointer to the task data.
+ *
+ * This function is only available if QuickSched was compiled with
+ * OpenMP support.
+ */
+ 
+void qsched_run ( struct qsched *s , int nr_threads , qsched_funtype fun ) {
+
+#if defined( HAVE_OPENMP )
+
+    /* Parallel loop. */
+    #pragma omp parallel num_threads( nr_threads )
+    {
+        /* Local variable. */
+        struct task *t;
+        
+        /* Get the ID of the current thread. */
+        int qid = omp_get_thread_num() % s->nr_queues;
+        
+        /* Loop as long as there are tasks. */
+        while ( ( t = qsched_gettask( s , qid ) ) != NULL ) {
+        
+            /* Call the user-supplied function on the task with its data. */
+            fun( t->type , t->data );
+            
+            /* Mark that task as done. */
+            qsched_done( s , t );
+            
+            } /* loop as long as there are tasks. */
+            
+        } /* parallel loop. */
+        
+#elif defined( HAVE_PTHREAD )
+
+    pthread_t threads[ nr_threads ];
+    int k;
+        
+    /* The runner function. */
+    void runner ( void *data ) {
+    
+        /* Local variable. */
+        struct task *t;
+        
+        /* Extract the runner's qid. */
+        int qid = (int)data;
+        
+        /* Loop as long as there are tasks. */
+        while ( ( t = qsched_gettask( s , qid ) ) != NULL ) {
+        
+            /* Call the user-supplied function on the task with its data. */
+            fun( t->type , t->data );
+            
+            /* Mark that task as done. */
[…]

### 2013-10-28 src/qsched.h

@@ -21,6 +21,7 @@
 #define qsched_flag_none                 0
 #define qsched_flag_dirty                1
 #define qsched_flag_ready                2
+#define qsched_flag_spin                 4
 
 /* Some sched-specific constants. */
 #define qsched_stretch                   1.41
@@ -32,6 +33,18 @@
 #define qsched_init_datapertask          2
 #define qsched_data_round                16
 #define qsched_res_none                  (-1)
+
+
+/** Type used for task identifiers. */
+typedef int qsched_task_t;
+
+
+/** Type used for resource identifiers. */
+typedef int qsched_res_t;
+
+
+/** Type definition for the execution function in #qsched_run. */
+typedef void (*qsched_funtype)( int , void * );
 
 
 /* The sched data structre. */
@@ -109,21 +122,24 @@ struct qsched {
 
 
 /* Function prototypes. */
-void qsched_init ( struct qsched *s , int nr_queues , int size );
+/* Internal functions. */
 void qsched_sort ( int *restrict data , int *restrict ind , int N , int min , int max );
 void qsched_sort_rec ( int *restrict data , int *restrict ind , int N , int min , int max );
-void qsched_prepare ( struct qsched *s );
-int qsched_addres ( struct qsched *s , int parent );
-void qsched_addlock ( struct qsched *s , int t , int res );
-void qsched_addunlock ( struct qsched *s , int ta , int tb );
-int qsched_newtask ( struct qsched *s , int type , unsigned int flags , void *data , int data_size , int cost );
 struct task *qsched_gettask ( struct qsched *s , int qid );
-void qsched_adduse ( struct qsched *s , int t , int res );
 void qsched_done ( struct qsched *s , struct task *t );
 void *qsched_getdata( struct qsched *s , struct task *t );
-void qsched_free ( struct qsched *s );
 int qsched_lockres ( struct qsched *s , int rid );
 void qsched_unlockres ( struct qsched *s , int rid );
 int qsched_locktask ( struct qsched *s , int tid );
 void qsched_unlocktask ( struct qsched *s , int tid );
 
+/* External functions. */
+void qsched_init ( struct qsched *s , int nr_queues , int size );
+void qsched_prepare ( struct qsched *s );
+qsched_res_t qsched_addres ( struct qsched *s , qsched_res_t parent );
+void qsched_addlock ( struct qsched *s , qsched_task_t t , qsched_res_t res );
[…]

### 2013-10-28 examples/test.c
use more specific type names.

@@ -61,7 +61,9 @@ void matmul ( int m , int n , int k , double *a , int lda , double *b , int ldb 
  
 void test2 ( int m , int n , int k , int nr_threads ) { 
 
-    int i, j, kk, qid, data[3], *d, tid, rid;
+    int i, j, kk, qid, data[3], *d;
+    qsched_task_t tid;
+    qsched_res_t rid;
     struct qsched s;
     struct task *t;
     double *a, *b, *c, *res, err = 0.0, irm = 1.0/RAND_MAX;
@@ -176,7 +178,9 @@ void test2 ( int m , int n , int k , int nr_threads ) {
  
 void test1 ( int m , int n , int k , int nr_threads ) { 
 
-    int i, j, qid, data[2], *d, tid, rid;
+    int i, j, qid, data[2], *d;
+    qsched_task_t tid;
+    qsched_res_t rid;
     struct qsched s;
     struct task *t;
     double *a, *b, *c, *res, err = 0.0, irm = 1.0/RAND_MAX;

### 2013-10-28 examples/test_bh.c

@@ -485,7 +485,8 @@ void iact_self ( struct cell *c ) {
  
 void create_tasks ( struct qsched *s , struct cell *ci , struct cell *cj ) {
 
-    int j, k, tid;
+    int j, k;
+    qsched_task_t tid;
     struct cell *data[2];
     double dx, r2;
     

### 2013-10-28 examples/test_qr.c

@@ -397,7 +397,8 @@ void test_qr ( int m , int n , int nr_threads ) {
     int k, j, i;
     double *A, *A_orig, *tau;
     struct qsched s;
-    int *tid, *rid, tid_new;
+    qsched_task_t *tid, tid_new;
+    qsched_res_t *rid;
     int data[3];
     
     enum task_types { task_DGEQRF , task_DLARFT , task_DTSQRF , task_DSSRFT };
@@ -425,8 +426,8 @@ void test_qr ( int m , int n , int nr_threads ) {
     qsched_init( &s , nr_threads , m*n );
     
     /* Allocate and init the task ID and resource ID matrix. */
-    if ( ( tid = (int *)malloc( sizeof(int) * m * n ) ) == NULL ||
-         ( rid = (int *)malloc( sizeof(int) * m * n ) ) == NULL )
+    if ( ( tid = (qsched_task_t *)malloc( sizeof(qsched_task_t) * m * n ) ) == NULL ||
+         ( rid = (qsched_res_t *)malloc( sizeof(qsched_res_t) * m * n ) ) == NULL )
         error( "Failed to allocate tid/rid matrix." );
     for ( k = 0 ; k < m * n ; k++ ) {
         tid[k] = -1;

### 2013-10-28 paper/paper.tex
backup.

@@
 
-\documentclass[prodmode,acmtecs]{acmsmall} % Aptara syntax
+\documentclass[prodmode,acmtopc]{acmsmall} % Aptara syntax
 
@@
 
-Conflicts can be modeled as exclusive locks on a shared resource
-which have to be obtained by a task before it can execute.
-Thus, in \fig{TaskConflicts}, before executing, task $F$ has
-to obtain an exclusive lock on the resource associated with
-the conflict between tasks $F$, $H$, and $I$.
-While task $F$ is being executed, neither $H$ nor $I$ can 
-lock the same resource, and therefore will not execute until
-task $F$ is done and the lock has been released.
-
 In dependency-only systems, such conflicts can be modelled
@@
 
+Conflicts can be modeled directly as exclusive locks on a shared resource
+which have to be obtained by a task before it can execute.
+Thus, in \fig{TaskConflicts}, before executing, task $F$ has
+to obtain an exclusive lock on the resource associated with
+the conflict between tasks $F$, $H$, and $I$.
+While task $F$ is being executed, neither $H$ nor $I$ can 
+lock the same resource, and therefore will not execute until
+task $F$ is done and the lock has been released.
+
 This paper presents QuickSched, a framework for task-based
 parallel programming with constraints.
+In order to be of practical use, and to scale well with modern
+multi-core shared-memory architectures, the task scheduler
+must be:
+\begin{itemize}
+    \item {\em Correct}: All constraints, i.e.~dependencies and
+        conflicts, must be correctly enforced,
+    \item {\em Fast}: The overheads associated with task management
+        should be as small as possible,
+    \item {\em Memory/cache-efficient}: Tasks accessing similar
+        sets of data should be preferentially executed on the
+        same core to preserve memory/cache locality as far as possible, and
+    \item {\em Parallel-efficient}: Tasks should be executed in an order
+        that sufficient work is available for all computatoinal
+        threads at all times.
+\end{itemize}
+\noindent This paper presents QuickSched, a framework for task-based
+parallel programming with constraints, which aims to achieve
+all of these goals.
 Section~2 describes the underlying algorithms and data structures,
@@
 Section~4 presents two test-cases:
-\begin{itemize}
+\begin{enumerate}
     \item The tiled QR
@@
     gravitational N-body problem,
-\end{itemize}
+\end{enumerate}
 These real-world examples show how QuickSched can be used in practice,
@@
 \subsection{Tasks}
+\label{sec:tasks}
 
@@
 \end{minipage}\end{center}
-\noindent where\dots
+\noindent where\dots the only additional field {\tt waiting} is
+used to keep track of the number of tasks that have not been
+executed.
 Note that for brevity, and to avoid conflicts with the naming
@@
 
+The tasks are executed as follows:
+\begin{center}\begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}
+void qsched_run ( qsched *s , void (*fun)( int , void * ) ) {
+    qsched_start( s );
+    #pragma omp parallel
+    {
+        int qid = omp_get_thread_num() % s->nr_queues;
+        struct task *t;
+        while ( ( t = qsched_gettask( s , qid ) ) != NULL ) {
+            fun( t->type , t->data );
+            qsched_done( s , t );
+            }
+        }
+    }
+    \end{lstlisting}
+\end{minipage}\end{center}
+\noindent where {\tt qsched\_start} initializes the tasks and
+fills the queues (line~1).
+For simplicity, OpenMP \cite{ref:Dagum1998}, which is available
+for most compilers, is used to create a parallel section
+in which the code between lines~4 and~12 is executed
+concurrently.
+The parallel section consists of a loop (lines~7--10) in
+which a task is acquired via the function {\tt qsched\_gettask}
+and its type and data are passed to a user-supplied execution
+function.
+Once the task has been executed, it is returned to the
+scheduler via the function {\tt qsched\_done}.
+The loop terminates when the scheduler runs out of tasks,
+i.e.~when {\tt qstack\_gettask} returns {\tt NULL}, and
+the function exits once all the threads have exited their
+loops.
+
+At the start of a parallel computation, {\tt qsched\_start}
+identifies the tasks that have no dependencies and sends them
+to queues:
 \begin{center}\begin{minipage}{0.9\textwidth}
@@
     for ( k = 0 ; k < s->nr_tasks && s->tasks[tid[k]].wait == 0 ; k++ )
-        qsched_enqueue( s , tid[k] );
+        qsched_enqueue( s , &s->tasks[tid[k]] );
     }
@@
 \end{minipage}\end{center}
+\noindent where the operations in line~2 are described in \sect{tasks}.
 
+The function {\tt qsched\_enqueue} tries to identify the best
+queue for a given task by looking at which queues last used
+the resources used and locked by the task, e.g.:
 \begin{center}\begin{minipage}{0.9\textwidth}
@@
 \end{minipage}\end{center}
+\noindent where the array {\tt score} keeps a count of the
+task resources ``owned'', or last used, by each queue.
+In lines~9--12 the queue with the highest such score is 
+chosen on which the task is then put (line~13).
 
+The function {\tt qsched\_gettask} fetches a task from
+one of the queues:
 \begin{center}\begin{minipage}{0.9\textwidth}
@@
 struct task *qsched_gettask ( qsched *s , int qid ) {
-    struct task *res;
+    struct task *res = NULL;
     int k;
@@
 \end{minipage}\end{center}
+\noindent where the parameter {\tt qid} is the index of the
+prefered queue.
+If the queue is empty, or all of the tasks in that queue had
+unresolved conflicts, the scheduler uses {\em work stealing}
+\cite{ref:Blumofe1999}, i.e.~it looks for the queue with the
+largest task count, and attempts to steal a task from it
+(line~7).
+If a task could be obtained from any queue (line~13),
+the resources it locks and uses are marked as now being owned
+by the prefered queue (lines~14--17).
+Finally, the task, or {\tt NULL} if no task could be obtained,
+is returned.
 
+The final step in a task's lifecycle is, on completion,
+to unlock the tasks which depend on it.
+This is handled by the function {\tt qsched\_done}:
 \begin{center}\begin{minipage}{0.9\textwidth}
     \begin{lstlisting}
-struct task *qsched_done ( qsched *s , struct task *t ) {
+void qsched_done ( qsched *s , struct task *t ) {
     int k;
@@
 \end{minipage}\end{center}
-
-\begin{center}\begin{minipage}{0.9\textwidth}
-    \begin{lstlisting}
-struct task *qsched_run ( qsched *s , void (*fun)( int , void * ) ) {
-    #pragma omp parallel
-    {
-        int qid = omp_get_thread_num() % s->nr_queues;
-        struct task *t;
-        while ( ( t = qsched_gettask( s , qid ) ) != NULL ) {
-            fun( t->type , t->data );
-            qsched_done( s , t );
-            }
-        }
-    }
-    \end{lstlisting}
-\end{minipage}\end{center}
+\noindent If any of the unlocked tasks' wait counters
+goes to zero (line~4), then the unlocked task is ready to
+run and is immediately dispatched via {\tt qsched\_enqueue},
+as described earlier.
+Once all the dependent tasks have been unlocked, the
+{\tt waiting} counter is decremented (line~6).
 
@@
 kernels in the task-based QR decomposition as part of his MSc Thesis
-at Durham University, as well as Lydia Heck of the Institute for
+at Durham University.
+The author also thanks Lydia Heck of the Institute for
 Computational Cosmology at Durham University for providing access
-to, and expertise on, the COSMA cluster.
+to, and expertise on, the COSMA cluster used in the performance
+evaluation.
 \end{acks}

### 2013-10-28 src/qsched.c
use flags for phtread/yield.

@@ -17,6 +17,9 @@
  * 
  ******************************************************************************/
 
+/* Config parameters. */
+#include "../config.h"
+
 /* Standard includes. */
 #include <stdio.h>
 #include <stdlib.h>
@@ -27,7 +30,7 @@
     #include <omp.h>
 #endif
 
-/* Pthread headers, only if available. */
+// /* Pthread headers, only if available. */
 #ifdef HAVE_PTHREAD
     #include <pthread.h>
 #endif
@@ -42,12 +45,30 @@
 #include "qsched.h"
 #include "queue.h"
 
-/* Max macro. */
-#define max(a,b) ( ((a) >= (b)) ? (a) : (b) )
+
+/**
+ * @brief Clear the tasks and resources in a scheduler.
+ *
+ * @param s Pointer to the #qsched.
+ */
+ 
+void qsched_reset ( struct qsched *s ) {
+
+    /* Simply clear the counts, leave the buffers allocated. */
+    s->count = 0;
+    s->waiting = 0;
+    s->count_data = 0;
+    s->count_deps = 0;
+    s->count_locks = 0;
+    s->count_uses = 0;
+    s->count_res = 0;
+
+    }
 
 
 /**
- * @brief Execute all the tasks in the current scheduler.
+ * @brief Execute all the tasks in the current scheduler using
+ *        OpenMP.
  *
  * @param s Pointer to the #qsched.
  * @param nr_threads Number of threads to use.
@@ -58,9 +79,12 @@
  * OpenMP support.
  */
  
-void qsched_run ( struct qsched *s , int nr_threads , qsched_funtype fun ) {
+void qsched_run_openmp ( struct qsched *s , int nr_threads , qsched_funtype fun ) {
 
 #if defined( HAVE_OPENMP )
+
+    /* Prepare the scheduler. */
+    qsched_prepare( s );
 
     /* Parallel loop. */
     #pragma omp parallel num_threads( nr_threads )
@@ -75,7 +99,7 @@ void qsched_run ( struct qsched *s , int nr_threads , qsched_funtype fun ) {
         while ( ( t = qsched_gettask( s , qid ) ) != NULL ) {
         
             /* Call the user-supplied function on the task with its data. */
-            fun( t->type , t->data );
+            fun( t->type , &s->data[ t->data ] );
             
             /* Mark that task as done. */
             qsched_done( s , t );
[…]

### 2013-10-28 src/qsched.h

@@ -21,18 +21,20 @@
 #define qsched_flag_none                 0
 #define qsched_flag_dirty                1
 #define qsched_flag_ready                2
-#define qsched_flag_spin                 4
+#define qsched_flag_yield                4
+#define qsched_flag_pthread              8
 
 /* Some sched-specific constants. */
-#define qsched_stretch                   1.41
-#define qsched_inc                       10
+#define qsched_stretch                   2
+#define qsched_size_init                 1000
 #define qsched_init_depspertask          2
 #define qsched_init_lockspertask         2
 #define qsched_init_usespertask          2
 #define qsched_init_respertask           2
-#define qsched_init_datapertask          2
+#define qsched_init_datapertask          8
 #define qsched_data_round                16
 #define qsched_res_none                  (-1)
+#define qsched_task_none                 (-1)
 
 
 /** Type used for task identifiers. */
@@ -117,7 +119,13 @@ struct qsched {
     
     /* A lock for the sched itself. */
     lock_type lock;
-
+    
+    /* Pthread stuff for condition variable for yielding threads. */
+    #ifdef HAVE_PTHREAD
+        pthread_cond_t cond;
+        pthread_mutex_t mutex;
+    #endif
+    
     };
 
 
@@ -132,14 +140,15 @@ int qsched_lockres ( struct qsched *s , int rid );
 void qsched_unlockres ( struct qsched *s , int rid );
 int qsched_locktask ( struct qsched *s , int tid );
 void qsched_unlocktask ( struct qsched *s , int tid );
+void qsched_prepare ( struct qsched *s );
 
 /* External functions. */
-void qsched_init ( struct qsched *s , int nr_queues , int size );
-void qsched_prepare ( struct qsched *s );
+void qsched_init ( struct qsched *s , int nr_queues , int flags );
 qsched_res_t qsched_addres ( struct qsched *s , qsched_res_t parent );
 void qsched_addlock ( struct qsched *s , qsched_task_t t , qsched_res_t res );
 void qsched_addunlock ( struct qsched *s , qsched_task_t ta , qsched_task_t tb );
-qsched_task_t qsched_newtask ( struct qsched *s , int type , unsigned int flags , void *data , int data_size , int cost );
+qsched_task_t qsched_addtask ( struct qsched *s , int type , unsigned int flags , void *data , int data_size , int cost );
 void qsched_adduse ( struct qsched *s , qsched_task_t t , qsched_res_t res );
[…]

### 2013-10-28 src/queue.c

@@ -17,6 +17,9 @@
  * 
  ******************************************************************************/
 
+/* Config parameters. */
+#include "../config.h"
+
 /* Standard includes. */
 #include <stdio.h>
 #include <stdlib.h>

### 2013-10-28 examples/test.c
use flags for phtread/yield.

@@ -18,6 +18,9 @@
  ******************************************************************************/
 
 
+/* Config parameters. */
+#include "../config.h"
+
 /* Standard includes. */
 #include <stdio.h>
 #include <stdlib.h>
@@ -61,13 +64,33 @@ void matmul ( int m , int n , int k , double *a , int lda , double *b , int ldb 
  
 void test2 ( int m , int n , int k , int nr_threads ) { 
 
-    int i, j, kk, qid, data[3], *d;
+    int i, j, kk, data[3];
     qsched_task_t tid;
     qsched_res_t rid;
     struct qsched s;
-    struct task *t;
     double *a, *b, *c, *res, err = 0.0, irm = 1.0/RAND_MAX;
     ticks tic_task, toc_task, tic_ref, toc_ref;
+    
+    
+    /* Runner function to pass to the scheduler. */
+    void runner ( int type , void *data ) {
+    
+        /* Decode the task data. */
+        int *d = (int *)data;
+        
+        /* Decode and execute the task. */
+        switch ( type ) {
+            case 1:
+                // message( "thread %i working on block [ %i , %i ] with k=%i, lock[0]=%i." , qid , d[0] , d[1] , d[2] , t->locks[0] ); fflush(stdout);
+                matmul( 32 , 32 , 32 , &a[ d[2]*32*m*32 + d[0]*32 ] , m*32 , &b[ k*32*d[1]*32 + d[2]*32 ] , k*32 , &c[ d[0]*32 + m*32*d[1]*32 ] , m*32 );
+                break;
+            default:
+                error( "Unknown task type." );
+            }
+
+        }
+        
+    
     
     /* Tell the user something about the test. */
     message( "computing a tiled matrix multiplication of the form "
@@ -75,7 +98,7 @@ void test2 ( int m , int n , int k , int nr_threads ) {
     
     /* Init the sched. */
     bzero( &s , sizeof(struct qsched) );
-    qsched_init( &s , nr_threads , m * n );
+    qsched_init( &s , nr_threads, qsched_flag_none );
 
     /* Allocate the matrices. */
     if ( ( a = (double *)malloc( sizeof(double) * m * k * 32 * 32 ) ) == NULL ||
@@ -99,50 +122,18 @@ void test2 ( int m , int n , int k , int nr_threads ) {
             data[0] = i; data[1] = j;
             for ( kk = 0 ; kk < k ; kk++ ) {
                 data[2] = kk;
[…]

### 2013-10-28 examples/test_bh.c

@@ -18,6 +18,9 @@
  ******************************************************************************/
 
 
+/* Config parameters. */
+#include "../config.h"
+
 /* Standard includes. */
 #include <stdio.h>
 #include <stdlib.h>
@@ -134,7 +137,7 @@ void cell_split ( struct cell *c , struct qsched *s ) {
         c->res = qsched_addres( s , qsched_res_none );
     
     /* Attach a center-of-mass task to the cell. */
-    c->com_tid = qsched_newtask( s , task_type_com , task_flag_none , &c , sizeof(struct cell *) , 1 );
+    c->com_tid = qsched_addtask( s , task_type_com , task_flag_none , &c , sizeof(struct cell *) , 1 );
     
     /* Does this cell need to be split? */
     if ( count > cell_maxparts ) {
@@ -520,7 +523,7 @@ void create_tasks ( struct qsched *s , struct cell *ci , struct cell *cj ) {
             data[0] = ci; data[1] = NULL;
             
             /* Create the task. */
-            tid = qsched_newtask( s , task_type_self , task_flag_none , data , sizeof(struct cell *) * 2 , ci->count*ci->count/2 );
+            tid = qsched_addtask( s , task_type_self , task_flag_none , data , sizeof(struct cell *) * 2 , ci->count*ci->count/2 );
             
             /* Add the resource. */
             qsched_addlock( s , tid , ci->res );
@@ -560,13 +563,13 @@ void create_tasks ( struct qsched *s , struct cell *ci , struct cell *cj ) {
             
                 /* Interact ci's parts with cj as a cell. */
                 data[0] = ci; data[1] = cj;
-                tid = qsched_newtask( s , task_type_pair_pc , task_flag_none , data , sizeof(struct cell *) * 2 , ci->count );
+                tid = qsched_addtask( s , task_type_pair_pc , task_flag_none , data , sizeof(struct cell *) * 2 , ci->count );
                 qsched_addlock( s , tid , ci->res );
                 qsched_addunlock( s , cj->com_tid , tid );
         
                 /* Interact cj's parts with ci as a cell. */
                 data[0] = cj; data[1] = ci;
-                tid = qsched_newtask( s , task_type_pair_pc , task_flag_none , data , sizeof(struct cell *) * 2 , ci->count );
+                tid = qsched_addtask( s , task_type_pair_pc , task_flag_none , data , sizeof(struct cell *) * 2 , ci->count );
[…]

### 2013-10-28 examples/test_qr.c

@@ -18,6 +18,9 @@
  ******************************************************************************/
 
 
+/* Config parameters. */
+#include "../config.h"
+
 /* Standard includes. */
 #include <stdio.h>
 #include <stdlib.h>
@@ -402,6 +405,37 @@ void test_qr ( int m , int n , int nr_threads ) {
     int data[3];
     
     enum task_types { task_DGEQRF , task_DLARFT , task_DTSQRF , task_DSSRFT };
+
+
+    /* Runner function to pass to the scheduler. */
+    void runner ( int type , void *data ) {
+    
+        /* Decode the task data. */
+        int *idata = (int *)data;
+        int i = idata[0], j = idata[1], k = idata[2];
+        double buff[ 2*32*32 ];
+        
+        /* Decode and execute the task. */
+        switch ( type ) {
+            case task_DGEQRF:
+                DGEQRF( &A[ j*m*32*32 + i*32 ] , &tau[ j*m*32 + i*32 ] , 32 , 32 , 32*m , buff );
+                break;
+            case task_DLARFT:
+                DLARFT( &A[ j*m*32*32 + i*32 ] , &A[ i*m*32*32 + i*32 ] , &tau[ i*m*32 + i*32 ] , 32 , 32 , 32*m );
+                break;
+            case task_DTSQRF:
+                DTSQRF( &A[ j*m*32*32 + j*32 ] , &A[ j*m*32*32 + i*32 ] , &tau[ j*m*32 + i*32 ] , 32 , 32 , 32 , 32*m , buff );
+                break;
+            case task_DSSRFT:
+                DSSRFT(	&A[ k*m*32 + i*32 ] , &A[ j*m*32*32 + k*32 ] , &A[ j*m*32*32 + i*32 ] , &tau[ k*m*32 + i*32 ] , 32 , 32 , 32*m );
+                break;
+            default:
+                error( "Unknown task type." );
+            }
+
+        }
+        
+    
     
     /* Allocate and fill the original matrix. */
     if ( ( A = (double *)malloc( sizeof(double) * m * n * 32 * 32 ) ) == NULL ||
@@ -423,7 +457,7 @@ void test_qr ( int m , int n , int nr_threads ) {
     printf( "];\n" ); */
     
     /* Initialize the scheduler. */
-    qsched_init( &s , nr_threads , m*n );
+    qsched_init( &s , nr_threads , qsched_flag_none );
     
     /* Allocate and init the task ID and resource ID matrix. */
[…]

### 2013-10-29 paper/paper.tex
add new figures, backup of text.

@@
 \usepackage{epsfig}
+%\usepackage{hyperref}
 \usepackage{listings}
@@
 \end{minipage}\end{center}
+\noindent where the {\tt data}, {\tt unlocks}, {\tt locks},
+and {\tt uses} arrays are pointers to the contents of other
+arrays, i.e.~they are not allocated individually.
 
@@
 The parallel section consists of a loop (lines~7--10) in
-which a task is acquired via the function {\tt qsched\_gettask}
+which a task is acquired via the {\em execution function}
+{\tt qsched\_gettask}
 and its type and data are passed to a user-supplied execution
@@
 void qsched_start ( qsched *s ) {
+  sort out pointers in tasks.
   set the task waits and weights.
@@
 \end{minipage}\end{center}
-\noindent where the operations in line~2 are described in \sect{tasks}.
+\noindent where line~2 sets the {\tt unlocks}, {\tt locks},
+and {\tt uses} pointers in the tasks.
+The operations in line~3 are described in \sect{tasks}.
 
@@
     if ( ( res = queue_get( s->queues[qid] ) ) == NULL ) {
-            k = id of queue with largest count.
+      loop over all other queues in random order with index k
         if ( ( res = queue_get( s->queues[k] ) ) != NULL )
@@
 unresolved conflicts, the scheduler uses {\em work stealing}
-\cite{ref:Blumofe1999}, i.e.~it looks for the queue with the
-largest task count, and attempts to steal a task from it
+\cite{ref:Blumofe1999}, i.e.~it loops over all other queues
+in a random order (line~6) and tries to get a task from them
 (line~7).
@@
 
+The algorithms, as described in the previous section, have
+all been implemented as part of the Open-Source C-language
+QuickSched library.\footnote{\url{http://sourceforge.net/projects/quicksched/}}
+This section describes the interface functions and how they
+are called, whereas the following section contains
+examples of how QuickSched can be used.
+
+As mentioned previously, the {\tt qsched} object is the main
+interface to the task scheduler.
+As such, it provieds functionality for task and resource
+creation, for assigning resources to tasks, either as locks
+or uses, and for assigning dependencies between tasks.
+The tasks and resources themselves are opaque to the
+user, and handles of the types {\tt qsched\_task\_t}
+and {\tt qsched\_res\_t} are used instead.
+
+The main functions for setting up the scheduler are:
+\begin{itemize}
+    \item {\tt void qsched\_init( struct qsched *s , int nr\_queues , int flags )} \\
+        Initializes a {\tt qsched} object with the given number of queues.
+        The {\tt flags} parameter can be set to any bitwise or combination
+        of {\tt qsched\_flag\_none},
+        {\tt qsched\_flag\_yield}, and {\tt qsched\_flag\_pthread},
+        which are described further below.
+        This function must be called before any of the other
+        functions are used.
+        \vspace{1mm}
+    \item {\tt void qsched\_free( struct qsched *s )} \\
+        Releases all the memory and other resources allocated by the
+        given {\tt qsched} object.
+        After this function has been called, the {\tt qsched} will
+        need to be re-initialized for reuse.
+        \vspace{1mm}
+    \item {\tt void qsched\_reset( struct qsched *s )} \\
+        Clears the tasks and resources in the given {\tt qsched},
+        but does not release the allocated memory or change
+        the number of queues.
+        \vspace{1mm}
+    \item {\tt qsched\_task\_t qsched\_addtask( struct qsched *s , int type , unsigned int flags , void *data , int data\_size , int cost )} \\
+        Creates a new task within the given {\tt qsched} and returns
+        its handle.
+        The {\tt type} and {\tt data} field are copied into the {\tt qsched}
+        and passed to the execution function when the {\tt qsched} is run.
+        The parameter flags is either {\tt task\_flag\_none} or
+        {\tt task\_flag\_virtual}.
+        Tasks marked as virtual do not have any action associated with them,
+        e.g. they are used only to group or otherwise dependencies, and
+        are not passed to the execution function in {\tt qsched\_run}.
+        \vspace{1mm}
+    \item {\tt qsched\_res\_t qsched\_addres( struct qsched *s , qsched\_res\_t parent )} \\
+        Creates a new resource within the given {\tt qsched} and returns
+        its handle.
+        The {\tt parent} field is the handle of the heirarchical parent of
+        the new resource or {\tt qsched\_res\_none} if the resource
+        has no hierarchical parent.
+        \vspace{1mm}
+    \item {\tt void qsched\_addlock( struct qsched *s , qsched\_task\_t t , qsched\_res\_t res )} \\
+        Append the resource {\tt res} to the task {\tt t}'s list of
+        locks.
+        The task {\tt t} will then conflict with any other task that
+        also locks the resource {\tt res}, its hierarchical parents, or
+        any resource hierarchically below it.
+        \vspace{1mm}
+    \item {\tt void qsched\_adduse( struct qsched *s , qsched\_task\_t t , qsched\_res\_t res )} \\
+        Similar to {\tt qsched\_addlock}, yet the resource is only used and
+        is not part of a conflict.
+        This information is used when assinging tasks to specific queues.
+        \vspace{1mm}
+    \item {\tt void qsched\_addunlock( struct qsched *s , qsched\_task\_t ta , qsched\_task\_t tb )} \\
+        Appends the task {\tt tb} to the list of tasks that the task {\tt ta}
+        unlocks.
+        The task {\tt tb} then depends on the task {\tt ta}.
+        \vspace{1mm}
+    \item {\tt void qsched\_run( struct qsched *s , int nr\_threads , qsched\_funtype fun )} \\
+        Executes the tasks in the given {\tt qsched} using {\tt nr\_threads}
+        threads via the execution function {\tt fun}, as described in the
+        previous section.
+        Once a {\tt qsched} has been initialized and filled with
+        tasks and resources, it can be run more than once.
+        \vspace{1mm}
+\end{itemize}
+
+The library can be compiled to use either OpenMP or the
+{\tt pthreads} library \cite{ref:Pthreads1995}.
+OpenMP is the default, but calling {\tt qsched\_init} with
+either the {\tt qsched\_flag\_yield}
+or the {\tt qsched\_flag\_pthread} switches to using {\tt pthreads}
+for the parallel loop.
+
+OpenMP has the advantage of being available for most compilers
+and also potentially providing some extra platform-specific
+scheduling features, e.g.~optimal thread location and/or affinity,
+and of integrating seamlessly with other parallel parts of the
+user application.
+The disadvantage of using OpenMP is that it does not provide
+any mechanism for yielding a thread if no tasks are available,
+i.e. the main loop in {\tt qsched\_gettask}, described in the
+previous section, will spin until a task becomes available.
+This may be a problem if other parts of the user application
+are running in the background simultaneously.
+Calling {\tt qsched\_init} with the {\tt qsched\_flag\_yield}
+forces the use of {\tt pthreads} and uses conditional variables
+to wait for a new task to be enqueued if obtaining a task
+from any of the queues fails.
+This relinquishes the waiting computational thread for other
+processes.
+
 \section{Validation}
 
+This section presents two test cases showing both
+how QuickSched can be used in real-world applications, and
+providing benchmarks to assess its efficiency and scalability.
+
+The source code of both examples is distributed with the
+QuickSched library.
+
 \subsection{Task-Based QR Decomposition}
 
+\citeN{ref:Buttari2009} introduced the concept of using task-based
+parallelsim for tile-based algorithms in numerical linear algebra,
+presenting parallel codes for the Cholesky, LU, and QR
+decompositions.
+These algorithms are now part of the PLASMA and MAGMA
+libraries for parallel linear algebra \cite{ref:Agullo2009}.
+The former uses the QUARK task scheduler, which was originally
+designed for it, while the latter currently uses the StarPU
+task scheduler.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/QR.pdf,width=0.8\textwidth}}
+    \caption{Task-based QR decomposition of a matrix consisting
+        of $4\times 4$ tiles.
+        Each circle represents a tile, and its colour represents
+        the type of taks on that tile at that level.
+        Empty circles have no task associated with them.
+        The arrows represent dependencies at each level, and
+        tasks at each level also implicitly depend on the
+        task at the same location in the previous level.}
+    \label{fig:QR}
+\end{figure}
+
+The tiled QR factorization is based on four basic tasks,
+or kernels, as shown in \fig{QR}.
+For a matrix consisting of $N\times N$ tiles, $N$ passes,
+or levels, are computed, each computing a column and row of the QR
+decomposition.
+The tasks can be defined in terms of the tuples $(i,j,k)$,
+where $i$ and $j$ are the row and column of the tile, respectively,
+and $k$ is its level:
+
+\begin{center}
+    \begin{tabular}{llll}
+        Task & cond. & depends on task & locks tile \\
+        \hline
+        \epsfig{file=figures/TaskRed.pdf,height=9pt} DGEQRF & $i=j=k$ & $(i,j,k-1)$ & $(i,j)$ \\
+        \epsfig{file=figures/TaskGreen.pdf,height=9pt} DLARFT & $i=k$, $j>k$ & $(i,j,k-1)$, $(k,k,k)$ & $(i,j)$ \\
+        \epsfig{file=figures/TaskBlue.pdf,height=9pt} DTSQRF & $i>k$, $j=k$ & $(i,j,k-1)$, $(i-1,j,k)$ & $(i,j)$, $(j,j)$ \\
+        \epsfig{file=figures/TaskYellow.pdf,height=9pt} DSSRFT & $i>k$, $j>k$ & $(i,j,k-1)$, $(k,j,k)$, $(i,k,k)$ & $(i,j)$ \\
+        \hline
+    \end{tabular}
+\end{center}
+
+\noindent where the task names are the BLAS-like operation
+performed on the given tiles.
+Every task depends on the task at the same position and the
+previous level, i.e.~the task $(i,j,k)$ always depends on
+$(i,j,k-1)$ for $k>1$.
+Each task also modifies its own tile $(i,j)$, and the DTSQRF
+task additionally modifies the lower triangular part of the $(j,j)$th tile.
+
+Setting up the dependencies and locks for a matrix of
+$m\times n$ tiles is implemented as shown in \fig{CodeQR},
+
+\begin{figure}
+\begin{center}\begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}
+enum { tDGEQRF , tDLARFT , tDTSQRF , tDSSRFT };
+void make_tasks ( struct qsched *s , int m , int n ) {
+  int i, j, k, data[3];
+  qsched_task_t tid[ m*n ], tid_new;
+  qsched_res_t rid[ m*n ];
+  for ( k = 0 ; k < m*n ; k++ ) {
+    tid[k] = qsched_task_none;
+    rid[k] = qsched_addres( s , qsched_res_none );
+    }
+  for ( k = 0 ; k < m && k < n ; k++ ) {
+    /* DGEQRF task at (k,k). */
+    data[0] = k; data[1] = k; data[2] = k;
+    tid_new = qsched_addtask( s , tDGEQRF , qsched_flags_none ,
+      data , sizeof(int)*3 , 2 );
+    qsched_addlock( s , tid_new , rid[ k*m + k ] )
+    if ( tid[ k*m + k ] != qsched_task_none )
+      qsched_addunlock( s , tid[ k*m + k ] , tid_new );
+    tid[ k*m + k ] = tid_new;
+    for ( j = k+1 ; j < n ; j++ ) {
+      /* DLARFT task at (k,j). */
+      data[0] = k; data[1] = j; data[2] = k;
+      tid_new = qsched_addtask( s , tDLARFT , qsched_flags_none ,
+        data , sizeof(int)*3 , 3 );
+      qsched_addlock( s , tid_new , rid[ j*m + k ] )
+      qsched_addunlock( s , tid[ k*m + k ] , tid_new );
+      if ( tid[ j*m + k ] != qsched_task_none )
+        qsched_addunlock( s , tid[ j*m + k ] , tid_new );
+      tid[ j*m + k ] = tid_new;
+      }
+    for ( i = k+1 ; i < m ; i++ ) {
+      /* DTSQRF task at (i,k). */
+      data[0] = i; data[1] = k; data[2] = k;
+      tid_new = qsched_addtask( s , tDTSQRF , qsched_flags_none ,
+        data , sizeof(int)*3 , 3 );
+      qsched_addlock( s , tid_new , rid[ k*m + i ] )
+      qsched_addlock( s , tid_new , rid[ k*m + k ] )
+      qsched_addunlock( s , tid[ k*m + k ] , tid_new );
+      if ( tid[ k*m + i ] != qsched_task_none )
+        qsched_addunlock( s , tid[ k*m + i ] , tid_new );
+      tid[ k*m + i ] = tid_new;
+      for ( j = k+1 ; j < n ; j++ ) {
+        /* DSSRFT task at (i,j). */
+        data[0] = i; data[1] = j; data[2] = k;
+        tid_new = qsched_addtask( s , tDSSRFT , qsched_flags_none ,
+          data , sizeof(int)*3 , 5 );
+        qsched_addlock( s , tid_new , rid[ j*m + i ] )
+        qsched_addunlock( s , tid[ j*m + k ] , tid_new );
+        qsched_addunlock( s , tid[ k*m + i ] , tid_new );
+        if ( tid[ j*m + i ] != qsched_task_none )
+          qsched_addunlock( s , tid[ j*m + i ] , tid_new );
+        tid[ j*m + i ] = tid_new;
+        }
+      }
+    }
+  }
+    \end{lstlisting}
+\end{minipage}\end{center}
+\caption{Example code to generate the tasks for the tiled
+    QR decomposition.}
+\label{fig:CodeQR}
+\end{figure}
+
+Setting up the dependencies and locks for a matrix of
+$m\times n$ tiles is implemented as shown in \fig{CodeQR},
+where the $m\times n$ matrix {\tt tid} stores the handles
+of the last task at position $(i,j)$ and is initialized with
+empty tasks (line~7).
+Similarly, {\tt rid} stores the handles of the resources for each
+tile of the matrix, allocated in line~8.
+
+The following loops mirror the task generation described in
+Algorithm~2 of \citeN{ref:Buttari2009}.
+E.g.~for each level {\tt k} (line~10), a DGEQRF task is created
+for tile $(k,k)$ (lines~13--14).
+A lock is added for the newly created task on the
+resource associated with the $(k,k)$th tile (line~15).
+If a task exists at that position at the previous level
+(line~16), a dependency is added between the old task and
+the new (line~17), and the new task is stored in {\tt tid}
+(line~18).
+The remaining tasks are generated in the same way, with
+their respective locks and dependencies.
+
+The execution function for these tasks simply calls the apropriate
+kernels on the matrix tiles given by the task data:
+\begin{center}\begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}
+void exec_fun ( int type , void *data ) {
+  int *idata = (int *)data;
+  int i = idata[0], j = idata[1], k = idata[2];
+  switch ( type ) {
+    case tDGEQRF:
+      DGEQRF( A[i,j] , ... );
+      break;
+    case tDLARFT:
+      DLARFT( A[i,j] , A[i,i] , ... );
+      break;
+    case tDTSQRF:
+      DTSQRF( A[i,j] , A[j,j]... );
+      break;
+    case tDSSRFT:
+      DSSRFT( A[i,j] , A[i,k] , A[k,j] , ... );
+      break;
+    default:
+      error( "Unknown task type." );
+    }
+  }
+    \end{lstlisting}
+\end{minipage}\end{center}
+\noindent where {\tt A} is the matrix over which the QR
+decmposition is executed.
+
 \subsection{Task-Based Barnes-Hut N-Body Solver}
 
+The Barnes-Hut tree-code is an algorithm to compute the
+interactions between a set of particles
+
 \section{Conclusions}
 
-Main points:
+Main points/differences to other schedulers:
 \begin{itemize}
-    \item Conflicts are better than dependencies.
+    \item Conflicts.
     \item Specifying the entire task tree before execution
@@
     \item Combination of locality, weight, and availability.
+    \item Compact implementation.
+\end{itemize}
+
+Possible improvements
+\begin{itemize}
+    \item Use number of locks as well when selecting from a queue.
+    \item Use total/max queue weight when selecting for work-stealing.
+    \item Use different resource sizes when computing queue score.
+    \item Dynamically add new tasks, but potentially wrecks havoc
+        with the weights.
 \end{itemize}
@@
 % Bibliography
-\bibliographystyle{hacm}
+\bibliographystyle{ACM-Reference-Format-Journals}
 \bibliography{quicksched}

### 2013-10-29 src/qsched.c
more specific type names, routines for dynamic task creation, as yet untested.

@@ -47,6 +47,190 @@
 
 
 /**
+ * @brief Make sure that the scheduler has enough memory
+ *        allocated for the tasks, resources, dependencies,
+ *        locks, uses, and data.
+ *
+ * @param s Pointer to the #qsched.
+ * @param nr_tasks Minimum number of tasks.
+ * @param nr_res Minimum number of reources.
+ * @param nr_deps Minimum number of dependencies.
+ * @param nr_locks Minimum number of locks.
+ * @param nr_uses Minimum number of uses.
+ * @param size_data Mnimum size of task data.
+ *
+ * This procedure pre-allocates resources in the sched. Although
+ * These buffers are grown automatically, this is not possible
+ * when dynamic tasks are spawned (see #qsched_addtask_dynamic),
+ * and insufficient buffers my cause dynamic task allocation to fail!
+ *
+ * Keep in mind that the individual blocks of task data are
+ * rounded up to 16 bytes for alignment.
+ * The number of dependencies, locks, and uses refers to the
+ * number of times #qsched_addunlock, #qsched_addlock, and
+ * #qsched_adduses are called, respectively.
+ */
+ 
+void qsched_ensure ( struct qsched *s , int nr_tasks , int nr_res , int nr_deps , int nr_locks , int nr_uses , int size_data ) {
+
+    int dirty = 0;
+    
+    /* Re-allocate tasks? */
+    if ( s->size < nr_tasks ) {
+        dirty = 1;
+        struct task *tasks_new;
+        if ( ( tasks_new = (struct task *)malloc( sizeof(struct task) * nr_tasks ) ) == NULL )
+            error( "Failed to allocate new task buffer." );
+        memcpy( tasks_new , s->tasks , sizeof(struct task) * s->count );
+        free( s->tasks );
+        s->tasks = tasks_new;
+        s->size = nr_tasks;
+        }
+        
+    /* Re-allocate resources? */
+    if ( s->size_res < nr_res ) {
+        dirty = 1;
+        struct res *res_new;
+        if ( ( res_new = (struct res *)malloc( sizeof(struct res) * nr_res ) ) == NULL )
+            error( "Failed to allocate new res buffer." );
+        memcpy( res_new , s->res , sizeof(struct res) * s->count_res );
+        free( s->res );
+        s->res = res_new;
[…]

### 2013-10-29 src/qsched.h

@@ -59,13 +59,13 @@ struct qsched {
     struct task *tasks;
     
     /* The dependency indices. */
-    int *deps, *deps_key;
+    qsched_task_t *deps;
     
     /* The conflict/lock array. */
-    int *locks, *locks_key;
+    qsched_res_t *locks, *uses;
     
-    /* The conflict/use array. */
-    int *uses, *uses_key;
+    /* Sorting indices. */
+    qsched_task_t *deps_key, *locks_key, *uses_key;
     
     /* The shared resources. */
     // lock_type *res;
@@ -141,6 +141,7 @@ void qsched_unlockres ( struct qsched *s , int rid );
 int qsched_locktask ( struct qsched *s , int tid );
 void qsched_unlocktask ( struct qsched *s , int tid );
 void qsched_prepare ( struct qsched *s );
+void qsched_enqueue ( struct qsched *s , struct task *t );
 
 /* External functions. */
 void qsched_init ( struct qsched *s , int nr_queues , int flags );
@@ -152,3 +153,5 @@ void qsched_adduse ( struct qsched *s , qsched_task_t t , qsched_res_t res );
 void qsched_free ( struct qsched *s );
 void qsched_run ( struct qsched *s , int nr_threads , qsched_funtype fun );
 void qsched_reset ( struct qsched *s );
+void qsched_addtask_dynamic ( struct qsched *s , int type , unsigned int flags , void *data , int data_size , int cost , qsched_res_t *locks , int nr_locks , qsched_res_t *uses , int nr_uses );
+void qsched_ensure ( struct qsched *s , int nr_tasks , int nr_res , int nr_deps , int nr_locks , int nr_uses , int size_data );

### 2013-10-29 examples/test_qr.c
use LAPACKE wherever i can get away with it.

@@ -29,12 +29,13 @@
 #include <math.h>
 #include <omp.h>
 
+/* LAPACKE header. */
+#include <lapacke.h>
+#include <cblas.h>
+
+
 /* Local includes. */
 #include "quicksched.h"
-
-
-/* Prototypes for BLAS function. */
-void dtrmm_ ( char * , char * , char * , char * , int * , int * , double * , double * , int * , double * , int * );
 
 
 /*
@@ -105,47 +106,6 @@ void calcvkDouble	(double topDiag,
 }
 
 
-/**
- * \brief Computes a Householder reflector \f$v\f$ of a vector \f$x\f$ for a single block
- *
- * Computes: \f[v = \textrm{sign}(x_1)||x||_2e_1+x\f]
- * Then does a non-standard normalisation \f$v\f$: \f[v = \frac{v}{v_1}\f]
- * 
- * \param x Pointer to an array containing a column vector to compute the Householder reflector of
- * \param l The number of elements in \f$x\f$
- * \param vk A pointer to an allocated array to store the resulting vector of size l - 1
- * due to the implied 1 as the first element
- *
- * \returns void
- */
-void calcvkSingle	(double* x,
-			int l,
-			double* vk)
-{
-	int sign, i;
-	double norm, div, beta;
-
-	sign = x[0] >= 0.0 ? 1 : -1;
-	beta = x[0];
-	//copy the values
-	for(i = 1; i < l; i++)
-		vk[i-1] = x[i];
-
-	//take the euclidian norm of the original vector
-	norm = do2norm(x, l);
-	//calculate the new normalisation
-	beta += norm * sign;
-
-	if(norm != 0.0)
-	{
-		//normalise 
-		div = 1/beta;
-	
-		for(i = 0; i < l-1; i++)
-			vk[i] *= div;
-	}
-}
-
 void updateDoubleQ_WY	(double* blockA,
 			double* blockB,
 			double* blockTau,
@@ -186,60 +146,6 @@ void updateDoubleQ_WY	(double* blockA,
 		blockB[(k*ldm) + i] = hhVector[i];
 
 	blockTau[k] = tau;
-}
-
-void updatekthSingleWY	(double* blockV,
-			double* tauBlock,
-			double beta,
-			int k,
-			int m, int n, int ldm,
-			double* w)
-{
-	/* Insert beta on the diagonal of Tau */
[…]

### 2013-10-30 src/qsched.c
fix bug in res re-allocation.

@@ -941,7 +941,7 @@ void qsched_prepare ( struct qsched *s ) {
  
 int qsched_addres ( struct qsched *s , int parent ) {
 
-    void *temp1, *temp2;
+    struct res *res_new;
     int id;
 
     /* Lock the sched. */
@@ -954,18 +954,17 @@ int qsched_addres ( struct qsched *s , int parent ) {
         s->size_res *= qsched_stretch;
         
         /* Allocate a new task list. */
-        if ( ( temp1 = malloc( sizeof(lock_type) * s->size_res ) ) == NULL ||
-             ( temp2 = malloc( sizeof(int) * s->size_res ) ) == NULL )
+        if ( ( res_new = malloc( sizeof(struct res) * s->size_res ) ) == NULL )
             error( "Failed to allocate new res lists." );
             
         /* Copy the res and owners over to the new list. */
-        memcpy( temp1 , (void *)s->res , sizeof(struct res) * s->count_res );
+        memcpy( res_new , s->res , sizeof(struct res) * s->count_res );
         
         /* Free the old res lists. */
-        free( (int *)s->res );
+        free( s->res );
         
         /* Set the new res lists. */
-        s->res = (struct res *)temp1;
+        s->res = res_new;
     
         }
         

### 2013-10-30 examples/test_bh.c
add timers to the BH example.

@@ -610,12 +610,13 @@ void create_tasks ( struct qsched *s , struct cell *ci , struct cell *cj ) {
  * @param nr_threads Number of threads to use.
  */
  
-void test_bh ( int N , int nr_threads ) {
+void test_bh ( int N , int nr_threads , int runs ) {
 
     int k;
     struct cell *root;
     struct part *parts;
     struct qsched s;
+    ticks tic, tot_setup = 0, tot_run = 0;
     
     /* Runner function. */
     void runner ( int type , void *data ) {
@@ -666,17 +667,30 @@ void test_bh ( int N , int nr_threads ) {
     root->count = N;
     root->parts = parts;
     cell_split( root , &s );
-    
+
     /* Create the tasks. */
+    tic = getticks();
     create_tasks( &s , root , NULL );
+    tot_setup += getticks() - tic;
+
+    /* Loop over the number of runs. */
+    for ( k = 0 ; k < runs ; k++ ) {
     
-    /* Execute the tasks. */
-    qsched_run( &s , nr_threads , runner );
+        /* Execute the tasks. */
+        tic = getticks();
+        qsched_run( &s , nr_threads , runner );
+        tot_run += getticks() - tic;
+        
+        }
         
     /* Dump the tasks. */
     for ( k = 0 ; k < s.count ; k++ )
         printf( " %i %i %lli %lli\n" , s.tasks[k].type , s.tasks[k].qid , s.tasks[k].tic , s.tasks[k].toc );
         
+    /* Dump the costs. */
+    message( "costs: setup=%lli ticks, run=%lli ticks." ,
+        tot_setup , tot_run/runs );
+    
     /* Clean up. */
     qsched_free( &s );
     
@@ -690,7 +704,7 @@ void test_bh ( int N , int nr_threads ) {
 int main ( int argc , char *argv[] ) {
 
     int c, nr_threads;
-    int N = 1000;
+    int N = 1000, runs = 1;
     
     /* Get the number of threads. */
     #pragma omp parallel shared(nr_threads)
@@ -700,19 +714,23 @@ int main ( int argc , char *argv[] ) {
     }
     
     /* Parse the options */
-    while ( ( c = getopt( argc , argv  , "m:n:k:t:" ) ) != -1 )
+    while ( ( c = getopt( argc , argv  , "n:r:t:" ) ) != -1 )
         switch( c ) {
 	        case 'n':
 	            if ( sscanf( optarg , "%d" , &N ) != 1 )
[…]

### 2013-10-30 examples/runs.sh
add script for benchmarking.

@@ -1,0 +1,16 @@
+#!/bin/bash
+
+# Executable and parameters
+prog = test_bh
+params = -n 1000000 -r 10
+
+# Main loop
+for n in {1..16}
+do
+
+    if [ ! -e ${prog}_${n}.dump ]
+    then
+        ./${prog} ${params} -t ${n} > ${prog}_${n}.dump
+    fi
+
+done

### 2013-10-30 examples/test_bh.c
better parameters.

@@ -36,7 +36,7 @@
 /* Some local constants. */
 #define cell_pool_grow 100
 #define cell_maxparts 100
-#define task_limit 1000
+#define task_limit 4000
 #define const_G 6.6738e-11
 #define dist_min 0.5
 
@@ -684,8 +684,8 @@ void test_bh ( int N , int nr_threads , int runs ) {
         }
         
     /* Dump the tasks. */
-    for ( k = 0 ; k < s.count ; k++ )
-        printf( " %i %i %lli %lli\n" , s.tasks[k].type , s.tasks[k].qid , s.tasks[k].tic , s.tasks[k].toc );
+    /* for ( k = 0 ; k < s.count ; k++ )
+        printf( " %i %i %lli %lli\n" , s.tasks[k].type , s.tasks[k].qid , s.tasks[k].tic , s.tasks[k].toc ); */
         
     /* Dump the costs. */
     message( "costs: setup=%lli ticks, run=%lli ticks." ,

### 2013-10-30 examples/runs.sh
minor fix.

@@ -1,8 +1,8 @@
 #!/bin/bash
 
 # Executable and parameters
-prog = test_bh
-params = -n 1000000 -r 10
+prog=test_bh
+params="-n 1000000 -r 10"
 
 # Main loop
 for n in {1..16}

### 2013-10-30 examples/test_bh.c
better reporting.

@@ -616,7 +616,7 @@ void test_bh ( int N , int nr_threads , int runs ) {
     struct cell *root;
     struct part *parts;
     struct qsched s;
-    ticks tic, tot_setup = 0, tot_run = 0;
+    ticks tic, toc_run, tot_setup = 0, tot_run = 0;
     
     /* Runner function. */
     void runner ( int type , void *data ) {
@@ -679,7 +679,9 @@ void test_bh ( int N , int nr_threads , int runs ) {
         /* Execute the tasks. */
         tic = getticks();
         qsched_run( &s , nr_threads , runner );
-        tot_run += getticks() - tic;
+	toc_run = getticks();
+	message( "%ith run took %lli ticks..." , k , toc_run - tic );
+        tot_run += toc_run - tic;
         
         }
         

### 2013-10-30 examples/runs_bh.sh
make runs files case-specific.

@@ -5,7 +5,7 @@ prog=test_bh
 params="-n 1000000 -r 10"
 
 # Main loop
-for n in {1..16}
+for n in {1..64}
 do
 
     if [ ! -e ${prog}_${n}.dump ]

### 2013-10-30 examples/test_qr.c
add better timing facilities.

@@ -233,7 +233,7 @@ void DSSRFT	(double* blockV,
  * @param nr_threads Number of threads to use.
  */
  
-void test_qr ( int m , int n , int nr_threads ) {
+void test_qr ( int m , int n , int nr_threads , int runs ) {
 
     int k, j, i;
     double *A, *A_orig, *tau;
@@ -241,6 +241,7 @@ void test_qr ( int m , int n , int nr_threads ) {
     qsched_task_t *tid, tid_new;
     qsched_res_t *rid;
     int data[3];
+    ticks tic, toc_run, tot_setup, tot_run = 0;
     
     enum task_types { task_DGEQRF , task_DLARFT , task_DTSQRF , task_DSSRFT };
 
@@ -302,6 +303,7 @@ void test_qr ( int m , int n , int nr_threads ) {
     qsched_init( &s , nr_threads , qsched_flag_none );
     
     /* Allocate and init the task ID and resource ID matrix. */
+    tic = getticks();
     if ( ( tid = (qsched_task_t *)malloc( sizeof(qsched_task_t) * m * n ) ) == NULL ||
          ( rid = (qsched_res_t *)malloc( sizeof(qsched_res_t) * m * n ) ) == NULL )
         error( "Failed to allocate tid/rid matrix." );
@@ -363,9 +365,20 @@ void test_qr ( int m , int n , int nr_threads ) {
             }
     
         } /* build the tasks. */
+    tot_setup = getticks() - tic;
         
-    /* Execute the the tasks. */
-    qsched_run( &s , nr_threads , runner );
+        
+    /* Loop over the number of runs. */
+    for ( k = 0 ; k < runs ; k++ ) {
+    
+        /* Execute the the tasks. */
+        tic = getticks();
+        qsched_run( &s , nr_threads , runner );
+        toc_run = getticks(); 
+	    message( "%ith run took %lli ticks..." , k , toc_run - tic );
+        tot_run += toc_run - tic;
+        
+        }
     
         
     /* Dump A. */
@@ -387,11 +400,15 @@ void test_qr ( int m , int n , int nr_threads ) {
     printf( "];\n" ); */
     
     /* Dump the tasks. */
-    for ( k = 0 ; k < s.count ; k++ ) {
+    /* for ( k = 0 ; k < s.count ; k++ ) {
         int *d = (int *)&s.data[ s.tasks[k].data ];
[…]

### 2013-10-30 examples/runs_qr.sh
test_qr now has a script too.

@@ -1,0 +1,16 @@
+#!/bin/bash
+
+# Executable and parameters
+prog=test_qr
+params="-m 32 -n 32 -r 10"
+
+# Main loop
+for n in {1..64}
+do
+
+    if [ ! -e ${prog}_${n}.dump ]
+    then
+        ./${prog} ${params} -t ${n} > ${prog}_${n}.dump
+    fi
+
+done

### 2013-10-30 examples/test_qr.c
add option to change tile size.

@@ -233,7 +233,7 @@ void DSSRFT	(double* blockV,
  * @param nr_threads Number of threads to use.
  */
  
-void test_qr ( int m , int n , int nr_threads , int runs ) {
+void test_qr ( int m , int n , int K , int nr_threads , int runs ) {
 
     int k, j, i;
     double *A, *A_orig, *tau;
@@ -252,26 +252,26 @@ void test_qr ( int m , int n , int nr_threads , int runs ) {
         /* Decode the task data. */
         int *idata = (int *)data;
         int i = idata[0], j = idata[1], k = idata[2];
-        double buff[ 2*32*32 ];
+        double buff[ 2*K*K ];
         
         /* Decode and execute the task. */
         switch ( type ) {
             case task_DGEQRF:
-                LAPACKE_dgeqrf_work( LAPACK_COL_MAJOR , 32, 32 ,
-                                &A[ j*m*32*32 + i*32 ] , m*32 , &tau[ j*m*32 + i*32 ] ,
-                                buff , 2*32*32 );
+                LAPACKE_dgeqrf_work( LAPACK_COL_MAJOR , K, K ,
+                                &A[ j*m*K*K + i*K ] , m*K , &tau[ j*m*K + i*K ] ,
+                                buff , 2*K*K );
                 break;
             case task_DLARFT:
                 LAPACKE_dlarft_work( LAPACK_COL_MAJOR , 'F' , 'C' ,
-                                32 , 32 , &A[ i*m*32*32 + i*32 ] ,
-                                m*32 , &tau[ i*m*32 + i*32 ] , &A[ j*m*32*32 + i*32 ] ,
-                                m*32 );
+                                K , K , &A[ i*m*K*K + i*K ] ,
+                                m*K , &tau[ i*m*K + i*K ] , &A[ j*m*K*K + i*K ] ,
+                                m*K );
                 break;
             case task_DTSQRF:
-                DTSQRF( &A[ j*m*32*32 + j*32 ] , &A[ j*m*32*32 + i*32 ] , &tau[ j*m*32 + i*32 ] , 32 , 32 , 32 , 32*m , buff );
+                DTSQRF( &A[ j*m*K*K + j*K ] , &A[ j*m*K*K + i*K ] , &tau[ j*m*K + i*K ] , K , K , K , K*m , buff );
                 break;
             case task_DSSRFT:
[…]

### 2013-10-30 src/qsched.c
add qsched_flag_noreown and function qsched_res_own.

@@ -47,6 +47,22 @@
 
 
 /**
+ * @brief Change the owner of a resource.
+ *
+ * @param Pointer to the #qsched.
+ * @param res Resource handle.
+ * @param owner The ID of the new owner.
+ */
+ 
+void qsched_res_own ( struct qsched *s , qsched_res_t res , int owner ) {
+
+    /* Force the new ownership. */
+    s->res[ res ].owner = owner;
+
+    }
+
+
+/**
  * @brief Make sure that the scheduler has enough memory
  *        allocated for the tasks, resources, dependencies,
  *        locks, uses, and data.
@@ -418,7 +434,7 @@ void *qsched_getdata( struct qsched *s , struct task *t ) {
  
 void qsched_enqueue ( struct qsched *s , struct task *t ) {
 
-    int j, qid, scores[ s->nr_queues ];
+    int j, qid, scores[ s->nr_queues ], oid;
     
     /* If this is a virtual task, just do its unlocks and leave. */
     if ( t->flags & task_flag_virtual ) {
@@ -438,9 +454,11 @@ void qsched_enqueue ( struct qsched *s , struct task *t ) {
 
         /* Loop over the locks and uses, and get their owners. */
         for ( j = 0 ; j < t->nr_locks ; j++ )
-            scores[ s->res[ t->locks[j] ].owner ] += 1;
+            if ( ( oid = s->res[ t->locks[j] ].owner ) != qsched_owner_none )
+                scores[ oid ] += 1;
         for ( j = 0 ; j < t->nr_uses ; j++ )
-            scores[ s->res[ t->uses[j] ].owner ] += 1;
+            if ( ( oid = s->res[ t->uses[j] ].owner ) != qsched_owner_none )
+                scores[ oid ] += 1;
 
         /* Find the queue with the highest score. */
         qid = 0;
@@ -695,10 +713,12 @@ struct task *qsched_gettask ( struct qsched *s , int qid ) {
             t = &s->tasks[tid];
         
             /* Own the resources. */
-            for ( k = 0 ; k < t->nr_locks ; k++ )
-                s->res[ t->locks[k] ].owner = qid;
-            for ( k = 0 ; k < t->nr_uses ; k++ )
-                s->res[ t->uses[k] ].owner = qid;
+            if ( !( s->flags & qsched_flag_noreown ) ) {
+                for ( k = 0 ; k < t->nr_locks ; k++ )
+                    s->res[ t->locks[k] ].owner = qid;
[…]

### 2013-10-30 src/qsched.h

@@ -23,6 +23,7 @@
 #define qsched_flag_ready                2
 #define qsched_flag_yield                4
 #define qsched_flag_pthread              8
+#define qsched_flag_noreown              16
 
 /* Some sched-specific constants. */
 #define qsched_stretch                   2
@@ -35,6 +36,7 @@
 #define qsched_data_round                16
 #define qsched_res_none                  (-1)
 #define qsched_task_none                 (-1)
+#define qsched_owner_none                (-1)
 
 
 /** Type used for task identifiers. */
@@ -145,7 +147,7 @@ void qsched_enqueue ( struct qsched *s , struct task *t );
 
 /* External functions. */
 void qsched_init ( struct qsched *s , int nr_queues , int flags );
-qsched_res_t qsched_addres ( struct qsched *s , qsched_res_t parent );
+qsched_res_t qsched_addres ( struct qsched *s , int owner , qsched_res_t parent );
 void qsched_addlock ( struct qsched *s , qsched_task_t t , qsched_res_t res );
 void qsched_addunlock ( struct qsched *s , qsched_task_t ta , qsched_task_t tb );
 qsched_task_t qsched_addtask ( struct qsched *s , int type , unsigned int flags , void *data , int data_size , int cost );
@@ -155,3 +157,4 @@ void qsched_run ( struct qsched *s , int nr_threads , qsched_funtype fun );
 void qsched_reset ( struct qsched *s );
 void qsched_addtask_dynamic ( struct qsched *s , int type , unsigned int flags , void *data , int data_size , int cost , qsched_res_t *locks , int nr_locks , qsched_res_t *uses , int nr_uses );
 void qsched_ensure ( struct qsched *s , int nr_tasks , int nr_res , int nr_deps , int nr_locks , int nr_uses , int size_data );
+void qsched_res_own ( struct qsched *s , qsched_res_t res , int owner );

### 2013-10-30 examples/test.c
add qsched_flag_noreown and function qsched_res_own.

@@ -118,7 +118,7 @@ void test2 ( int m , int n , int k , int nr_threads ) {
     /* Build a task for each tile of the matrix c. */
     for ( i = 0 ; i < m ; i++ )
         for ( j = 0 ; j < n ; j++ ) {
-            rid = qsched_addres( &s , -1 );
+            rid = qsched_addres( &s , qsched_owner_none , qsched_res_none );
             data[0] = i; data[1] = j;
             for ( kk = 0 ; kk < k ; kk++ ) {
                 data[2] = kk;
@@ -222,7 +222,7 @@ void test1 ( int m , int n , int k , int nr_threads ) {
     for ( i = 0 ; i < m ; i++ )
         for ( j = 0 ; j < n ; j++ ) {
             data[0] = i; data[1] = j;
-            rid = qsched_addres( &s , -1 );
+            rid = qsched_addres( &s , qsched_owner_none , qsched_res_none );
             tid = qsched_addtask( &s , 1 , task_flag_none , data , 2*sizeof(int) , 1 );
             qsched_addlock( &s , tid , rid );
             }

### 2013-10-30 examples/test_bh.c

@@ -121,6 +121,7 @@ struct cell *cell_get ( ) {
  *        to the sched.
  *
  * @param c The #cell to be split.
+ * @param N The total number of parts.
  * @param s The #sched to store the resources.
  */
  
@@ -131,10 +132,15 @@ void cell_split ( struct cell *c , struct qsched *s ) {
     struct cell *cp;
     int left[8], right[8];
     double pivot[3];
+    static struct cell *root = NULL;
+    
+    /* Set the root cell. */
+    if ( root == NULL )
+        root = c;
     
     /* Add a resource for this cell if it doesn't have one yet. */
     if ( c->res == qsched_res_none ) 
-        c->res = qsched_addres( s , qsched_res_none );
+        c->res = qsched_addres( s , qsched_owner_none , qsched_res_none );
     
     /* Attach a center-of-mass task to the cell. */
     c->com_tid = qsched_addtask( s , task_type_com , task_flag_none , &c , sizeof(struct cell *) , 1 );
@@ -155,7 +161,7 @@ void cell_split ( struct cell *c , struct qsched *s ) {
             cp->h[0] = c->h[0]/2;
             cp->h[1] = c->h[1]/2;
             cp->h[2] = c->h[2]/2;
-            cp->res = qsched_addres( s , c->res );
+            cp->res = qsched_addres( s , qsched_owner_none , c->res );
             if ( k & 4 )
                 cp->loc[0] += cp->h[0];
             if ( k & 2 )
@@ -229,6 +235,9 @@ void cell_split ( struct cell *c , struct qsched *s ) {
             qsched_addunlock( s , c->progeny[k]->com_tid , c->com_tid );
             
         } /* does the cell need to be split? */
+        
+    /* Set this cell's resources ownership. */
+    qsched_res_own( s , c->res , s->nr_queues * ( c->parts - root->parts ) / root->count );
         
     }
     
@@ -644,7 +653,7 @@ void test_bh ( int N , int nr_threads , int runs ) {
         }
     
     /* Initialize the scheduler. */
-    qsched_init( &s , nr_threads , qsched_flag_yield );
+    qsched_init( &s , nr_threads , qsched_flag_noreown );
     
     /* Init and fill the particle array. */
     if ( ( parts = (struct part *)malloc( sizeof(struct part) * N ) ) == NULL )
@@ -679,8 +688,8 @@ void test_bh ( int N , int nr_threads , int runs ) {
         /* Execute the tasks. */
[…]

### 2013-10-30 examples/test_qr.c

@@ -309,7 +309,7 @@ void test_qr ( int m , int n , int K , int nr_threads , int runs ) {
         error( "Failed to allocate tid/rid matrix." );
     for ( k = 0 ; k < m * n ; k++ ) {
         tid[k] = qsched_task_none;
-        rid[k] = qsched_addres( &s , qsched_res_none );
+        rid[k] = qsched_addres( &s , qsched_owner_none , qsched_res_none );
         }
     
     /* Build the tasks. */

### 2013-10-31 examples/test_bh.c
fixed bug in task type creation.

@@ -552,10 +552,26 @@ void create_tasks ( struct qsched *s , struct cell *ci , struct cell *cj ) {
             r2 += dx*dx;
             }
             
+        /* Are the cells sufficiently well separated? */
+        if ( r2/ci->h[0] > dist_min*dist_min ) {
+
+            /* Interact ci's parts with cj as a cell. */
+            data[0] = ci; data[1] = cj;
+            tid = qsched_addtask( s , task_type_pair_pc , task_flag_none , data , sizeof(struct cell *) * 2 , ci->count );
+            qsched_addlock( s , tid , ci->res );
+            qsched_addunlock( s , cj->com_tid , tid );
+
+            /* Interact cj's parts with ci as a cell. */
+            data[0] = cj; data[1] = ci;
+            tid = qsched_addtask( s , task_type_pair_pc , task_flag_none , data , sizeof(struct cell *) * 2 , ci->count );
+            qsched_addlock( s , tid , cj->res );
+            qsched_addunlock( s , ci->com_tid , tid );
+
+            }
+            
         /* Does this task need to be broken-down further? */
-        if ( r2/ci->h[0] < dist_min*dist_min &&
-             ci->split && cj->split &&
-             ci->count > task_limit && cj->count > task_limit ) {
+        else if ( ci->split && cj->split &&
+                  ci->count > task_limit && cj->count > task_limit ) {
              
             /* Loop over all pairs between ci and cj's progeny. */
             for ( j = 0 ; j < 8 ; j++ )
@@ -564,49 +580,27 @@ void create_tasks ( struct qsched *s , struct cell *ci , struct cell *cj ) {
         
             }
             
-        /* Otherwise, generate a task. */
+        /* Otherwise, generate a part-part task. */
         else {
         
-            /* Can we get away with cell-part interactions? */
-            if ( r2/ci->h[0] < dist_min*dist_min ) {
-            
-                /* Interact ci's parts with cj as a cell. */
-                data[0] = ci; data[1] = cj;
-                tid = qsched_addtask( s , task_type_pair_pc , task_flag_none , data , sizeof(struct cell *) * 2 , ci->count );
-                qsched_addlock( s , tid , ci->res );
-                qsched_addunlock( s , cj->com_tid , tid );
-        
-                /* Interact cj's parts with ci as a cell. */
-                data[0] = cj; data[1] = ci;
-                tid = qsched_addtask( s , task_type_pair_pc , task_flag_none , data , sizeof(struct cell *) * 2 , ci->count );
[…]

### 2013-10-31 examples/test_bh.c
fixed same bug, yet in a different place.

@@ -353,68 +353,69 @@ void iact_pair ( struct cell *ci , struct cell *cj ) {
             for ( k = j+1 ; k < 8 ; k++ )
                 iact_pair( ci->progeny[j] , cj->progeny[k] );
             
-    /* Otherwise, compute interactions. */
+    /* Get the minimum distance between both cells. */
+    for ( r2 = 0.0, k = 0 ; k < 3 ; k++ ) {
+        dx[k] = fabs( ci->loc[k] - cj->loc[k] );
+        if ( dx[k] > 0 )
+            dx[k] -= ci->h[k];
+        r2 += dx[k]*dx[k];
+        }
+
+    /* Sufficiently well-separated? */
+    if ( r2/ci->h[0] > dist_min*dist_min ) {
+
+        /* Compute the center of mass interactions. */
+        iact_pair_pc( ci , cj );
+        iact_pair_pc( cj , ci );
+
+        }
+
+    /* Recurse? */
+    else if ( ci->split && cj->split )
+        for ( j = 0 ; j < 8 ; j++ )
+            for ( k = j+1 ; k < 8 ; k++ )
+                iact_pair( ci->progeny[j] , cj->progeny[k] );
+            
+    /* Otherwise, do direct interactions. */
     else {
-    
-        /* Get the minimum distance between both cells. */
-        for ( r2 = 0.0, k = 0 ; k < 3 ; k++ ) {
-            dx[k] = fabs( ci->loc[k] - cj->loc[k] );
-            if ( dx[k] > 0 )
-                dx[k] -= ci->h[k];
-            r2 += dx[k]*dx[k];
-            }
-            
-        /* Sufficiently well-separated? */
-        if ( r2/ci->h[0] < dist_min*dist_min ) {
-        
-            /* Compute the center of mass interactions. */
-            iact_pair_pc( ci , cj );
-            iact_pair_pc( cj , ci );
-        
-            }
-            
-        /* Otherwise, do direct interactions. */
-        else {
-            
-            /* Loop over all particles... */
-            for ( i = 0 ; i < count_i ; i++ ) {
 
-                /* Init the ith particle's data. */
-                for ( k = 0 ; k < 3 ; k++ ) {
-                    xi[k] = parts_i[i].x[k];
-                    ai[k] = 0.0;
+        /* Loop over all particles... */
+        for ( i = 0 ; i < count_i ; i++ ) {
+
+            /* Init the ith particle's data. */
[…]

### 2013-11-01 src/qsched.c
added timers.

@@ -46,6 +46,11 @@
 #include "queue.h"
 
 
+/** Timer names. */
+char *qsched_timer_names[ qsched_timer_count ] = 
+    { "queue" , "lock" , "gettask" , "done" , "prepare" };
+    
+    
 /**
  * @brief Change the owner of a resource.
  *
@@ -263,6 +268,11 @@ void qsched_reset ( struct qsched *s ) {
     s->count_uses = 0;
     s->count_res = 0;
 
+    /* Clear the timers. */
+    #ifdef TIMERS
+        bzero( s->timers , sizeof(ticks) * qsched_timer_count );
+    #endif
+    
     }
 
 
@@ -486,6 +496,8 @@ void qsched_done ( struct qsched *s , struct task *t ) {
 
     int k;
     struct task *t2;
+        
+    TIMER_TIC
     
     /* Release this task's locks. */
     for ( k = 0 ; k < t->nr_locks ; k++ )
@@ -518,6 +530,10 @@ void qsched_done ( struct qsched *s , struct task *t ) {
             pthread_mutex_unlock( &s->mutex );
             }
     #endif
+    
+    /* Careful, this may pick up duplicate timers if virtual
+       tasks are used. */
+    TIMER_TOC( s , qsched_timer_done )
                 
     }
     
@@ -610,6 +626,8 @@ int qsched_locktask ( struct qsched *s , int tid ) {
 
     int k;
     struct task *t;
+    
+    TIMER_TIC
 
     /* Get a pointer on the task. */
     t = &s->tasks[tid];
@@ -627,13 +645,16 @@ int qsched_locktask ( struct qsched *s , int tid ) {
             qsched_unlockres( s , t->locks[k] );
 
         /* Fail. */
+        TIMER_TOC( s , qsched_timer_lock )
         return 0;
 
         }
         
     /* Otherwise, all went well. */
-    else
+    else {
+        TIMER_TOC( s , qsched_timer_lock )
         return 1;
+        }
             
     }
     
@@ -649,6 +670,8 @@ void qsched_unlocktask ( struct qsched *s , int tid ) {
 
     int k;
     struct task *t;
+    
+    TIMER_TIC
 
     /* Get a pointer on the task. */
     t = &s->tasks[tid];
@@ -656,6 +679,8 @@ void qsched_unlocktask ( struct qsched *s , int tid ) {
     /* Unlock the used resources. */
[…]

### 2013-11-01 src/qsched.h

@@ -51,6 +51,30 @@ typedef int qsched_res_t;
 typedef void (*qsched_funtype)( int , void * );
 
 
+/** Timer types. */
+enum qsched_timer {
+    qsched_timer_queue = 0,
+    qsched_timer_lock,
+    qsched_timer_gettask,
+    qsched_timer_done,
+    qsched_timer_prepare,
+    qsched_timer_count
+    };
+extern char *qsched_timer_names[];
+    
+    
+/* Timer macros. */
+#ifdef TIMERS
+    #define TIMER_TIC ticks __tic = getticks();
+    #define TIMER_TIC2 __tic = getticks();
+    #define TIMER_TOC(s,tid) atomic_add( &s->timers[tid] , getticks() - __tic );
+#else
+    #define TIMER_TIC
+    #define TIMER_TIC2
+    #define TIMER_TOC
+#endif
+
+
 /* The sched data structre. */
 struct qsched {
 
@@ -126,6 +150,11 @@ struct qsched {
     #ifdef HAVE_PTHREAD
         pthread_cond_t cond;
         pthread_mutex_t mutex;
+    #endif
+    
+    /* Timers. */
+    #ifdef TIMERS
+        ticks timers[ qsched_timer_count ];
     #endif
     
     };

### 2013-11-01 examples/mkplots_tasks.m
added timers.

@@ -58,7 +58,7 @@ maxd = max( [ tasks(:,3) ; tasks(:,4) ] );
 % Init the plot
 clf;
 subplot('position',[ 0.05 , 0.1 , 0.9 , 0.8 ]);
-colours = [ 1 0 0 ; 1 1 0 ; 0 1 0 ; 0 0 1 ];
+colours = [ 255 34 0 ; 130 255 0 ; 0 184 255 ; 255 237 0 ] / 255;
 hold on;
 
 % Plot the tasks

### 2013-11-01 examples/test_bh.c

@@ -696,6 +696,10 @@ void test_bh ( int N , int nr_threads , int runs ) {
     /* Dump the costs. */
     message( "costs: setup=%lli ticks, run=%lli ticks." ,
         tot_setup , tot_run/runs );
+        
+    /* Dump the timers. */
+    for ( k = 0 ; k < qsched_timer_count ; k++ )
+        message( "timer %s is %lli ticks." , qsched_timer_names[k] , s.timers[k]/runs );
     
     /* Clean up. */
     qsched_free( &s );

### 2013-11-01 examples/test_qr.c

@@ -366,6 +366,13 @@ void test_qr ( int m , int n , int K , int nr_threads , int runs ) {
     
         } /* build the tasks. */
     tot_setup = getticks() - tic;
+    
+    /* Dump the number of tasks. */
+    message( "total nr of tasks: %i." , s.count );
+    message( "total nr of deps: %i." , s.count_deps );
+    message( "total nr of res: %i." , s.count_res );
+    message( "total nr of locks: %i." , s.count_locks );
+    message( "total nr of uses: %i." , s.count_uses );    
         
         
     /* Loop over the number of runs. */
@@ -408,6 +415,10 @@ void test_qr ( int m , int n , int K , int nr_threads , int runs ) {
     /* Dump the costs. */
     message( "costs: setup=%lli ticks, run=%lli ticks." ,
         tot_setup , tot_run/runs );
+    
+    /* Dump the timers. */
+    for ( k = 0 ; k < qsched_timer_count ; k++ )
+        message( "timer %s is %lli ticks." , qsched_timer_names[k] , s.timers[k]/runs );
     
     /* Clean up. */
     qsched_free( &s );

### 2013-11-01 examples/test_qr_ompss.c
add ompss version of the qr test.

@@ -1,0 +1,394 @@
+/*******************************************************************************
+ * This file is part of QuickSched.
+ * Coypright (c) 2013 Pedro Gonnet (pedro.gonnet@durham.ac.uk)
+ * 
+ * This program is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as published
+ * by the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ * 
+ * You should have received a copy of the GNU Lesser General Public License
+ * along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ * 
+ ******************************************************************************/
+
+
+/* Config parameters. */
+#include "../config.h"
+
+/* Standard includes. */
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <unistd.h>
+#include <math.h>
+#include <omp.h>
+
+/* LAPACKE header. */
+#include <lapacke.h>
+#include <cblas.h>
+
+
+/* Local includes. */
+#include "quicksched.h"
+
+
+/*
+ * Sam's routines for the tiled QR decomposition.
+ */
+
+/*
+  \brief Computes 2-norm of a vector \f$x\f$
+  
+  Computes the 2-norm by computing the following: \f[\textrm{2-norm}=\sqrt_0^lx(i)^2\f]
+ */
+double do2norm(double* x, int l)
+{
+	double sum = 0, norm;
+	int i;
+
+	for(i = 0; i < l; i++)
+		sum += x[i] * x[i];
+
+	norm = sqrt(sum);
+
+	return norm;
+}
+
+/**
+ * \brief Computes a Householder reflector from a pair of vectors from coupled blocks
+ *
+ * Calculates the Householder vector of the vector formed by a column in a pair of coupled blocks.
+ * There is a single non-zero element, in the first row, of the top vector. This is passed as topDiag
+ *
[…]

### 2013-11-01 examples/test_qr_ompss.c
add timers for the tasks, taskID always zero for now.

@@ -38,6 +38,15 @@
 #include "quicksched.h"
 
 
+/* Stuff to collect task data. */
+struct timer {
+    int threadID, type;
+    ticks tic, toc;
+    };
+struct timer *timers;
+int nr_timers = 0;
+
+
 /*
  * Sam's routines for the tiled QR decomposition.
  */
@@ -161,6 +170,11 @@ void DTSQRF	(double* blockA,
 	double* xVectA, *xVectB;
     double hhVector[ 2*ma ];
 	
+    int ind = atomic_inc( &nr_timers );
+    timers[ind].threadID = omp_get_thread_num();
+    timers[ind].type = 2;
+    timers[ind].tic = getticks();
+    
 	xVectA = blockA;
 	xVectB = blockB;
 
@@ -181,6 +195,7 @@ void DTSQRF	(double* blockA,
 		xVectA += ldm + 1;
 		xVectB += ldm;
 	}
+    timers[ind].toc = getticks();
 }
 
 #pragma omp task in( blockV[0] ) in( blockB[0] ) inout( blockA[0] ) in( blockTau[0] )
@@ -193,6 +208,11 @@ void DSSRFT	(double* blockV,
 
 	double tau, beta;
 
+    int ind = atomic_inc( &nr_timers );
+    timers[ind].threadID = omp_get_thread_num();
+    timers[ind].type = 3;
+    timers[ind].tic = getticks();
+    
 	/* Compute b_j = b_j - tau*v*v'*b_j for each column j of blocks A & B,
 	   and for each householder vector v of blockV */
 
@@ -224,6 +244,7 @@ void DSSRFT	(double* blockV,
 				blockB[(j*ldm) + i] -= beta * blockV[(k*ldm) + i];
 		}
 	}
+    timers[ind].toc = getticks();
 }
 
 
@@ -234,7 +255,12 @@ void DSSRFT	(double* blockV,
 #pragma omp task inout( a[0] ) inout( tau[0] )
 void DGEQRF ( int matrix_order, lapack_int m, lapack_int n,
                                 double* a, lapack_int lda, double* tau ) {
+    int ind = atomic_inc( &nr_timers );
+    timers[ind].threadID = omp_get_thread_num();
+    timers[ind].type = 0;
+    timers[ind].tic = getticks();
     LAPACKE_dgeqrf( matrix_order, m, n, a, lda, tau );
+    timers[ind].toc = getticks();
     }
 
 
[…]

### 2013-11-01 examples/test_qr_ompss.c
change way tasks are defined.

@@ -157,7 +157,7 @@ void updateDoubleQ_WY	(double* blockA,
 	blockTau[k] = tau;
 }
 
-#pragma omp task in( blockA[0] ) inout( blockB[0] ) in( blockTau[0] )
+// #pragma omp task in( blockA[0] ) inout( blockB[0] ) in( blockTau[0] )
 void DTSQRF	(double* blockA,
 		double* blockB,
 		double* blockTau,
@@ -198,7 +198,7 @@ void DTSQRF	(double* blockA,
     timers[ind].toc = getticks();
 }
 
-#pragma omp task in( blockV[0] ) in( blockB[0] ) inout( blockA[0] ) in( blockTau[0] )
+// #pragma omp task in( blockV[0] ) in( blockB[0] ) inout( blockA[0] ) in( blockTau[0] )
 void DSSRFT	(double* blockV,
 		double* blockA, double* blockB,
 		double* blockTau,
@@ -252,7 +252,7 @@ void DSSRFT	(double* blockV,
  * @breif Wrapper to get the dependencies right.
  */
  
-#pragma omp task inout( a[0] ) inout( tau[0] )
+// #pragma omp task inout( a[0] ) inout( tau[0] )
 void DGEQRF ( int matrix_order, lapack_int m, lapack_int n,
                                 double* a, lapack_int lda, double* tau ) {
     int ind = atomic_inc( &nr_timers );
@@ -268,7 +268,7 @@ void DGEQRF ( int matrix_order, lapack_int m, lapack_int n,
  * @breif Wrapper to get the dependencies right.
  */
  
-#pragma omp task inout( v[0] ) in( tau[0] ) inout( t[0] )
+// #pragma omp task inout( v[0] ) in( tau[0] ) inout( t[0] )
 void DLARFT ( int matrix_order, char direct, char storev,
                                 lapack_int n, lapack_int k, const double* v,
                                 lapack_int ldv, const double* tau, double* t,
@@ -295,6 +295,7 @@ void test_qr ( int m , int n , int K , int nr_threads , int runs ) {
     int k, j, i, r;
     double *A, *A_orig, *tau;
     ticks tic, toc_run, tot_setup, tot_run = 0;
+    int tid[ m*n ];
 
 
     /* Allocate and fill the original matrix. */
@@ -327,10 +328,12 @@ void test_qr ( int m , int n , int K , int nr_threads , int runs ) {
[…]

### 2013-11-01 paper/paper.tex
new figures and backup.

@@
 into hierarchical cells which are used at different levels.
+The owner field is the ID of the queue to which this
+resource has been preferentially assigned.
 
@@
     }
-  if ( res != NULL ) {
+  if ( res != NULL && s->reown ) {
     for ( k = 0 ; k < res->nr_locks ; k++ )
@@
 (line~7).
-If a task could be obtained from any queue (line~13),
+If a task could be obtained from any queue and task re-owning
+is switched on (line~13),
 the resources it locks and uses are marked as now being owned
@@
         \vspace{1mm}
-    \item {\tt qsched\_res\_t qsched\_addres( struct qsched *s , qsched\_res\_t parent )} \\
+    \item {\tt qsched\_res\_t qsched\_addres( struct qsched *s , int owner , qsched\_res\_t parent )} \\
         Creates a new resource within the given {\tt qsched} and returns
         its handle.
+        The owner field is the initial queue ID to which this resource
+        should be assigned, or {\tt qsched\_owner\_none}.
         The {\tt parent} field is the handle of the heirarchical parent of
@@
 
+The first test is the tiled QR decomposition originally
+from \citeN{ref:Buttari2009}, which has been used as a benchmark
+by other authors \cite{ref:Agullo2009b,ref:Badia2009,ref:Bosilca2012}.
+This example only requires dependencies and is presented
+only as a point of comparison to existing task-based
+parallel infrastructures.
+
+The second example is a Barnes-Hut tree-code, a problem
+similar to the Fast Multipole Method described in both
+\citeN{ref:Ltaief2012} and \citeN{ref:Agullo2013}.
+This example shows how conflicts, modeled
+via hierarchical resources, can be useful.
+
 The source code of both examples is distributed with the
-QuickSched library.
+QuickSched library, along with scripts to run the benchmarks
+and generate the plots used in the following.
 
@@
 
+The QR decomposition was computed for a $2048\times 2048$
+random matrix using tiles of size $64\times 64$ using QuickSched
+as described above.
+For the DGEQRF and DLARFT kernels, the respective routines
+from the LAPACKE\footnote{\url{http://www.netlib.org/lapack/lapacke.html}}
+library were used.
+The DTSQRF and DSSRFT kernels were implemented by Sam Townsend
+as part of his MSc thesis in Computer Science at Durham University.
+For this matrix, a total of 11440 tasks with 32240 dependencies
+were generated.
+
+All tests were run on a 64-core AMD Opteron 6376 machine running
+at 2.6\,GHz.
+For these tests, OpenMP parallelism and resource re-owning
+were used.
+The QR decomposition was computed 10 times for each number of
+cores, and the average thereof taken for the scaling and
+efficiency results in \fig{QRResults}.
+The timings are for {\tt qsched\_run}, including the cost of
+{\tt qsched\_start}, which does not run in parallel.
+Setting up the scheduler, tasks, and resources took, in all
+cases, an average of 7.2\,ms.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/QR_scaling.pdf,width=0.9\textwidth}}
+    \caption{Strong scaling and parallel efficiency of the tiled QR decomposition
+        computed over a $2048\times 2048$ matrix with tiles of size
+        $64\times 64$.
+        The QR decomposition takes 233\,ms, achieving 73\% parallel
+        efficiency, over all 64 cores.
+        }
+    \label{fig:QRResults}
+\end{figure}
+
 \subsection{Task-Based Barnes-Hut N-Body Solver}
 
-The Barnes-Hut tree-code is an algorithm to compute the
-interactions between a set of particles
+The Barnes-Hut tree-code is an algorithm to approximate the
+solution of an $N$-body problem, i.e.~computing all the
+pairwise interactions between a set of $N$ particles,
+in \oh{N\log N} operations, as opposed to the \oh{N^2}
+naive direct computation.
+
+The algorithm is based on a recursive octree decomposition:
+Starting from a cubic cell containing all the particles,
+the cell is recursivel bisected along all three spatial dimensions,
+resulting in eight sub-cells, until the number of particles
+per cell is smaller than some limit $n_\mathsf{max}$.
+
+The particle interactions can also be formulated recursively:
+Given a particle an a cell of particles, if the particle and cell
+are sufficiently well separated, the particle-cell interactions
+are approximated by interacting the particle with the cell's
+centre of mass.
+If the particle and the cell are too close, and the cell
+has sub-cells, i.e.~it contained more than $n_\mathsf{max}$
+particles and was split in the recursive octree decomposition,
+then the particle is interacted with each of the sub-cells
+recursively.
+Finally, if the cell is not split, i.e.~it is a leaf cell
+in the octree, then the particle is interacted with all
+particles in the cell, except for the particle itself if
+it is in the same cell.
+This operation is performed for each particle, starting
+with the root-level cell containing all the particles.
+
+The cells have the following structure:
+\begin{center}\begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}
+struct cell {
+  double loc[3], h[3], com[3], mass;
+  int split, count;
+  struct part *parts;
+  struct cell *progeny[8];
+  qsched_res_t res;
+  qsched_task_t task_com;
+  };
+    \end{lstlisting}
+\end{minipage}\end{center}
+\noindent where {\tt loc} and {\tt h} are the location
+and size of the cell, respectively.
+The {\tt com} and {\tt mass} fields represent the cell's
+center of mass, which will be used in the interactions.
+The {\tt res} filed is the hieararchical resource representing
+the cell's particles, and it is the parent resource of the cell
+progeny's {\tt res}.
+Similarly, the {\tt task\_com} is a task handle to
+compute the center of mass of the cell's particles, and 
+it depends on the {\tt task\_com} of all the progeny if
+the cell is split.
+{\tt parts} is a pointer to an array of {\tt count} 
+particle strutures, which contain all the particle
+data of the form:
+\begin{center}\begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}
+struct part {
+  double x[3], a[3], mass;
+  int id;
+  };
+    \end{lstlisting}
+\end{minipage}\end{center}
+\noindent i.e.~the particle position, accelleration, mass,
+and ID, respectively.
+
+The particle data is sorted hierarchically, following the
+octree structure.
+Unlike in many codes, where the leaves store an array of
+pointers to the underlying particles, the cells, at all
+levels, store only a pointer to the first of their own particles,
+and the total number of particles.
+The current approach, illustrated in \fig{CellParts} is not
+only more compact, it also allows a direct and more cache-efficient access
+to the list of particles for any inner node of the tree.
+The cost of sorting the particles, with a recursive
+partitioning similar to QuickSort \cite{ref:Hoare1962},
+is in \oh{N\log N}.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/CellParts.pdf,width=0.9\textwidth}}
+    \caption{Hierarchical ordering of the particle data structures
+    (right) according to their cell (left).
+    Each cell has a pointer to the first of its particles (same colour
+    as cells) in the same global parts array.}
+    \label{fig:CellParts}
+\end{figure}
+
+The task-based implementation will consist of four
+types of tasks:
+\begin{itemize}
+    \item {\em Self}-interactions in which all particles
+        in a single cell interact with all other particles in the
+        same cell,
+    \item {\em Particle-particle pair}-interactions in which
+        all particles in a pair of cells interact with all
+        particles in the opposite cell, and
+    \item {\em Particle-cell pair}-interactions in which
+        all particles in one cell are interacted with the
+        center of mass of another cell.
+    \item {\em Center of mass} tasks, which compute
+        the center of mass of a single cell either from
+        the sum of the centers of mass of its sub-cells
+        if it has been split, or directly from the
+        particles otherwise.
+\end{itemize}
+
+\begin{figure}
+\begin{center}\begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}
+enum { tSELF , tPAIR_PP , tPAIR_PC , tCOM };
+void make_tasks ( struct qsched *s , struct cell *ci , struct cell *cj ) {
+  int j, k;
+  qsched_task_t tid;
+  struct cell *data[2];
+  if ( cj == NULL ) {
+    if ( ci->split && ci->count > n_task )
+      for ( j = 0 ; j < 8 ; j++ ) {
+        make_tasks( s , ci->progeny[j] , NULL );
+        for ( k = j+1 ; k < 8 ; k++ )
+          make_tasks( s , ci->progeny[j] , ci->progeny[k] );
+        }
+    else {
+      tid = qsched_addtask( s , tSELF , qsched_flags_none ,
+        &ci , sizeof(struct cell *) , ci->count*ci->count );
+      qsched_addlock( s , tid , ci->res );
+      if ( ci->split )
+        qsched_addunlock( s , ci->com , tid );
+      }
+    }
+  else if ( ci and cj are well separated ) {
+      data[0] = ci; data[1] = cj;
+      tid = qsched_addtask( s , tPAIR_PC , qsched_flags_none ,
+        data , sizeof(struct cell *)*2 , ci->count );
+      qsched_addlock( s , tid , ci->res );
+      qsched_addunlock( s , cj->com , tid );
+      data[0] = cj; data[1] = ci;
+      tid = qsched_addtask( s , tPAIR_PC , qsched_flags_none ,
+        data , sizeof(struct cell *)*2 , cj->count );
+      qsched_addlock( s , tid , cj->res );
+      qsched_addunlock( s , ci->com , tid );
+      }
+  else if ( ci->split && cj->split && 
+     ci->count*cj->count > n_task*n_task )
+    for ( j = 0 ; j < 8 ; j++ )
+      for ( k = 0 ; k < 8 ; k++ )
+        make_tasks( s , ci->progeny[j] , cj->progeny[k] );
+  else {
+    data[0] = ci; data[1] = cj;
+    tid = qsched_addtask( s , tPAIR_PP , qsched_flags_none ,
+      data , sizeof(struct cell *)*2 , ci->count*cj->count );
+    qsched_addlock( s , tid , ci->res );
+    qsched_addlock( s , tid , cj->res );
+    if ( ci->split && cj->split ) {
+      qsched_addunlock( s , ci->com , tid );
+      qsched_addunlock( s , cj->com , tid );
+      }
+    }
+  }
+    \end{lstlisting}
+\end{minipage}\end{center}
+    \caption{C-like pseudo-code for recursive task creation
+        for the Barnes-Hut tree-code.}
+    \label{fig:MakeTasks}
+\end{figure}
+
+These tasks can be created recursively over the cell hierarchy
+as shown in the function {\tt make\_tasks} in \fig{MakeTasks}.
+The function is called on the root cell with the root cell
+and {\tt NULL} as its two cell parameters.
+The function recurses as follows:
+\begin{itemize}
+    \item If called with a single (line~6), split (line~7) cell,
+        recurse over all the cell's sub cells (line~9), and all
+        pairs of the cell's sub cells (line~11),
+    \item If called with a single unslplit cell (line~13),
+        create a self-interaction task on that cell (line~14),
+    \item If called with two cells that are sufficiently well
+        separated (line~21), create two particle-cell pair
+        interactions (lines~23 and~28) over both cells in
+        opposite orders, which depend on the center of mass
+        task of each cell,
+    \item If called with two cells that are not well
+        separated and both cells are split (line~33),
+        recurse over all pairs of sub-cells spanning
+        both cells (line~37), and
+    \item If called with two cells that are not well separated
+        and either of the cells are not split, create
+        a particle-particle pair task over both cells.
+\end{itemize}
+\noindent where every interaction task additionally locks
+the cells on which it operates (lines~16, 25, 30, and 42--43).
+
+In order to reduce the number of tasks, and to prevent generating
+too many very small tasks, the task generation only recurses
+if the cells contain more than a minimum number $n_\mathsf{task}$
+of threads each (lines~7 and~34).
+The tasks themselves are then left to recurse over the sub-trees,
+which is why in these cases, the tasks are made to depend on the
+center of mass tasks (lines~17--18 and~41--47)
+which may be used in the ensuing interactions.
+
+\begin{figure}
+\begin{center}\begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}
+void comp_com ( struct cell *c ) {
+  int j, k;
+  c->com[0] = 0.0; c->com[1] = 0.0; c->com[2] = 0.0;
+  c->mas = 0.0;
+  if ( c->split )
+    for ( k = 0 ; k < 8 ; k++ ) {
+      struct cell *cp = c->progeny[k]
+      for ( j = 0 ; j < 3 ; j++ ) c->com[j] += cp->com[j]*cp->mass;
+      c->mass += cp->mass;
+      }
+  else
+    for ( k = 0 ; k < 8 ; k++ ) {
+      struct part *p = &c->parts[k];
+      for ( j = 0 ; j < 3 ; j++ ) c->com[j] += p->x[j]*p->mass;
+      c->mass += p->mass;
+      }
+  c->com[0] /= c->mass; c->com[1] /= c->mass; c->com[2] /= c->mass; 
+  }
+
+void comp_self ( struct cell *c ) {
+  int j, k;
+  if ( c->split )
+    for ( j = 0 ; j < 8 ; j++ ) {
+      comp_self( c->progeny[j] );
+      for ( k = j+1 ; k < 8 ; k++ )
+        comp_pair( c->progeny[j] , c->progeny[k] );
+      }
+  else
+    for ( j = 0 ; j < c->count ; j++ )
+      for ( k = j+1 ; k < c->count ; k++ )
+        interact c->parts[j] and c->parts[k].
+  }
+
+void comp_pair ( struct cell *ci , struct cell *cj ) {
+  int j, k;
+  if ( ci and cj well separated ) {
+    comp_pair_pc( ci , cj );
+    comp_pair_pc( cj , ci );
+    }
+  else if ( ci->split && cj->split )
+    for ( j = 0 ; j < 8 ; j++ )
+      for ( k = 0 ; k < 8 ; k++ )
+        comp_pair( ci->progeny[j] , cj->progeny[k] );
+  else
+    for ( j = 0 ; j < ci->count ; j++ )
+      for ( k = 0 ; k < cj->count ; k++ )
+        interact ci->parts[j] and cj->parts[k].
+  }
+
+void comp_pair_cp ( struct cell *ci , struct cell *cj ) {
+  int k;
+  for ( k = 0 ; k < ci->count ; k++ )
+    interact ci->parts[k] and cj center of mass.
+  }
+    \end{lstlisting}
+\end{minipage}\end{center}
+    \caption{Task functions for the Barnes-Hut tree-code.}
+    \label{fig:BHTasks}
+\end{figure}
+
+The functions for the task themselves are relatively
+straight-forward and shown in \fig{BHTasks}, and the
+execution function can be written as:
+\begin{center}\begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}
+void exec_fun ( int type , void *data ) {
+  struct cell **cells = (struct cell **)data;
+  switch ( type ) {
+    case tSELF:
+      comp_self( cells[0] );
+      break;
+    case tPAIR_PP:
+      comp_pair( cells[0] , cells[1] );
+      break;
+    case tPAIR_PC:
+      comp_pair_pc( cells[0] , cells[1] );
+      break;
+    case tCOM:
+      comp_com( cells[0] );
+      break;
+    default:
+      error( "Unknown task type." );
+    }
+  }
+    \end{lstlisting}
+\end{minipage}\end{center}
+
+This Barnes-Hut tree-code was used to approximate the gravitational
+N-Body problem for 1\,000\,000 particles with random coordinates the
+parameters $n_\mathsf{max}=100$ and $n_\mathsf{task}=5000$.
+Cell pairs were considered well separated if not adjecent.
+Using the above scheme generated 161\,613 tasks, of which
+512 self-interaction tasks, 18\,532 particle-particle interaction
+task, 105\,120 particle-cell interaction tasks, and 37\,449
+center of mass tasks.
+Additionally 179\,632 dependencies were generated, along with
+142'696 locks on 37\,449 resources.
+
+As with the previous example, all
+tests were run on a 64-core AMD Opteron 6376 machine running
+at 2.6\,GHz.
+For these tests, OpenMP parallelism was used and resource
+re-owning was switched off.
+The interactions  computed 10 times for each number of
+cores, and the average thereof taken for the scaling and
+efficiency results in \fig{BHResults}.
+The timings are for {\tt qsched\_run}, including the cost of
+{\tt qsched\_start}, which does not run in parallel.
+Setting up the scheduler, tasks, and resources took, in all
+cases, an average of 51.3\,ms.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/BH_scaling.pdf,width=0.9\textwidth}}
+    \caption{Strong scaling and parallel efficiency of the Barnes-Hut tree-code
+        computed over 1\,000\,000 particles.
+        Solving the N-Body problem takes 3.5\,s, achieving 52\% parallel
+        efficiency, over all 64 cores.
+        }
+    \label{fig:BHResults}
+\end{figure}
+
+Unlike the QR decomposition, the results scale well only to
+32 cores, achieving 90\% parallel efficiency, and then
+level off for increasing numbers of cores.
+This, however, is not a problem of the task-based parallel
+algorith, or of QuickSched, but of the cache hierarchies
+of the underlying hardware.
+On the AMD Opteron 6376, the cores are grouped into pairs
+which each share a 2\,MB L2 cache.
+Each group of four pairs, or eight cores, shares a common
+6\,MB L3 cache.
+It is the difference between the sum of the L2 caches and
+the (smaller) L3 cache which causes problems.
 

### 2013-11-10 paper/paper.tex
final draft now ready.

@@
 \begin{abstract}
-Bla.
+This paper describes QuickSched, a compact and efficient Open-Source
+C-language library for task-based shared-memory parallel programming.
+QuickSched extends the standard dependency-only scheme of task-based
+programming with the concept of task conflicts, i.e.~sets of tasks
+that can be executed in any order, yet not concurrently.
+These conflicts are modelled using exclusively lockable
+hierarchical resources.
+The scheduler is shown to perform and scale well on a 64-core parallel
+shared-memory machine for two example problems: A tiled QR
+decomposition and a task-based Barnes-Hut tree code.
 \end{abstract}
@@
 \begin{bottomstuff}
-This work was supported by a Durham Universiy Seedcorn Grant.
+This work was supported by a Durham University Seedcorn Grant.
 
@@
 Task-based parallelism is a conceptually simple paradigm for
-shared-memory paralelism in which a computation is broken-down
+shared-memory parallelism in which a computation is broken-down
 into a set of inter-dependent tasks which are executed
@@
 
-This computational model is trival to parallelize.
+This computational model is trivial to parallelize.
 Given a set of inter-dependent tasks and a set of computational
@@
 new tasks, or until all tasks in the DAG have been executed.
-
-\fig{Tasks} shows the DAG for a set of tasks.
+\fig{Tasks} shows such a DAG for a set of tasks.
 The arrows indicate the direction of the dependency, i.e.~an
@@
         on task $A$.
-        Tasks $A$, $G$, and $J$ have no unsatisifed dependencies and
+        Tasks $A$, $G$, and $J$ have no unsatisfied dependencies and
         can therefore be executed.
-        Once task $G$ has completed, tasks $F$ and $H$ become available.}
+        Once task $G$ has completed, tasks $F$ and $H$ become available,
+        and task $E$ only becomes available once both tasks $D$ and $F$
+        have completed.}
     \label{fig:Tasks}
@@
 would create implicit dependencies between the lowest-level
-tasks $C$, $E$, and $K$.
+tasks $C$, $E$, and $K$.\footnote{
+The main thread spawns tasks $A$, $G$ and $J$, $A$ spawns $B$ and $D$,
+$G$ spawns $F$, $H$, and then $I$, $B$ spawns $C$.
+The main thread then has to {\tt sync} for $A$, $G$, and $J$,
+and thus implicitly all their spawned tasks, before executing
+$E$ and $K$.}
 
@@
 QUARK \cite{ref:Yarkhan2011}, and KAAPI \cite{ref:Gautier2007}
-the programmer spcifies
+the programmer specifies
 what shared data each task will access, and how that data will
@@
 dependencies explicitly.
-Intel's Threding Building Blocks (TBB)
+Intel's Threading Building Blocks (TBB)
 \cite{ref:Reinders2010}
@@
 OmpSs \cite{ref:Duran2011} extends this scheme with automatic
-dependency generation as in SMP superscalar, along with
+dependency generation as in SMP superscalar, of which it
+is a direct descendant, along with
 the ability to explicitly wait on certain tasks.
@@
 In order to avoid concurrent access to that resource, it is
-imperative that the execution of both tasks does not overlap,
-yet the order in which the tasks are exectued does not matter.
-In the following, such a relationship will be refered to
+imperative that the execution of both tasks do not overlap,
+yet the order in which the tasks are executed is irrelevant.
+In the following, such a relationship will be referred to
 as a ``conflict'' between two tasks.
@@
 joined by thick dashed lines.
-None of tasks $F$, $H$, and $I$ cannot be executed concurrently,
+None of tasks $F$, $H$, and $I$ can be executed concurrently,
 i.e. they must be serialized, yet in no particular order.
@@
 ordering on conflicting tasks.
-This, however, imposes unnecessary restriction on the order
-in which tasks can be scheduled, especially in the presence
+This unnecessary restriction on the order
+in which tasks can be scheduled can severely limit the
+parallelizability of a computation, especially in the presence
 of multiple conflicts per task.
@@
 this problem in their respective implementations of the Fast Multipole
-Method, in which forces computed in different tasks are
+Method (FMM), in which forces computed in different tasks are
 accumulated on a set of particles.
@@
 
-This paper presents QuickSched, a framework for task-based
-parallel programming with constraints.
 In order to be of practical use, and to scale well with modern
-multi-core shared-memory architectures, the task scheduler
-must be:
+multi-core shared-memory architectures, any task scheduler
+implementing constraints must be:
 \begin{itemize}
@@
     \item {\em Parallel-efficient}: Tasks should be executed in an order
-        that sufficient work is available for all computatoinal
+        that sufficient work is available for all computational
         threads at all times.
@@
     decomposition described in \cite{ref:Buttari2009} and for
-    which the QUARK scheduler was originally developed,
+    which the QUARK scheduler was originally developed, and
     \item A task-based Barnes-Hut tree-code to compute the
-    gravitational N-body problem,
+    gravitational N-body problem similar to the FMM codes
+    of \citeN{ref:Ltaief2012} and \citeN{ref:Agullo2013},
 \end{enumerate}
@@
 and can be used to assess its efficiency.
-Section~5 concludes with some general observations and future work.
+Section~5 concludes with some general observations and future work
+directions.
 
@@
 
-The QuickSched task scheduler consits of four main
+The QuickSched task scheduler consist of four main
 objects types: {\em task}, {\em resource}, {\em scheduler},
@@
 
-The division of labour regarding {\em correctness}
+The division of labor regarding {\em correctness}
 between the scheduler and
@@
 in \figs{Tasks}{TaskConflicts}.
-
 Conversely, {\tt wait} is the number of unresolved dependencies
@@
 
-The {\tt locks} field points to the first element of
+The {\tt locks} field of each task points to the first element of
 an array of {\tt nr\_locks} pointers to {\em resources}
@@
 \noindent where the array {\tt top} contains the task indices
-in reverse topological order. 
-The test in line~10 is a convenient check if the tasks and teir
+in topological order. 
+The test in line~10 is a convenient check if the tasks and their
 dependencies actually do form an acyclic graph.
@@
 topological order, computing the recursive weight as the sum of the
-task cost and the maximum weight of the tasks it unlocks,
-and recomputing the task waits at the same time.
+task cost and the maximum weight of the tasks it unlocks (line~6),
+and recomputing the task waits at the same time (line~3).
 
@@
 
-The data structure for the resources is as follows:
+Resources consist of the following data structure:
 \begin{center}\begin{minipage}{0.9\textwidth}
@@
 into hierarchical cells which are used at different levels.
-The owner field is the ID of the queue to which this
+The {\tt owner} field is the ID of the queue to which this
 resource has been preferentially assigned.
@@
 The {\tt hold} field is a counter indicating how many
-sub-resources of the current resouce are locked.
+sub-resources of the current resource are locked.
 If a resource's hold counter is not zero, then it is
@@
     \centerline{\epsfig{file=figures/Resources.pdf,width=0.6\textwidth}}
-    \caption{A hierarchicy of cells (left) and the hierarchy of
+    \caption{A hierarchy of cells (left) and the hierarchy of
         corresponding hierarchical resources at each level.
@@
         arrows indicate the resource's parent.
-        Resources coloured red are locked, resources coloured orange
+        Resources colored red are locked, resources colored orange
         are held, where the number in the square indicates the
@@
 \noindent where the resource itself is unlocked (line~3)
-and the hold counter of its parents is decremented (lines~4--5).
+and the hold counters of all of its hierarchical parents
+are decremented (lines~4--5).
 
@@
 One possible strategy would be to maintain an array of tasks
-sorted by their weights, and to trverse that list in descending
+sorted by their weights, and to traverse that list in descending
 order, trying to lock the resources of each task, until
@@
 or removing an entry is in \oh{n} for $n$ elements.
-
-Using an unsorted array requires only \oh{1} operations for
-insertion and deletion, but is undesireable as it completely
+Using an unsorted array would require only \oh{1} operations for
+insertion and deletion, but is undesirable as it completely
 ignores the task weights.
@@
 As a compromise, the queue stores the tasks in an array
-organized as a max-heap, with the task with maximum weight
+organized as a max-heap, i.e.~where the $k$th entry is ``larger''
+than both the $2k+1$st and the $2k+2$nd entry,
+with the task with maximum weight
 in the first position.
-Maintainig this heap structure thus requires \oh{\log n}
+Maintaining this heap structure thus requires \oh{\log n}
 operations for both insertion and deletion, i.e. for the
@@
 Although the first task in the array will be the task with
-maximum weight, the following tasks are only losely ordered,
+maximum weight, the following tasks are only loosely ordered,
 where the $k$th of $n$ tasks has a larger weight than at least
@@
 and the heap order is fixed (line~4).
-Before exiting, the lock on the queue is released (line~5).
+Before returning, the lock on the queue is released (line~5).
 
@@
     for ( j = 0 ; j < q->tasks[k]->nr_locks ; j++ )
-      if ( !resource_lock( q->tasks[k]->lock[j] )
+      if ( !resource_lock( q->tasks[k]->lock[j] ) )
         break;
@@
 (line~13).
-If all the locks on a task could be obtained (line~14), the
+If all the locks on a task could be obtained (line~15), the
 task pointer is replaced by the last pointer in the heap (line~17)
@@
 QuickSched task scheduler, and as such contains the instances
-other three object types:
+of the other three object types:
 \begin{center}\begin{minipage}{0.9\textwidth}
@@
 \end{minipage}\end{center}
-\noindent where\dots the only additional field {\tt waiting} is
+\noindent where the only additional field {\tt waiting} is
 used to keep track of the number of tasks that have not been
@@
 schemes of other standard libraries, the type name {\tt qsched}
-is used.
+is used for the scheduler data type.
 
@@
 for most compilers, is used to create a parallel section
-in which the code between lines~4 and~12 is executed
+in which the code between lines~4 and~11 is executed
 concurrently.
 The parallel section consists of a loop (lines~7--10) in
-which a task is acquired via the {\em execution function}
-{\tt qsched\_gettask}
-and its type and data are passed to a user-supplied execution
-function.
+which a task is acquired via {\tt qsched\_gettask}
+and its type and data are passed to a user-supplied
+{\em execution function} {\tt fun}.
 Once the task has been executed, it is returned to the
-scheduler via the function {\tt qsched\_done}.
+scheduler via the function {\tt qsched\_done}, i.e.~to
+unlock its resources and unlock dependent tasks.
 The loop terminates when the scheduler runs out of tasks,
-i.e.~when {\tt qstack\_gettask} returns {\tt NULL}, and
+i.e.~when {\tt qsched\_gettask} returns {\tt NULL}, and
 the function exits once all the threads have exited their
@@
 The operations in line~3 are described in \sect{tasks}.
-
 The function {\tt qsched\_enqueue} tries to identify the best
@@
 task resources ``owned'', or last used, by each queue.
-In lines~9--12 the queue with the highest such score is 
-chosen on which the task is then put (line~13).
+In lines~9--11 the queue with the highest such score is 
+chosen on which the task is then put (line~12).
 
@@
 \noindent where the parameter {\tt qid} is the index of the
-prefered queue.
+preferred queue.
 If the queue is empty, or all of the tasks in that queue had
@@
 the resources it locks and uses are marked as now being owned
-by the prefered queue (lines~14--17).
+by the preferred queue (lines~14--17).
 Finally, the task, or {\tt NULL} if no task could be obtained,
@@
 
-The final step in a task's lifecycle is, on completion,
-to unlock the tasks which depend on it.
+The final step in a task's life cycle is, on completion,
+to unlock the resources and tasks which depend on it.
 This is handled by the function {\tt qsched\_done}:
@@
   int k;
+  for ( k = 0 ; k < t->nr_locks ; k++ )
+    resource_unlock( t->locks[k] );
   for ( k = 0 ; k < t->nr_unlocks ; k++ )
@@
 \noindent If any of the unlocked tasks' wait counters
-goes to zero (line~4), then the unlocked task is ready to
+goes to zero (line~6), then the unlocked task is ready to
 run and is immediately dispatched via {\tt qsched\_enqueue},
@@
 Once all the dependent tasks have been unlocked, the
-{\tt waiting} counter is decremented (line~6).
+{\tt waiting} counter is decremented (line~8).
 
@@
 This section describes the interface functions and how they
-are called, whereas the following section contains
+are called, and the following section contains
 examples of how QuickSched can be used.
@@
 interface to the task scheduler.
-As such, it provieds functionality for task and resource
+As such, it provides functionality for task and resource
 creation, for assigning resources to tasks, either as locks
@@
         should be assigned, or {\tt qsched\_owner\_none}.
-        The {\tt parent} field is the handle of the heirarchical parent of
+        The {\tt parent} field is the handle of the hierarchical parent of
         the new resource or {\tt qsched\_res\_none} if the resource
@@
         is not part of a conflict.
-        This information is used when assinging tasks to specific queues.
+        This information is used when assigning tasks to specific queues.
         \vspace{1mm}
@@
 
-The library can be compiled to use either OpenMP or the
+The library can be compiled to use OpenMP and/or the
 {\tt pthreads} library \cite{ref:Pthreads1995}.
@@
 either the {\tt qsched\_flag\_yield}
-or the {\tt qsched\_flag\_pthread} switches to using {\tt pthreads}
-for the parallel loop.
+or the {\tt qsched\_flag\_pthread} switches to using {\tt pthreads},
+if available, for the parallel loop.
 
@@
 This may be a problem if other parts of the user application
-are running in the background simultaneously.
+are running concurrently in the background.
 Calling {\tt qsched\_init} with the {\tt qsched\_flag\_yield}
@@
 
-This section presents two test cases showing both
+This section presents two test cases showing
 how QuickSched can be used in real-world applications, and
@@
 The first test is the tiled QR decomposition originally
-from \citeN{ref:Buttari2009}, which has been used as a benchmark
+described in \citeN{ref:Buttari2009}, which has been used as a benchmark
 by other authors \cite{ref:Agullo2009b,ref:Badia2009,ref:Bosilca2012}.
 This example only requires dependencies and is presented 
-only as a point of comparison to existing task-based
-parallel infrastructures.
+as a point of comparison to existing task-based parallel
+programming infrastructures.
 
@@
 and generate the plots used in the following.
+All examples were compiled with gcc v.\,4.8.1 using the
+{\tt -O2 -march=native} flags and run on a 
+tests were run on a 64-core AMD Opteron 6376 machine running
+at 2.6\,GHz.
 
@@
 \citeN{ref:Buttari2009} introduced the concept of using task-based
-parallelsim for tile-based algorithms in numerical linear algebra,
+parallelism for tile-based algorithms in numerical linear algebra,
 presenting parallel codes for the Cholesky, LU, and QR
@@
         of $4\times 4$ tiles.
-        Each circle represents a tile, and its colour represents
-        the type of taks on that tile at that level.
+        Each circle represents a tile, and its color represents
+        the type of task on that tile at that level.
         Empty circles have no task associated with them.
@@
     \begin{tabular}{llll}
-        Task & cond. & depends on task & locks tile \\
+        Task & where & depends on task(s) & locks tile(s) \\
         \hline
@@
 task additionally modifies the lower triangular part of the $(j,j)$th tile.
-
-Setting up the dependencies and locks for a matrix of
-$m\times n$ tiles is implemented as shown in \fig{CodeQR},
 
@@
 Similarly, {\tt rid} stores the handles of the resources for each
-tile of the matrix, allocated in line~8.
+tile of the matrix, which are allocated in line~8.
 
@@
 Algorithm~2 of \citeN{ref:Buttari2009}.
-E.g.~for each level {\tt k} (line~10), a DGEQRF task is created
+For each level {\tt k} (line~10), a DGEQRF task is created
 for tile $(k,k)$ (lines~13--14).
@@
 
-The execution function for these tasks simply calls the apropriate
+The execution function for these tasks simply calls the appropriate
 kernels on the matrix tiles given by the task data:
@@
 \noindent where {\tt A} is the matrix over which the QR
-decmposition is executed.
+decomposition is executed.
 
@@
 
-All tests were run on a 64-core AMD Opteron 6376 machine running
-at 2.6\,GHz.
 For these tests, OpenMP parallelism and resource re-owning
-were used.
+were used with one queue per core.
 The QR decomposition was computed 10 times for each number of
@@
 
+The same decomposition was implementing using OmpSs v.\,1.99.0,
+calling the kernels directly using {\tt \#pragma omp task}
+annotations with the respective dependencies.
+The scaling and efficiency relative to QuickSched are 
+shown in \fig{QRResults} as well.
+The difference in timings is the result of the different
+task scheduling policies, as well as a smaller lag between the
+individual tasks, as shown in \fig{QRTasks},
+for a smaller $1024\times 1024$ matrix on 16 cores of the
+same hardware.
+The most visible difference between both schedulers is that
+the DGEQRF tasks (in red) are scheduled as soon as they
+become available in QuickSched, thus preventing bottlenecks
+near the end of the computation.
+
 \begin{figure}
@@
         $64\times 64$.
-        The QR decomposition takes 233\,ms, achieving 73\% parallel
-        efficiency, over all 64 cores.
+        The QR decomposition with QuickSched takes 233\,ms,
+        achieving 73\% parallel efficiency, over all 64 cores.
+        The scaling and efficiency for OmpSs are computed relative to QuickSched.
         }
     \label{fig:QRResults}
+\end{figure}
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/tasks_qr.pdf,width=0.9\textwidth}}
+    \centerline{\epsfig{file=figures/tasks_qr_ompss.pdf,width=0.9\textwidth}}
+    \caption{Task scheduling in QuickSched (above) and OmpSs (below)
+        for a $1024\times 1024$ matrix on 16 cores.
+        The task colors correspond to those in \fig{QR}.}
+    \label{fig:QRTasks}
 \end{figure}
@@
 naive direct computation.
-
 The algorithm is based on a recursive octree decomposition:
 Starting from a cubic cell containing all the particles,
-the cell is recursivel bisected along all three spatial dimensions,
+the cell is recursively bisected along all three spatial dimensions,
 resulting in eight sub-cells, until the number of particles
@@
 The particle interactions can also be formulated recursively:
-Given a particle an a cell of particles, if the particle and cell
+Given a particle and a set of particles in a cell,
+if the particle and cell
 are sufficiently well separated, the particle-cell interactions
 are approximated by interacting the particle with the cell's
-centre of mass.
+center of mass.
 If the particle and the cell are too close, and the cell
@@
 
-The cells have the following structure:
+The cells themselves are implemented using the following 
+data structure:
 \begin{center}\begin{minipage}{0.9\textwidth}
@@
 The {\tt com} and {\tt mass} fields represent the cell's
-center of mass, which will be used in the interactions.
-The {\tt res} filed is the hieararchical resource representing
+center of mass, which will be used in the particle-cell interactions.
+The {\tt res} filed is the hierarchical resource representing
 the cell's particles, and it is the parent resource of the cell
@@
 {\tt parts} is a pointer to an array of {\tt count} 
-particle strutures, which contain all the particle
+particle structures, which contain all the particle
 data of the form:
@@
 \end{minipage}\end{center}
-\noindent i.e.~the particle position, accelleration, mass,
+\noindent i.e.~the particle position, acceleration, mass,
 and ID, respectively.
@@
     (right) according to their cell (left).
-    Each cell has a pointer to the first of its particles (same colour
+    Each cell has a pointer to the first of its particles (same color
     as cells) in the same global parts array.}
@@
         same cell,
-    \item {\em Particle-particle pair}-interactions in which
-        all particles in a pair of cells interact with all
-        particles in the opposite cell, and
-    \item {\em Particle-cell pair}-interactions in which
+    \item {\em Particle-particle} pair interactions in which
+        all particles in a cell interact with all
+        particles in another cell,
+    \item {\em Particle-cell} pair interactions in which
         all particles in one cell are interacted with the
-        center of mass of another cell.
+        center of mass of another cell, and
     \item {\em Center of mass} tasks, which compute
@@
     \item If called with a single (line~6), split (line~7) cell,
-        recurse over all the cell's sub cells (line~9), and all
-        pairs of the cell's sub cells (line~11),
-    \item If called with a single unslplit cell (line~13),
+        recurse over all the cell's sub-cells (line~9), and all
+        pairs of the cell's sub-cells (line~11),
+    \item If called with a single unsplit cell (line~13),
         create a self-interaction task on that cell (line~14),
@@
 
-In order to reduce the number of tasks, and to prevent generating
-too many very small tasks, the task generation only recurses
+In order to prevent generating
+a large number of very small tasks, the task generation only recurses
 if the cells contain more than a minimum number $n_\mathsf{task}$
@@
 This Barnes-Hut tree-code was used to approximate the gravitational
-N-Body problem for 1\,000\,000 particles with random coordinates the
-parameters $n_\mathsf{max}=100$ and $n_\mathsf{task}=5000$.
-Cell pairs were considered well separated if not adjecent.
+N-Body problem for 1\,000\,000 particles with random coordinates
+in $[0,1]^3$.
+The parameters $n_\mathsf{max}=100$ and $n_\mathsf{task}=5000$
+were used to generate the tasks, and cell pairs were considered
+well separated if not directly adjacent.
 Using the above scheme generated 161\,613 tasks, of which
@@
 center of mass tasks.
-Additionally 179\,632 dependencies were generated, along with
-142'696 locks on 37\,449 resources.
-
-As with the previous example, all
-tests were run on a 64-core AMD Opteron 6376 machine running
-at 2.6\,GHz.
+A total of 179\,632 dependencies were generated, along with
+142\,696 locks on 37\,449 resources.
+
 For these tests, OpenMP parallelism was used and resource
 re-owning was switched off.
+Resource ownership was attributed by dividing the global
+{\tt parts} array by the number of queues and assigning each cell's
+{\tt res} to the fraction of the {\tt parts} array to which
+the first of its own {\tt parts} belong.
 The interactions  computed 10 times for each number of
@@
         computed over 1\,000\,000 particles.
-        Solving the N-Body problem takes 3.5\,s, achieving 52\% parallel
+        Solving the N-Body problem takes 3.3\,s, achieving 60\% parallel
         efficiency, over all 64 cores.
@@
 This, however, is not a problem of the task-based parallel
-algorith, or of QuickSched, but of the cache hierarchies
+algorithm, or of QuickSched, but of the memory bandwidth
 of the underlying hardware.
-On the AMD Opteron 6376, the cores are grouped into pairs
-which each share a 2\,MB L2 cache.
-Each group of four pairs, or eight cores, shares a common
-6\,MB L3 cache.
-It is the difference between the sum of the L2 caches and
-the (smaller) L3 cache which causes problems.
-
-\section{Conclusions}
-
-Main points/differences to other schedulers:
+\fig{BHTasks} shows the accumulated cost of each task type and of 
+QuickSched over the number of cores.
+At 64 cores, the scheduler overheads account for only 4.7\% of
+the total computational cost, whereas,
+as of 30 cores, the cost of both pair types grow by up to
+50\%.
+This is most probably due to memory bandwidth restrictions, as
+the cost of the self interaction tasks, which do twice as much
+computation per memory access, only grow by up to 15\%.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/BH_times.pdf,width=0.8\textwidth}}
+    \caption{Accumulated cost of each task type and of the overheads
+        associated with {\tt qsched\_gettask}.
+        As of $\sim 30$ cores, the cost of both pair interaction task
+        types grow by up to 50\%.
+        The cost of the self interactions, which entail twice as much
+        computation per memory access, grow only by at most 15\%.
+        The scheduler overheads make up less than 5\% of the total
+        time.
+        }
+    \label{fig:BHTimes}
+\end{figure}
+
+\section{Discussion and Conclusions}
+
+The task scheduler described in the previous sections, QuickSched,
+differs from existing task-based programming schemes
+in a number of ways.
+The most obvious such difference is the addition of {\em conflicts},
+modeled using exclusive locks on hierarchical resources.
+This extension to the standard dependencies-only model
+of task-based parallelism allows for more complex task relations,
+such as in the Barnes-Hut tree-code described earlier.
+
+Another significant difference is that the tasks, their
+dependencies, and their conflicts must be described
+{\em explicitly} before the parallel computation starts.
+This as opposed to implicit dependencies generated
+by task spawning, e.g.~as in Cilk, or to extracting the
+dependencies from the task parameters, e.g.~in QUARK or OmpSs.
+Explicitly defining dependencies has the advantage that 
+more elaborate dependency structures can be generated.
+Furthermore, knowing the structure of the entire task
+graph from the start of the computation provides valuable
+information when scheduling the tasks, e.g.~using the 
+critical path along the dependencies to compute the
+task weight.
+
+Finally, as opposed to the most other task-based
+programming environments which rely on compiler extensions
+and/or code pre-processors, QuickSched operates as a regular
+C-language library, based on standard parallel functionality
+provided by OpenMP and/or {\tt pthreads}.
+This ensures a maximum amount of portability on existing
+and future architectures.
+The interfaces are also kept as simple
+as possible in order to reduce the burden on the programmer
+when implementing task-based codes.
+
+The QuickSched library itself is remarkably simple, consisting of
+less than 3\,000 lines of code, including comments.
+Both examples, which are distributed with QuickSched,
+require less than 1\,000 lines of code each.
+
+In both examples, QuickSched performs extremely well, even
+on a large number of shared-memory cores.
+This performance is due, on the one hand, to the
+division of labor between the scheduler and the queues,
+and on the other hand due to the simple yet efficient
+algorithms for task selection and resource locking.
+The task weighting based on the length of the critical
+path of the dependencies delivers, in the examples shown,
+good parallel efficiency.
+
+There are several possible improvements to QuickSched which
+have not been addressed in this paper.
+The most obvious of which are the following:
 \begin{itemize}
-    \item Conflicts.
-    \item Specifying the entire task tree before execution
-        gives better scheduling.
-    \item Combination of locality, weight, and availability.
-    \item Compact implementation.
+    \item {\em Priorities}: The current implementation of
+        QuickSched does not take the resource locks into
+        account when selecting tasks in the queues,
+    \item {\em Work-stealing}: During work-stealing, the
+        queues are probed in a random order although
+        the total relative cost of the tasks in the queue,
+        as well as the length of their critical paths are
+        known,
+    \item {\em Costs}: The size of the resources used by
+        a task are currently not taken into account when
+        assigning it to the queues in {\tt qsched\_enqueue},
+        or when approximating the cost of a task.
 \end{itemize}
 
-Possible improvements
-\begin{itemize}
-    \item Use number of locks as well when selecting from a queue.
-    \item Use total/max queue weight when selecting for work-stealing.
-    \item Use different resource sizes when computing queue score.
-    \item Dynamically add new tasks, but potentially wrecks havoc
-        with the weights.
-\end{itemize}
+QuickSched is distributed under the GNU Lesser General Public Licence
+v\,3.0 and is available for download via
+\url{http://quicksched.sourceforge.net}.
 

### 2013-11-10 examples/mkplots_tasks.m
minor changes.

@@ -48,7 +48,7 @@ print -depsc2 tasks_mm_dynamic.eps
 
 %% Plot the task timelines for tasks allocation
 % Load the data
-tasks = importdata( 'test.dump' );
+tasks = importdata( 'test_qr.tasks' );
 tasks(:,6) = ( tasks(:,6) - tasks(:,5) ) * tpms;
 start = min( tasks(:,5) );
 tasks(:,5) = ( tasks(:,5) - start ) * tpms;
@@ -72,15 +72,50 @@ end
 hold off;
 xlabel('time (ms)');
 ylabel('core ID');
-set(gca,'YTick',1:(max(tasks(:,1))+1))
-title('tiled QR decomposition tasks');
-axis([ 0 , max( tasks(:,5) + tasks(:,6) ) , -0.5 , nr_cores-0.5 ]);
+set(gca,'YTick',0:(max(tasks(:,2))))
+title('QuickSched tiled QR decomposition tasks');
+axis([ 0 , 110 , -0.5 , nr_cores-0.5 ]);
 
 % Print this plot
 set( gcf , 'PaperSize' , 2.3*[ 16 4 ] );
 set( gcf , 'PaperPosition' , 2.3*[ 0.25 0.25 16 4 ] );
-print -depsc2 tasks_qr_dynamic.eps
-!epstopdf tasks_qr_dynamic.eps 
+print -depsc2 figures/tasks_qr.eps
+
+
+
+%% Plot the task timelines for tasks allocation (OmpSs)
+% Load the data
+tasks = importdata( 'test_qr_ompss.tasks' );
+tasks(:,4) = ( tasks(:,4) - tasks(:,3) ) * tpms;
+start = min( tasks(:,3) );
+tasks(:,3) = ( tasks(:,3) - start ) * tpms;
+nr_cores = max( tasks(:,1) ) + 1;
+
+% Init the plot
+clf;
+subplot('position',[ 0.05 , 0.1 , 0.9 , 0.8 ]);
+colours = [ 255 34 0 ; 130 255 0 ; 0 184 255 ; 255 237 0 ] / 255;
+hold on;
+
+% Plot the tasks
+for k=1:size(tasks,1)
+    c = colours( tasks(k,2)+1 , : );
+    rectangle( 'Position' , [ tasks(k,3) , tasks(k,1)-0.5 , tasks(k,4) , 1 ] , ...
[…]

### 2013-11-10 examples/runs_qr.sh
change tile size.

@@ -2,7 +2,7 @@
 
 # Executable and parameters
 prog=test_qr
-params="-m 32 -n 32 -r 10"
+params="-m 32 -n 32 -k 64 -r 10"
 
 # Main loop
 for n in {1..64}

### 2013-11-12 paper/paper.tex
fixed download url.
updated ompss results.

@@
 
-The same decomposition was implementing using OmpSs v.\,1.99.0,
+The same decomposition was implemented using OmpSs v.\,1.99.0,
 calling the kernels directly using {\tt \#pragma omp task}
-annotations with the respective dependencies.
-The scaling and efficiency relative to QuickSched are 
+annotations with the respective dependencies, and
+the runtime parameters
+\begin{quote}
+    \tt --disable-yield --schedule=socket --cores-per-socket=16 --num-sockets=4
+\end{quote}
+\noindent The scaling and efficiency relative to QuickSched are 
 shown in \fig{QRResults} as well.
@@
 individual tasks, as shown in \fig{QRTasks},
-for a smaller $1024\times 1024$ matrix on 16 cores of the
-same hardware.
+which shows the assignment of the different tasks to cores for the
+64 core run.
 The most visible difference between both schedulers is that
@@
     \caption{Task scheduling in QuickSched (above) and OmpSs (below)
-        for a $1024\times 1024$ matrix on 16 cores.
+        for a $2048\times 2048$ matrix on 64 cores.
         The task colors correspond to those in \fig{QR}.}
@@
 v\,3.0 and is available for download via
-\url{http://quicksched.sourceforge.net}.
+\url{https://sourceforge.net/projects/quicksched/files/}.
 

### 2013-11-13 src/queue.c
check heap in both directions.

@@ -89,20 +89,33 @@ int queue_get ( struct queue *q , struct qsched *s , int insist ) {
     
         /* Fix the heap. */
         w = tasks[ inds[k] ].weight;
-        while ( 1 ) {
-            if ( ( j = 2*k + 1 ) >= count )
-                break;
-            if ( j+1 < count && tasks[ inds[j] ].weight < tasks[ inds[j+1] ].weight )
-                j = j + 1;
-            if ( tasks[ inds[j] ].weight > w ) {
-                temp = inds[j];
-                inds[j] = inds[k];
-                inds[k] = temp;
-                k = j;
+        if ( k > 0 && w > tasks[ inds[k/2] ].weight )
+            while ( k > 0 ) {
+                j = k / 2;
+                if ( w > tasks[ inds[j] ].weight ) {
+                    temp = inds[j];
+                    inds[j] = inds[k];
+                    inds[k] = temp;
+                    k = j;
+                    }
+                else
+                    break;
                 }
-            else
-                break;
-            }
+        else
+            while ( 1 ) {
+                if ( ( j = 2*k + 1 ) >= count )
+                    break;
+                if ( j+1 < count && tasks[ inds[j] ].weight < tasks[ inds[j+1] ].weight )
+                    j = j + 1;
+                if ( tasks[ inds[j] ].weight > w ) {
+                    temp = inds[j];
+                    inds[j] = inds[k];
+                    inds[k] = temp;
+                    k = j;
+                    }
+                else
+                    break;
+                }
         
         } /* did we get a task? */
         

### 2013-11-13 src/queue.c
fix indexing errror.

@@ -89,9 +89,9 @@ int queue_get ( struct queue *q , struct qsched *s , int insist ) {
     
         /* Fix the heap. */
         w = tasks[ inds[k] ].weight;
-        if ( k > 0 && w > tasks[ inds[k/2] ].weight )
+        if ( k > 0 && w > tasks[ inds[(k-1)/2] ].weight )
             while ( k > 0 ) {
-                j = k / 2;
+                j = (k - 1) / 2;
                 if ( w > tasks[ inds[j] ].weight ) {
                     temp = inds[j];
                     inds[j] = inds[k];

### 2014-07-29 examples/test_bh_sorted.c
Added more logic in iact_pair() and create_tasks() to only create pair tasks when we can't decide between a p-p and p-m interaction.

@@ -37,7 +37,7 @@
 
 /* Some local constants. */
 #define cell_pool_grow 1000
-#define cell_maxparts 16
+#define cell_maxparts 32
 #define task_limit 5000
 #define const_G 6.6738e-8
 #define dist_min 0.5 /* Used fpr legacy walk only */
@@ -989,12 +989,44 @@ void iact_pair(struct cell *ci, struct cell *cj) {
     if (ci->split && cj->split) {
 
       /* Let's split both cells and build all possible pairs */
-      for (cp = ci->firstchild; cp != ci->sibling; cp = cp->sibling)
-	for (cps = cj->firstchild; cps != cj->sibling; cps = cps->sibling)
-	  iact_pair(cp, cps);
+      for (cp = ci->firstchild; cp != ci->sibling; cp = cp->sibling) {
+	for (cps = cj->firstchild; cps != cj->sibling; cps = cps->sibling) {
+	  
+	  /* Check the distance between the two children */
+	  for (k = 0; k < 3; k++) {
+	    center_i = cp->loc[k] + 0.5 * cp->h;
+	    center_j = cps->loc[k] + 0.5 * cps->h;
+	    dx[k] = fabs(center_i - center_j);
+	  }
 
+	  min_dist = cp->h + cps->h;
+
+	  /* Are the cells NOT neighbours ? */
+	  if ((dx[0] > min_dist) || (dx[1] > min_dist) || (dx[2] > min_dist)) {
+	    
+	    iact_pair_pc(cp, cps);
+	    iact_pair_pc(cps, cp);
+
+	  } else {
+
+	    /* Are both children split ? */
+	    if (cp->split && cps->split) {
+
+	      /* Recurse */
+	      iact_pair(cp, cps);
+	      
+	    } else {
+
+	      /* Use direct summation */
+	      iact_pair_direct(cp, cps);
+      
+	    }
+	  }
+	}
+      }
     } else {/* Ok one of the cells is not split */
 
+      /* Use direct summation */
       iact_pair_direct(ci, cj);
     }
   }
@@ -1097,7 +1129,7 @@ void create_tasks(struct qsched *s, struct cell *ci, struct cell *cj) {
   /* Single cell? */
   if (cj == NULL) {
 
-    /* Is this cell split? */
+    /* Is this cell split and above the task limit ? */
     if (ci->split && ci->count > task_limit) {
 
       /* Loop over each of this cell's progeny. */
[…]

### 2014-07-29 examples/test_bh_sorted.c
changed the way task_limit is used, reformatted.

@@ -37,8 +37,8 @@
 
 /* Some local constants. */
 #define cell_pool_grow 1000
-#define cell_maxparts 32
-#define task_limit 5000
+#define cell_maxparts 100
+#define task_limit 1.0e9
 #define const_G 6.6738e-8
 #define dist_min 0.5 /* Used fpr legacy walk only */
 #define iact_pair_direct iact_pair_direct_sorted
@@ -263,7 +263,7 @@ void cell_sort(struct cell *c, float *axis, int aid) {
     struct {
       struct index *index;
       int offset;
-      } temp, pindex[8];
+    } temp, pindex[8];
     int k = 0;
 
     /* First, make sure all the progeny have been sorted, and get the pointers
@@ -274,7 +274,7 @@ void cell_sort(struct cell *c, float *axis, int aid) {
       pindex[k].offset = cp->parts - c->parts;
       k++;
     }
-    
+
     /* Fill any remaining gaps with the sentinel. */
     c->indices[(c->count + 1) * aid + c->count].d = FLT_MAX;
     for (; k < 8; k++)
@@ -284,8 +284,9 @@ void cell_sort(struct cell *c, float *axis, int aid) {
     for (k = 3; k >= 0; k--) {
       int j = k;
       while (j < 4) {
-        int jj = 2*j + 1;
-        if (jj + 1 < 8 && pindex[jj + 1].index->d < pindex[jj].index->d) jj = jj + 1;
+        int jj = 2 * j + 1;
+        if (jj + 1 < 8 && pindex[jj + 1].index->d < pindex[jj].index->d)
+          jj = jj + 1;
         if (pindex[jj].index->d < pindex[j].index->d) {
           temp = pindex[jj];
           pindex[jj] = pindex[j];
@@ -311,8 +312,9 @@ void cell_sort(struct cell *c, float *axis, int aid) {
       /* Fix the heap. */
       int j = 0;
       while (j < 4) {
-        int jj = 2*j + 1;
-        if (jj + 1 < 8 && pindex[jj + 1].index->d < pindex[jj].index->d) jj = jj + 1;
+        int jj = 2 * j + 1;
+        if (jj + 1 < 8 && pindex[jj + 1].index->d < pindex[jj].index->d)
+          jj = jj + 1;
         if (pindex[jj].index->d < pindex[j].index->d) {
           temp = pindex[jj];
           pindex[jj] = pindex[j];
[…]

### 2014-07-29 paper/paper.tex
refactoring introduction, added design considerations.

@@
-% v2-acmsmall-sample.tex, dated March 6 2012
-% This is a sample file for ACM small trim journals
-%
-% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
-% (c) 2010 Association for Computing Machinery (ACM)
-%
-% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
-% Users can also go through the FAQs available on the journal's submission webpage.
-%
-% Steps to compile: latex, bibtex, latex latex
-%
-% For tracking purposes => this is v1.3 - March 2012
-
-\documentclass[prodmode,acmtopc]{acmsmall} % Aptara syntax
+\documentclass[preprint]{elsarticle}
 
@@
 
-% Metadata Information
-\acmVolume{0}
-\acmNumber{0}
-\acmArticle{0}
-\acmYear{2014}
-\acmMonth{0}
-
 % My own macros
 \usepackage{epsfig}
-%\usepackage{hyperref}
+\usepackage{url}
 \usepackage{listings}
@@
 
-% Page heads
-\markboth{P.~Gonnet}{QuickSched: Task-based parallelism with dependencies and conflicts}
-
 % Title portion
 \title{QuickSched: Task-based parallelism with dependencies and conflicts}
-\author{PEDRO GONNET
-\affil{Durham University}}
-% NOTE! Affiliations placed here should be for the institution where the
-%       BULK of the research was done. If the author has gone to a new
-%       institution, before publication, the (above) affiliation should NOT be changed.
-%       The authors 'current' address may be given in the "Author's addresses:" block (below).
-%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
-%       currently affiliated with NASA.
+\author[ecs,ggl]{Pedro Gonnet\corref{cor}} \ead{pedro.gonnet@durham.ac.uk}
+\author[ecs]{Aidan B. G. Chalk} \ead{aidan.chalk@durham.ac.uk}
+\author[icc]{Matthieu Schaller} \ead{matthieu.schaller@durham.ac.uk}
+\cortext[cor]{Corresponding author}
+\address[ecs]{School of Engineering and Computing Sciences, Durham University, South Road, Durham DH1 3LE, United Kingdom}
+\address[icc]{Institute for Computational Cosmology, Durham University, South Road, Durham DH1 3LE, United Kingdom}
+\address[ggl]{Google Switzerland GmbH, Brandshenkestr. 110, 8002 Z\"urich, Switzerland}
 
@@
 
-\category{???}{Computing methodologies}{Shared memory algorithms \and Concurrent algorithms}
-
-\terms{Algorithms, Design, Performance}
-
-\keywords{Task-based parallelism}
-
-\acmformat{Pedro Gonnet, 2014.
-QuickSched: Task-based parallelism with dependencies and conflicts.}
-% At a minimum you need to supply the author names, year and a title.
-% IMPORTANT:
-% Full first names whenever they are known, surname last, followed by a period.
-% In the case of two authors, 'and' is placed between them.
-% In the case of three or more authors, the serial comma is used, that is, all author names
-% except the last one but including the penultimate author's name are followed by a comma,
-% and then 'and' is placed before the final author's name.
-% If only first and middle initials are known, then each initial
-% is followed by a period and they are separated by a space.
-% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.
-
-\begin{bottomstuff}
-This work was supported by a Durham University Seedcorn Grant.
-
-Author's address: P. Gonnet, School of Engineering and Computing Sciences,
-Durham University, South Road, Durham, DH1 3LE, United Kingdom.
-\end{bottomstuff}
+\begin{keyword}
+    task-based parallelism \sep shared-memory parallelism \sep high performance computing \sep parallel computing
+\end{keyword}
 
@@
 to complete.
-Although simple to use, this implicit dependency management
-limits the types of DAGs that can be represented, i.e.~for
-the example in \fig{Tasks}, using such a spawning model
-would create implicit dependencies between the lowest-level
-tasks $C$, $E$, and $K$.\footnote{
-The main thread spawns tasks $A$, $G$ and $J$, $A$ spawns $B$ and $D$,
-$G$ spawns $F$, $H$, and then $I$, $B$ spawns $C$.
-The main thread then has to {\tt sync} for $A$, $G$, and $J$,
-and thus implicitly all their spawned tasks, before executing
-$E$ and $K$.}
 
@@
 of multiple conflicts per task.
-Both \citeN{ref:Ltaief2012} and \citeN{ref:Agullo2013} note
+Both \cite{ref:Ltaief2012} and \cite{ref:Agullo2013} note
 this problem in their respective implementations of the Fast Multipole
@@
 all of these goals.
-Section~2 describes the underlying algorithms and data structures,
-and Section~3 describes their specific implementation in QuickSched.
-Section~4 presents two test-cases:
+Section~2 describes the main design considerations, Section~3 the
+underlying algorithms and data structures, and
+Section~4 their specific implementation in QuickSched.
+Section~5 presents two test-cases:
 \begin{enumerate}
@@
     gravitational N-body problem similar to the FMM codes
-    of \citeN{ref:Ltaief2012} and \citeN{ref:Agullo2013},
+    of \cite{ref:Ltaief2012} and \cite{ref:Agullo2013},
 \end{enumerate}
@@
 and can be used to assess its efficiency.
-Section~5 concludes with some general observations and future work
+Section~6 concludes with some general observations and future work
 directions.
@@
 \end{figure}
+
+\section{Design Considerations}
+
+\begin{itemize}
+  \item Two flavours of dependency generation: Through spawning as in Cilk,
+    or automatically as per QUARK, OmpSS, etc...
+  \item Limitation of spawning with complex task hierarchies.
+  \item Limitation of dependency guessing is sometimes over-protective dependencies.
+  \item We let the user define dependencies explictly.
+  \item Con: Some effort needed to get dependencies right,
+  \item Pro: All dependencies known at the start of a computation, better
+    scheduling decisions can be made (see QR).
+  \item A word on task granularity...
+\end{itemize}
+
+From a programmer's perspective, there are two main paradigms for generating
+task dependencies:
+\begin{itemize}
+  \item Implicitly via spawning and waiting, e.g.~as in Cilk, or
+  \item Automatic extraction from data dependencies.
+\end{itemize}
+
+The first scheme, spawning and waiting, is arguably the simplest to
+use.
+For simple depedency structures in which each task depends on only a
+single task, i.e.~if the task DAG is a tree, each task {\em spawns}, or
+creates, its dependent tasks after its completion.
+Hence for the tasks $A$--$E$ in \fig{Tasks}, task $A$ would spawn
+tasks $B$ and $D$, task $B$ would spawn task $C$, and task $D$ would
+spawn task $E$.
+If a task has more than one dependency, e.g. tasks $D$--$F$ in \fig{Tasks},
+then the task generation order is reversed: Task $E$ is executed first,
+and first spawns tasks $D$ and $F$, and waits for both their completion
+before doing its own computations.
+
+Although simple to use, this implicit dependency management
+limits the types of DAGs that can be represented, i.e.~for
+all the tasks in \fig{Tasks}, using such a spawning and waiting model
+would create implicit dependencies between the lowest-level
+tasks $C$, $E$, and $K$.
+The main thread would spawn tasks $A$, $G$ and $J$, $A$ spawns $B$ and $D$,
+$G$ spawns $F$, $H$, and then $I$, $B$ spawns $C$.
+The main thread then has to wait for $A$, $G$, and $J$,
+and thus implicitly all their spawned tasks, before executing
+$E$ and $K$.
+
+Automatic dependency extraction works by enforcing dependencies
+between tasks that access the same data.
+These data dependencies are provided explicitly by the programmer, e.g.~by
+describing which parameters to a task are input, output, and input/output.
+The dependencies are enforced in the order in which the tasks are 
+created.
+This approach usually relies on compiler extensions, e.g.~ {\tt pragma}s
+in C/C++, or a system of runtime wrappers, to describe the task parameters
+and their intent.
+
+This approach allows programmers to specify rich dependency hierarchies
+with very little effort, they still only allow for one type of relationship,
+i.e.~dependencies, and lack the ability to deal with conflicts as
+described in the previous section.
+
+Here, QuickSched takes a drastically different approach.
+The programmer creates the task DAG explicitly before execution.
 
@@
 the tasks topologically according to their dependencies, i.e.
-as per \citeN{ref:Kahn1962}:
+as per \cite{ref:Kahn1962}:
 \begin{center}\begin{minipage}{0.9\textwidth}
@@
 
+\section{Validation}
+
+This section presents two test cases showing
+how QuickSched can be used in real-world applications, and
+providing benchmarks to assess its efficiency and scalability.
+
+The first test is the tiled QR decomposition originally
+described in \cite{ref:Buttari2009}, which has been used as a benchmark
+by other authors \cite{ref:Agullo2009b,ref:Badia2009,ref:Bosilca2012}.
+This example only requires dependencies and is presented 
+as a point of comparison to existing task-based parallel
+programming infrastructures.
+
+The second example is a Barnes-Hut tree-code, a problem
+similar to the Fast Multipole Method described in both
+\cite{ref:Ltaief2012} and \cite{ref:Agullo2013}.
+This example shows how conflicts, modeled
+via hierarchical resources, can be useful.
+
+The source code of both examples is distributed with the
+QuickSched library, along with scripts to run the benchmarks
+and generate the plots used in the following.
+All examples were compiled with gcc v.\,4.8.1 using the
+{\tt -O2 -march=native} flags and run on a 
+tests were run on a 64-core AMD Opteron 6376 machine running
+at 2.6\,GHz.
+
+\subsection{Task-Based QR Decomposition}
+
+\cite{ref:Buttari2009} introduced the concept of using task-based
+parallelism for tile-based algorithms in numerical linear algebra,
+presenting parallel codes for the Cholesky, LU, and QR
+decompositions.
+These algorithms are now part of the PLASMA and MAGMA
+libraries for parallel linear algebra \cite{ref:Agullo2009}.
+The former uses the QUARK task scheduler, which was originally
+designed for it, while the latter currently uses the StarPU
+task scheduler.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/QR.pdf,width=0.8\textwidth}}
+    \caption{Task-based QR decomposition of a matrix consisting
+        of $4\times 4$ tiles.
+        Each circle represents a tile, and its color represents
+        the type of task on that tile at that level.
+        Empty circles have no task associated with them.
+        The arrows represent dependencies at each level, and
+        tasks at each level also implicitly depend on the
+        task at the same location in the previous level.}
+    \label{fig:QR}
+\end{figure}
+
+The tiled QR factorization is based on four basic tasks,
+or kernels, as shown in \fig{QR}.
+For a matrix consisting of $N\times N$ tiles, $N$ passes,
+or levels, are computed, each computing a column and row of the QR
+decomposition.
+The tasks can be defined in terms of the tuples $(i,j,k)$,
+where $i$ and $j$ are the row and column of the tile, respectively,
+and $k$ is its level:
+
+\begin{center}
+    \begin{tabular}{llll}
+        Task & where & depends on task(s) & locks tile(s) \\
+        \hline
+        \epsfig{file=figures/TaskRed.pdf,height=9pt} DGEQRF & $i=j=k$ & $(i,j,k-1)$ & $(i,j)$ \\
+        \epsfig{file=figures/TaskGreen.pdf,height=9pt} DLARFT & $i=k$, $j>k$ & $(i,j,k-1)$, $(k,k,k)$ & $(i,j)$ \\
+        \epsfig{file=figures/TaskBlue.pdf,height=9pt} DTSQRF & $i>k$, $j=k$ & $(i,j,k-1)$, $(i-1,j,k)$ & $(i,j)$, $(j,j)$ \\
+        \epsfig{file=figures/TaskYellow.pdf,height=9pt} DSSRFT & $i>k$, $j>k$ & $(i,j,k-1)$, $(i-1,j,k)$, $(i,k,k)$ & $(i,j)$ \\
+        \hline
+    \end{tabular}
+\end{center}
+
+\noindent where the task names are the BLAS-like operation
+performed on the given tiles.
+Every task depends on the task at the same position and the
+previous level, i.e.~the task $(i,j,k)$ always depends on
+$(i,j,k-1)$ for $k>1$.
+Each task also modifies its own tile $(i,j)$, and the DTSQRF
+task additionally modifies the lower triangular part of the $(j,j)$th tile.
+
+The QR decomposition was computed for a $2048\times 2048$
+random matrix using tiles of size $64\times 64$ using QuickSched
+as described above.
+For the DGEQRF and DLARFT kernels, the respective routines
+from the LAPACKE\footnote{\url{http://www.netlib.org/lapack/lapacke.html}}
+library were used.
+The DTSQRF and DSSRFT kernels were implemented by Sam Townsend
+as part of his MSc thesis in Computer Science at Durham University.
+For this matrix, a total of 11440 tasks with 32240 dependencies
+were generated.
+
+For these tests, OpenMP parallelism and resource re-owning
+were used with one queue per core.
+The QR decomposition was computed 10 times for each number of
+cores, and the average thereof taken for the scaling and
+efficiency results in \fig{QRResults}.
+The timings are for {\tt qsched\_run}, including the cost of
+{\tt qsched\_start}, which does not run in parallel.
+Setting up the scheduler, tasks, and resources took, in all
+cases, an average of 7.2\,ms.
+
+The same decomposition was implemented using OmpSs v.\,1.99.0,
+calling the kernels directly using {\tt \#pragma omp task}
+annotations with the respective dependencies, and
+the runtime parameters
+\begin{quote}
+    \tt --disable-yield --schedule=socket --cores-per-socket=16 --num-sockets=4
+\end{quote}
+\noindent The scaling and efficiency relative to QuickSched are 
+shown in \fig{QRResults} as well.
+The difference in timings is the result of the different
+task scheduling policies, as well as a smaller lag between the
+individual tasks, as shown in \fig{QRTasks},
+which shows the assignment of the different tasks to cores for the
+64 core run.
+The most visible difference between both schedulers is that
+the DGEQRF tasks (in red) are scheduled as soon as they
+become available in QuickSched, thus preventing bottlenecks
+near the end of the computation.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/QR_scaling.pdf,width=0.9\textwidth}}
+    \caption{Strong scaling and parallel efficiency of the tiled QR decomposition
+        computed over a $2048\times 2048$ matrix with tiles of size
+        $64\times 64$.
+        The QR decomposition with QuickSched takes 233\,ms,
+        achieving 73\% parallel efficiency, over all 64 cores.
+        The scaling and efficiency for OmpSs are computed relative to QuickSched.
+        }
+    \label{fig:QRResults}
+\end{figure}
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/tasks_qr.pdf,width=0.9\textwidth}}
+    \centerline{\epsfig{file=figures/tasks_qr_ompss.pdf,width=0.9\textwidth}}
+    \caption{Task scheduling in QuickSched (above) and OmpSs (below)
+        for a $2048\times 2048$ matrix on 64 cores.
+        The task colors correspond to those in \fig{QR}.}
+    \label{fig:QRTasks}
+\end{figure}
+
+\subsection{Task-Based Barnes-Hut N-Body Solver}
+
+The Barnes-Hut tree-code is an algorithm to approximate the
+solution of an $N$-body problem, i.e.~computing all the
+pairwise interactions between a set of $N$ particles,
+in \oh{N\log N} operations, as opposed to the \oh{N^2}
+naive direct computation.
+The algorithm is based on a recursive octree decomposition:
+Starting from a cubic cell containing all the particles,
+the cell is recursively bisected along all three spatial dimensions,
+resulting in eight sub-cells, until the number of particles
+per cell is smaller than some limit $n_\mathsf{max}$.
+
+The particle interactions can also be formulated recursively:
+Given a particle and a set of particles in a cell,
+if the particle and cell
+are sufficiently well separated, the particle-cell interactions
+are approximated by interacting the particle with the cell's
+center of mass.
+If the particle and the cell are too close, and the cell
+has sub-cells, i.e.~it contained more than $n_\mathsf{max}$
+particles and was split in the recursive octree decomposition,
+then the particle is interacted with each of the sub-cells
+recursively.
+Finally, if the cell is not split, i.e.~it is a leaf cell
+in the octree, then the particle is interacted with all
+particles in the cell, except for the particle itself if
+it is in the same cell.
+This operation is performed for each particle, starting
+with the root-level cell containing all the particles.
+
+The particle data is sorted hierarchically, following the
+octree structure.
+Unlike in many codes, where the leaves store an array of
+pointers to the underlying particles, the cells, at all
+levels, store only a pointer to the first of their own particles,
+and the total number of particles.
+The current approach, illustrated in \fig{CellParts} is not
+only more compact, it also allows a direct and more cache-efficient access
+to the list of particles for any inner node of the tree.
+The cost of sorting the particles, with a recursive
+partitioning similar to QuickSort \cite{ref:Hoare1962},
+is in \oh{N\log N}.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/CellParts.pdf,width=0.9\textwidth}}
+    \caption{Hierarchical ordering of the particle data structures
+    (right) according to their cell (left).
+    Each cell has a pointer to the first of its particles (same color
+    as cells) in the same global parts array.}
+    \label{fig:CellParts}
+\end{figure}
+
+The task-based implementation will consist of four
+types of tasks:
+\begin{itemize}
+    \item {\em Self}-interactions in which all particles
+        in a single cell interact with all other particles in the
+        same cell,
+    \item {\em Particle-particle} pair interactions in which
+        all particles in a cell interact with all
+        particles in another cell,
+    \item {\em Particle-cell} pair interactions in which
+        all particles in one cell are interacted with the
+        center of mass of another cell, and
+    \item {\em Center of mass} tasks, which compute
+        the center of mass of a single cell either from
+        the sum of the centers of mass of its sub-cells
+        if it has been split, or directly from the
+        particles otherwise.
+\end{itemize}
+
+These tasks can be created recursively over the cell hierarchy
+as shown in the function {\tt make\_tasks} in \fig{MakeTasks}.
+The function is called on the root cell with the root cell
+and {\tt NULL} as its two cell parameters.
+The function recurses as follows:
+\begin{itemize}
+    \item If called with a single (line~6), split (line~7) cell,
+        recurse over all the cell's sub-cells (line~9), and all
+        pairs of the cell's sub-cells (line~11),
+    \item If called with a single unsplit cell (line~13),
+        create a self-interaction task on that cell (line~14),
+    \item If called with two cells that are sufficiently well
+        separated (line~21), create two particle-cell pair
+        interactions (lines~23 and~28) over both cells in
+        opposite orders, which depend on the center of mass
+        task of each cell,
+    \item If called with two cells that are not well
+        separated and both cells are split (line~33),
+        recurse over all pairs of sub-cells spanning
+        both cells (line~37), and
+    \item If called with two cells that are not well separated
+        and either of the cells are not split, create
+        a particle-particle pair task over both cells.
+\end{itemize}
+\noindent where every interaction task additionally locks
+the cells on which it operates (lines~16, 25, 30, and 42--43).
+
+In order to prevent generating
+a large number of very small tasks, the task generation only recurses
+if the cells contain more than a minimum number $n_\mathsf{task}$
+of threads each (lines~7 and~34).
+The tasks themselves are then left to recurse over the sub-trees,
+which is why in these cases, the tasks are made to depend on the
+center of mass tasks (lines~17--18 and~41--47)
+which may be used in the ensuing interactions.
+
+This Barnes-Hut tree-code was used to approximate the gravitational
+N-Body problem for 1\,000\,000 particles with random coordinates
+in $[0,1]^3$.
+The parameters $n_\mathsf{max}=100$ and $n_\mathsf{task}=5000$
+were used to generate the tasks, and cell pairs were considered
+well separated if not directly adjacent.
+Using the above scheme generated 97\,553 tasks, of which
+512 self-interaction tasks, ??? particle-particle interaction
+task, ??? particle-cell interaction tasks, and 37\,449
+center of mass tasks.
+A total of 141\,840 dependencies were generated, along with
+104\,392 locks on 37\,449 resources.
+
+For these tests, OpenMP parallelism was used and resource
+re-owning was switched off.
+Resource ownership was attributed by dividing the global
+{\tt parts} array by the number of queues and assigning each cell's
+{\tt res} to the fraction of the {\tt parts} array to which
+the first of its own {\tt parts} belong.
+The interactions were computed 10 times for each number of
+cores, and the average thereof taken for the scaling and
+efficiency results in \fig{BHResults}.
+The timings are for {\tt qsched\_run}, including the cost of
+{\tt qsched\_start}, which does not run in parallel.
+Setting up the scheduler, tasks, and resources took, in all
+cases, an average of 51.3\,ms.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/BH_scaling.pdf,width=0.9\textwidth}}
+    \caption{Strong scaling and parallel efficiency of the Barnes-Hut tree-code
+        computed over 1\,000\,000 particles.
+        Solving the N-Body problem takes 3.3\,s, achieving 60\% parallel
+        efficiency, over all 64 cores.
+        }
+    \label{fig:BHResults}
+\end{figure}
+
+Unlike the QR decomposition, the results scale well only to
+32 cores, achieving 90\% parallel efficiency, and then
+level off for increasing numbers of cores.
+This, however, is not a problem of the task-based parallel
+algorithm, or of QuickSched, but of the memory bandwidth
+of the underlying hardware.
+\fig{BHTasks} shows the accumulated cost of each task type and of 
+QuickSched over the number of cores.
+At 64 cores, the scheduler overheads account for only 4.7\% of
+the total computational cost, whereas,
+as of 30 cores, the cost of both pair types grow by up to
+50\%.
+This is most probably due to memory bandwidth restrictions, as
+the cost of the self interaction tasks, which do twice as much
+computation per memory access, only grow by up to 15\%.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/BH_times.pdf,width=0.8\textwidth}}
+    \caption{Accumulated cost of each task type and of the overheads
+        associated with {\tt qsched\_gettask}.
+        As of $\sim 30$ cores, the cost of both pair interaction task
+        types grow by up to 50\%.
+        The cost of the self interactions, which entail twice as much
+        computation per memory access, grow only by at most 15\%.
+        The scheduler overheads make up less than 5\% of the total
+        time.
+        }
+    \label{fig:BHTimes}
+\end{figure}
+
+\section{Discussion and Conclusions}
+
+The task scheduler described in the previous sections, QuickSched,
+differs from existing task-based programming schemes
+in a number of ways.
+The most obvious such difference is the addition of {\em conflicts},
+modeled using exclusive locks on hierarchical resources.
+This extension to the standard dependencies-only model
+of task-based parallelism allows for more complex task relations,
+such as in the Barnes-Hut tree-code described earlier.
+
+Another significant difference is that the tasks, their
+dependencies, and their conflicts must be described
+{\em explicitly} before the parallel computation starts.
+This as opposed to implicit dependencies generated
+by task spawning, e.g.~as in Cilk, or to extracting the
+dependencies from the task parameters, e.g.~in QUARK or OmpSs.
+Explicitly defining dependencies has the advantage that 
+more elaborate dependency structures can be generated.
+Furthermore, knowing the structure of the entire task
+graph from the start of the computation provides valuable
+information when scheduling the tasks, e.g.~using the 
+critical path along the dependencies to compute the
+task weight.
+
+Finally, as opposed to the most other task-based
+programming environments which rely on compiler extensions
+and/or code pre-processors, QuickSched operates as a regular
+C-language library, based on standard parallel functionality
+provided by OpenMP and/or {\tt pthreads}.
+This ensures a maximum amount of portability on existing
+and future architectures.
+The interfaces are also kept as simple
+as possible in order to reduce the burden on the programmer
+when implementing task-based codes.
+
+The QuickSched library itself is remarkably simple, consisting of
+less than 3\,000 lines of code, including comments.
+Both examples, which are distributed with QuickSched,
+require less than 1\,000 lines of code each.
+
+In both examples, QuickSched performs extremely well, even
+on a large number of shared-memory cores.
+This performance is due, on the one hand, to the
+division of labor between the scheduler and the queues,
+and on the other hand due to the simple yet efficient
+algorithms for task selection and resource locking.
+The task weighting based on the length of the critical
+path of the dependencies delivers, in the examples shown,
+good parallel efficiency.
+
+There are several possible improvements to QuickSched which
+have not been addressed in this paper.
+The most obvious of which are the following:
+\begin{itemize}
+    \item {\em Priorities}: The current implementation of
+        QuickSched does not take the resource locks into
+        account when selecting tasks in the queues,
+    \item {\em Work-stealing}: During work-stealing, the
+        queues are probed in a random order although
+        the total relative cost of the tasks in the queue,
+        as well as the length of their critical paths are
+        known,
+    \item {\em Costs}: The size of the resources used by
+        a task are currently not taken into account when
+        assigning it to the queues in {\tt qsched\_enqueue},
+        or when approximating the cost of a task.
+\end{itemize}
+
+QuickSched is distributed under the GNU Lesser General Public Licence
+v\,3.0 and is available for download via
+\url{https://sourceforge.net/projects/quicksched/files/}.
+
+% Acknowledgments
+\section*{Acknowledgments}
+The author would like to thank Sam Townsend for implementing the
+kernels in the task-based QR decomposition as part of his MSc Thesis
+at Durham University.
+The author also thanks Lydia Heck of the Institute for
+Computational Cosmology at Durham University for providing access
+to, and expertise on, the COSMA cluster used in the performance
+evaluation.
+This work was supported by a Durham University Seedcorn Grant.
+
+% Bibliography
+\bibliographystyle{elsarticle-num}
+\bibliography{quicksched}
+
+% Appendices
+\appendix
 \section{User Interface}
@@
 
-\section{Validation}
-
-This section presents two test cases showing
-how QuickSched can be used in real-world applications, and
-providing benchmarks to assess its efficiency and scalability.
-
-The first test is the tiled QR decomposition originally
-described in \citeN{ref:Buttari2009}, which has been used as a benchmark
-by other authors \cite{ref:Agullo2009b,ref:Badia2009,ref:Bosilca2012}.
-This example only requires dependencies and is presented 
-as a point of comparison to existing task-based parallel
-programming infrastructures.
-
-The second example is a Barnes-Hut tree-code, a problem
-similar to the Fast Multipole Method described in both
-\citeN{ref:Ltaief2012} and \citeN{ref:Agullo2013}.
-This example shows how conflicts, modeled
-via hierarchical resources, can be useful.
-
-The source code of both examples is distributed with the
-QuickSched library, along with scripts to run the benchmarks
-and generate the plots used in the following.
-All examples were compiled with gcc v.\,4.8.1 using the
-{\tt -O2 -march=native} flags and run on a 
-tests were run on a 64-core AMD Opteron 6376 machine running
-at 2.6\,GHz.
-
-\subsection{Task-Based QR Decomposition}
-
-\citeN{ref:Buttari2009} introduced the concept of using task-based
-parallelism for tile-based algorithms in numerical linear algebra,
-presenting parallel codes for the Cholesky, LU, and QR
-decompositions.
-These algorithms are now part of the PLASMA and MAGMA
-libraries for parallel linear algebra \cite{ref:Agullo2009}.
-The former uses the QUARK task scheduler, which was originally
-designed for it, while the latter currently uses the StarPU
-task scheduler.
+\section{Tiled QR Implementation}
 
 \begin{figure}
-    \centerline{\epsfig{file=figures/QR.pdf,width=0.8\textwidth}}
-    \caption{Task-based QR decomposition of a matrix consisting
-        of $4\times 4$ tiles.
-        Each circle represents a tile, and its color represents
-        the type of task on that tile at that level.
-        Empty circles have no task associated with them.
-        The arrows represent dependencies at each level, and
-        tasks at each level also implicitly depend on the
-        task at the same location in the previous level.}
-    \label{fig:QR}
-\end{figure}
-
-The tiled QR factorization is based on four basic tasks,
-or kernels, as shown in \fig{QR}.
-For a matrix consisting of $N\times N$ tiles, $N$ passes,
-or levels, are computed, each computing a column and row of the QR
-decomposition.
-The tasks can be defined in terms of the tuples $(i,j,k)$,
-where $i$ and $j$ are the row and column of the tile, respectively,
-and $k$ is its level:
-
-\begin{center}
-    \begin{tabular}{llll}
-        Task & where & depends on task(s) & locks tile(s) \\
-        \hline
-        \epsfig{file=figures/TaskRed.pdf,height=9pt} DGEQRF & $i=j=k$ & $(i,j,k-1)$ & $(i,j)$ \\
-        \epsfig{file=figures/TaskGreen.pdf,height=9pt} DLARFT & $i=k$, $j>k$ & $(i,j,k-1)$, $(k,k,k)$ & $(i,j)$ \\
-        \epsfig{file=figures/TaskBlue.pdf,height=9pt} DTSQRF & $i>k$, $j=k$ & $(i,j,k-1)$, $(i-1,j,k)$ & $(i,j)$, $(j,j)$ \\
-        \epsfig{file=figures/TaskYellow.pdf,height=9pt} DSSRFT & $i>k$, $j>k$ & $(i,j,k-1)$, $(k,j,k)$, $(i,k,k)$ & $(i,j)$ \\
-        \hline
-    \end{tabular}
-\end{center}
-
-\noindent where the task names are the BLAS-like operation
-performed on the given tiles.
-Every task depends on the task at the same position and the
-previous level, i.e.~the task $(i,j,k)$ always depends on
-$(i,j,k-1)$ for $k>1$.
-Each task also modifies its own tile $(i,j)$, and the DTSQRF
-task additionally modifies the lower triangular part of the $(j,j)$th tile.
-
-\begin{figure}
-\begin{center}\begin{minipage}{0.9\textwidth}
-    \begin{lstlisting}
-enum { tDGEQRF , tDLARFT , tDTSQRF , tDSSRFT };
-void make_tasks ( struct qsched *s , int m , int n ) {
+\begin{center}\begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}[basicstyle=\scriptsize\tt]
+enum {tDGEQRF, tDLARFT, tDTSQRF, tDSSRFT};
+void make_tasks(struct qsched *s, int m, int n) {
   int i, j, k, data[3];
-  qsched_task_t tid[ m*n ], tid_new;
-  qsched_res_t rid[ m*n ];
-  for ( k = 0 ; k < m*n ; k++ ) {
+  qsched_task_t tid[m * n], tid_new;
+  qsched_res_t rid[m * n];
+  for (k = 0; k < m * n; k++) {
     tid[k] = qsched_task_none;
-    rid[k] = qsched_addres( s , qsched_res_none );
-    }
-  for ( k = 0 ; k < m && k < n ; k++ ) {
+    rid[k] = qsched_addres(s, qsched_res_none);
+  }
+  for (k = 0; k < m && k < n; k++) {
     /* DGEQRF task at (k,k). */
     data[0] = k; data[1] = k; data[2] = k;
-    tid_new = qsched_addtask( s , tDGEQRF , qsched_flags_none ,
-      data , sizeof(int)*3 , 2 );
-    qsched_addlock( s , tid_new , rid[ k*m + k ] )
-    if ( tid[ k*m + k ] != qsched_task_none )
-      qsched_addunlock( s , tid[ k*m + k ] , tid_new );
-    tid[ k*m + k ] = tid_new;
-    for ( j = k+1 ; j < n ; j++ ) {
+    tid_new = qsched_addtask(s, tDGEQRF, qsched_flags_none, data,
+                             sizeof(int) * 3, 2);
+    qsched_addlock(s, tid_new, rid[k * m + k])
+    if (tid[k * m + k] != qsched_task_none)
+      qsched_addunlock(s, tid[k * m + k], tid_new);
+    tid[k * m + k] = tid_new;
+    for (j = k + 1; j < n; j++) {
       /* DLARFT task at (k,j). */
       data[0] = k; data[1] = j; data[2] = k;
-      tid_new = qsched_addtask( s , tDLARFT , qsched_flags_none ,
-        data , sizeof(int)*3 , 3 );
-      qsched_addlock( s , tid_new , rid[ j*m + k ] )
-      qsched_addunlock( s , tid[ k*m + k ] , tid_new );
-      if ( tid[ j*m + k ] != qsched_task_none )
-        qsched_addunlock( s , tid[ j*m + k ] , tid_new );
-      tid[ j*m + k ] = tid_new;
-      }
-    for ( i = k+1 ; i < m ; i++ ) {
+      tid_new = qsched_addtask(s, tDLARFT, qsched_flags_none, data,
+                               sizeof(int) * 3, 3);
+      qsched_addlock(s, tid_new, rid[j * m + k])
+      qsched_addunlock(s, tid[k * m + k], tid_new);
+      if (tid[j * m + k] != qsched_task_none)
+        qsched_addunlock(s, tid[j * m + k], tid_new);
+      tid[j * m + k] = tid_new;
+    }
+    for (i = k + 1; i < m; i++) {
       /* DTSQRF task at (i,k). */
       data[0] = i; data[1] = k; data[2] = k;
-      tid_new = qsched_addtask( s , tDTSQRF , qsched_flags_none ,
-        data , sizeof(int)*3 , 3 );
-      qsched_addlock( s , tid_new , rid[ k*m + i ] )
-      qsched_addlock( s , tid_new , rid[ k*m + k ] )
-      qsched_addunlock( s , tid[ k*m + k ] , tid_new );
-      if ( tid[ k*m + i ] != qsched_task_none )
-        qsched_addunlock( s , tid[ k*m + i ] , tid_new );
-      tid[ k*m + i ] = tid_new;
-      for ( j = k+1 ; j < n ; j++ ) {
+      tid_new = qsched_addtask(s, tDTSQRF, qsched_flags_none, data,
+                               sizeof(int) * 3, 3);
+      qsched_addlock(s, tid_new, rid[k * m + i])
+      qsched_addlock(s, tid_new, rid[k * m + k])
+      qsched_addunlock(s, tid[k * m + k], tid_new);
+      if (tid[k * m + i] != qsched_task_none)
+        qsched_addunlock(s, tid[k * m + i], tid_new);
+      tid[k * m + i] = tid_new;
+      for (j = k + 1; j < n; j++) {
         /* DSSRFT task at (i,j). */
         data[0] = i; data[1] = j; data[2] = k;
-        tid_new = qsched_addtask( s , tDSSRFT , qsched_flags_none ,
-          data , sizeof(int)*3 , 5 );
-        qsched_addlock( s , tid_new , rid[ j*m + i ] )
-        qsched_addunlock( s , tid[ j*m + k ] , tid_new );
-        qsched_addunlock( s , tid[ k*m + i ] , tid_new );
-        if ( tid[ j*m + i ] != qsched_task_none )
-          qsched_addunlock( s , tid[ j*m + i ] , tid_new );
-        tid[ j*m + i ] = tid_new;
+        tid_new = qsched_addtask(s, tDSSRFT, qsched_flags_none, data,
+                                 sizeof(int) * 3, 5);
+        qsched_addlock(s, tid_new, rid[j * m + i])
+        qsched_addunlock(s, tid[j * m + k], tid_new);
+        qsched_addunlock(s, tid[k * m + i], tid_new);
+        if (tid[j * m + i] != qsched_task_none)
+          qsched_addunlock(s, tid[j * m + i], tid_new);
+        tid[j * m + i] = tid_new;
       }
@@
 The following loops mirror the task generation described in
-Algorithm~2 of \citeN{ref:Buttari2009}.
+Algorithm~2 of \cite{ref:Buttari2009}.
 For each level {\tt k} (line~10), a DGEQRF task is created
@@
 
-The QR decomposition was computed for a $2048\times 2048$
-random matrix using tiles of size $64\times 64$ using QuickSched
-as described above.
-For the DGEQRF and DLARFT kernels, the respective routines
-from the LAPACKE\footnote{\url{http://www.netlib.org/lapack/lapacke.html}}
-library were used.
-The DTSQRF and DSSRFT kernels were implemented by Sam Townsend
-as part of his MSc thesis in Computer Science at Durham University.
-For this matrix, a total of 11440 tasks with 32240 dependencies
-were generated.
-
-For these tests, OpenMP parallelism and resource re-owning
-were used with one queue per core.
-The QR decomposition was computed 10 times for each number of
-cores, and the average thereof taken for the scaling and
-efficiency results in \fig{QRResults}.
-The timings are for {\tt qsched\_run}, including the cost of
-{\tt qsched\_start}, which does not run in parallel.
-Setting up the scheduler, tasks, and resources took, in all
-cases, an average of 7.2\,ms.
-
-The same decomposition was implemented using OmpSs v.\,1.99.0,
-calling the kernels directly using {\tt \#pragma omp task}
-annotations with the respective dependencies, and
-the runtime parameters
-\begin{quote}
-    \tt --disable-yield --schedule=socket --cores-per-socket=16 --num-sockets=4
-\end{quote}
-\noindent The scaling and efficiency relative to QuickSched are 
-shown in \fig{QRResults} as well.
-The difference in timings is the result of the different
-task scheduling policies, as well as a smaller lag between the
-individual tasks, as shown in \fig{QRTasks},
-which shows the assignment of the different tasks to cores for the
-64 core run.
-The most visible difference between both schedulers is that
-the DGEQRF tasks (in red) are scheduled as soon as they
-become available in QuickSched, thus preventing bottlenecks
-near the end of the computation.
-
-\begin{figure}
-    \centerline{\epsfig{file=figures/QR_scaling.pdf,width=0.9\textwidth}}
-    \caption{Strong scaling and parallel efficiency of the tiled QR decomposition
-        computed over a $2048\times 2048$ matrix with tiles of size
-        $64\times 64$.
-        The QR decomposition with QuickSched takes 233\,ms,
-        achieving 73\% parallel efficiency, over all 64 cores.
-        The scaling and efficiency for OmpSs are computed relative to QuickSched.
-        }
-    \label{fig:QRResults}
-\end{figure}
-
-\begin{figure}
-    \centerline{\epsfig{file=figures/tasks_qr.pdf,width=0.9\textwidth}}
-    \centerline{\epsfig{file=figures/tasks_qr_ompss.pdf,width=0.9\textwidth}}
-    \caption{Task scheduling in QuickSched (above) and OmpSs (below)
-        for a $2048\times 2048$ matrix on 64 cores.
-        The task colors correspond to those in \fig{QR}.}
-    \label{fig:QRTasks}
-\end{figure}
-
-\subsection{Task-Based Barnes-Hut N-Body Solver}
-
-The Barnes-Hut tree-code is an algorithm to approximate the
-solution of an $N$-body problem, i.e.~computing all the
-pairwise interactions between a set of $N$ particles,
-in \oh{N\log N} operations, as opposed to the \oh{N^2}
-naive direct computation.
-The algorithm is based on a recursive octree decomposition:
-Starting from a cubic cell containing all the particles,
-the cell is recursively bisected along all three spatial dimensions,
-resulting in eight sub-cells, until the number of particles
-per cell is smaller than some limit $n_\mathsf{max}$.
-
-The particle interactions can also be formulated recursively:
-Given a particle and a set of particles in a cell,
-if the particle and cell
-are sufficiently well separated, the particle-cell interactions
-are approximated by interacting the particle with the cell's
-center of mass.
-If the particle and the cell are too close, and the cell
-has sub-cells, i.e.~it contained more than $n_\mathsf{max}$
-particles and was split in the recursive octree decomposition,
-then the particle is interacted with each of the sub-cells
-recursively.
-Finally, if the cell is not split, i.e.~it is a leaf cell
-in the octree, then the particle is interacted with all
-particles in the cell, except for the particle itself if
-it is in the same cell.
-This operation is performed for each particle, starting
-with the root-level cell containing all the particles.
+\section{Barnes-Hut N-body Solver Implementation}
 
@@
 
-The particle data is sorted hierarchically, following the
-octree structure.
-Unlike in many codes, where the leaves store an array of
-pointers to the underlying particles, the cells, at all
-levels, store only a pointer to the first of their own particles,
-and the total number of particles.
-The current approach, illustrated in \fig{CellParts} is not
-only more compact, it also allows a direct and more cache-efficient access
-to the list of particles for any inner node of the tree.
-The cost of sorting the particles, with a recursive
-partitioning similar to QuickSort \cite{ref:Hoare1962},
-is in \oh{N\log N}.
-
 \begin{figure}
-    \centerline{\epsfig{file=figures/CellParts.pdf,width=0.9\textwidth}}
-    \caption{Hierarchical ordering of the particle data structures
-    (right) according to their cell (left).
-    Each cell has a pointer to the first of its particles (same color
-    as cells) in the same global parts array.}
-    \label{fig:CellParts}
+\begin{center}\begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}[basicstyle=\scriptsize\tt]
+void comp_com(struct cell *c) {
+  int j, k;
+  c->com[0] = 0.0; c->com[1] = 0.0; c->com[2] = 0.0;
+  c->mass = 0.0;
+  if (c->split)
+    for (k = 0; k < 8; k++) {
+      struct cell *cp = c->progeny[k];
+      for (j = 0; j < 3; j++)
+        c->com[j] += cp->com[j] * cp->mass;
+      c->mass += cp->mass;
+    }
+  else
+    for (k = 0; k < 8; k++) {
+      struct part *p = &c->parts[k];
+      for (j = 0; j < 3; j++)
+        c->com[j] += p->x[j] * p->mass;
+      c->mass += p->mass;
+    }
+  c->com[0] /= c->mass; c->com[1] /= c->mass; c->com[2] /= c->mass;
+}
+
+void comp_self(struct cell *c) {
+  int j, k;
+  if (c->split)
+    for (j = 0; j < 8; j++) {
+      comp_self(c->progeny[j]);
+      for (k = j + 1; k < 8; k++)
+        comp_pair(c->progeny[j], c->progeny[k]);
+    }
+  else
+    for (j = 0; j < c->count; j++)
+      for (k = j + 1; k < c->count; k++)
+        interact c->parts[j] and c->parts[k].
+}
+
+void comp_pair(struct cell *ci, struct cell *cj) {
+  int j, k;
+  if (ci and cj well separated) {
+    comp_pair_pc(ci, cj);
+    comp_pair_pc(cj, ci);
+  } else if (ci->split && cj->split)
+    for (j = 0; j < 8; j++)
+      for (k = 0; k < 8; k++)
+        comp_pair(ci->progeny[j], cj->progeny[k]);
+  else
+    for (j = 0; j < ci->count; j++)
+      for (k = 0; k < cj->count; k++)
+        interact ci->parts[j] and cj->parts[k].
+}
+
+void comp_pair_cp(struct cell *ci, struct cell *cj) {
+  int k;
+  for (k = 0; k < ci->count; k++)
+    interact ci->parts[k] and cj center of mass.
+}
+    \end{lstlisting}
+\end{minipage}\end{center}
+    \caption{Task functions for the Barnes-Hut tree-code.}
+    \label{fig:BHTasks}
 \end{figure}
 
-The task-based implementation will consist of four
-types of tasks:
-\begin{itemize}
-    \item {\em Self}-interactions in which all particles
-        in a single cell interact with all other particles in the
-        same cell,
-    \item {\em Particle-particle} pair interactions in which
-        all particles in a cell interact with all
-        particles in another cell,
-    \item {\em Particle-cell} pair interactions in which
-        all particles in one cell are interacted with the
-        center of mass of another cell, and
-    \item {\em Center of mass} tasks, which compute
-        the center of mass of a single cell either from
-        the sum of the centers of mass of its sub-cells
-        if it has been split, or directly from the
-        particles otherwise.
-\end{itemize}
-
 \begin{figure}
 \begin{center}\begin{minipage}{0.9\textwidth}
-    \begin{lstlisting}
+    \begin{lstlisting}[basicstyle=\scriptsize\tt]
 enum { tSELF , tPAIR_PP , tPAIR_PC , tCOM };
-void make_tasks ( struct qsched *s , struct cell *ci , struct cell *cj ) {
+void make_tasks(struct qsched *s, struct cell *ci, struct cell *cj) {
   int j, k;
@@
   struct cell *data[2];
-  if ( cj == NULL ) {
-    if ( ci->split && ci->count > n_task )
-      for ( j = 0 ; j < 8 ; j++ ) {
-        make_tasks( s , ci->progeny[j] , NULL );
-        for ( k = j+1 ; k < 8 ; k++ )
-          make_tasks( s , ci->progeny[j] , ci->progeny[k] );
+  if (cj == NULL) {
+    if (ci->split && ci->count > n_task)
+      for (j = 0; j < 8; j++) {
+        make_tasks(s, ci->progeny[j], NULL);
+        for (k = j + 1; k < 8; k++)
+          make_tasks(s, ci->progeny[j], ci->progeny[k]);
       }
     else {
-      tid = qsched_addtask( s , tSELF , qsched_flags_none ,
-        &ci , sizeof(struct cell *) , ci->count*ci->count );
-      qsched_addlock( s , tid , ci->res );
-      if ( ci->split )
-        qsched_addunlock( s , ci->com , tid );
-      }
-    }
-  else if ( ci and cj are well separated ) {
+      tid = qsched_addtask(s, tSELF, qsched_flags_none, &ci,
+                           sizeof(struct cell *), ci->count * ci->count);
+      qsched_addlock(s, tid, ci->res);
+      if (ci->split) qsched_addunlock(s, ci->com, tid);
+    }
+  } else if (ci and cj are well separated) {
     data[0] = ci; data[1] = cj;
-      tid = qsched_addtask( s , tPAIR_PC , qsched_flags_none ,
-        data , sizeof(struct cell *)*2 , ci->count );
-      qsched_addlock( s , tid , ci->res );
-      qsched_addunlock( s , cj->com , tid );
+    tid = qsched_addtask(s, tPAIR_PC, qsched_flags_none, data,
+                         sizeof(struct cell *) * 2, ci->count);
+    qsched_addlock(s, tid, ci->res);
+    qsched_addunlock(s, cj->com, tid);
     data[0] = cj; data[1] = ci;
-      tid = qsched_addtask( s , tPAIR_PC , qsched_flags_none ,
-        data , sizeof(struct cell *)*2 , cj->count );
-      qsched_addlock( s , tid , cj->res );
-      qsched_addunlock( s , ci->com , tid );
-      }
-  else if ( ci->split && cj->split && 
-     ci->count*cj->count > n_task*n_task )
-    for ( j = 0 ; j < 8 ; j++ )
-      for ( k = 0 ; k < 8 ; k++ )
-        make_tasks( s , ci->progeny[j] , cj->progeny[k] );
+    tid = qsched_addtask(s, tPAIR_PC, qsched_flags_none, data,
+                         sizeof(struct cell *) * 2, cj->count);
+    qsched_addlock(s, tid, cj->res);
+    qsched_addunlock(s, ci->com, tid);
+  } else if (ci->split && cj->split && ci->count * cj->count > n_task * n_task)
+    for (j = 0; j < 8; j++)
+      for (k = 0; k < 8; k++) make_tasks(s, ci->progeny[j], cj->progeny[k]);
   else {
     data[0] = ci; data[1] = cj;
-    tid = qsched_addtask( s , tPAIR_PP , qsched_flags_none ,
-      data , sizeof(struct cell *)*2 , ci->count*cj->count );
-    qsched_addlock( s , tid , ci->res );
-    qsched_addlock( s , tid , cj->res );
-    if ( ci->split && cj->split ) {
-      qsched_addunlock( s , ci->com , tid );
-      qsched_addunlock( s , cj->com , tid );
+    tid = qsched_addtask(s, tPAIR_PP, qsched_flags_none, data,
+                         sizeof(struct cell *) * 2, ci->count * cj->count);
+    qsched_addlock(s, tid, ci->res);
+    qsched_addlock(s, tid, cj->res);
+    if (ci->split && cj->split) {
+      qsched_addunlock(s, ci->com, tid);
+      qsched_addunlock(s, cj->com, tid);
     }
@@
     \label{fig:MakeTasks}
-\end{figure}
-
-These tasks can be created recursively over the cell hierarchy
-as shown in the function {\tt make\_tasks} in \fig{MakeTasks}.
-The function is called on the root cell with the root cell
-and {\tt NULL} as its two cell parameters.
-The function recurses as follows:
-\begin{itemize}
-    \item If called with a single (line~6), split (line~7) cell,
-        recurse over all the cell's sub-cells (line~9), and all
-        pairs of the cell's sub-cells (line~11),
-    \item If called with a single unsplit cell (line~13),
-        create a self-interaction task on that cell (line~14),
-    \item If called with two cells that are sufficiently well
-        separated (line~21), create two particle-cell pair
-        interactions (lines~23 and~28) over both cells in
-        opposite orders, which depend on the center of mass
-        task of each cell,
-    \item If called with two cells that are not well
-        separated and both cells are split (line~33),
-        recurse over all pairs of sub-cells spanning
-        both cells (line~37), and
-    \item If called with two cells that are not well separated
-        and either of the cells are not split, create
-        a particle-particle pair task over both cells.
-\end{itemize}
-\noindent where every interaction task additionally locks
-the cells on which it operates (lines~16, 25, 30, and 42--43).
-
-In order to prevent generating
-a large number of very small tasks, the task generation only recurses
-if the cells contain more than a minimum number $n_\mathsf{task}$
-of threads each (lines~7 and~34).
-The tasks themselves are then left to recurse over the sub-trees,
-which is why in these cases, the tasks are made to depend on the
-center of mass tasks (lines~17--18 and~41--47)
-which may be used in the ensuing interactions.
-
-\begin{figure}
-\begin{center}\begin{minipage}{0.9\textwidth}
-    \begin{lstlisting}
-void comp_com ( struct cell *c ) {
-  int j, k;
-  c->com[0] = 0.0; c->com[1] = 0.0; c->com[2] = 0.0;
-  c->mas = 0.0;
-  if ( c->split )
-    for ( k = 0 ; k < 8 ; k++ ) {
-      struct cell *cp = c->progeny[k]
-      for ( j = 0 ; j < 3 ; j++ ) c->com[j] += cp->com[j]*cp->mass;
-      c->mass += cp->mass;
-      }
-  else
-    for ( k = 0 ; k < 8 ; k++ ) {
-      struct part *p = &c->parts[k];
-      for ( j = 0 ; j < 3 ; j++ ) c->com[j] += p->x[j]*p->mass;
-      c->mass += p->mass;
-      }
-  c->com[0] /= c->mass; c->com[1] /= c->mass; c->com[2] /= c->mass; 
-  }
-
-void comp_self ( struct cell *c ) {
-  int j, k;
-  if ( c->split )
-    for ( j = 0 ; j < 8 ; j++ ) {
-      comp_self( c->progeny[j] );
-      for ( k = j+1 ; k < 8 ; k++ )
-        comp_pair( c->progeny[j] , c->progeny[k] );
-      }
-  else
-    for ( j = 0 ; j < c->count ; j++ )
-      for ( k = j+1 ; k < c->count ; k++ )
-        interact c->parts[j] and c->parts[k].
-  }
-
-void comp_pair ( struct cell *ci , struct cell *cj ) {
-  int j, k;
-  if ( ci and cj well separated ) {
-    comp_pair_pc( ci , cj );
-    comp_pair_pc( cj , ci );
-    }
-  else if ( ci->split && cj->split )
-    for ( j = 0 ; j < 8 ; j++ )
-      for ( k = 0 ; k < 8 ; k++ )
-        comp_pair( ci->progeny[j] , cj->progeny[k] );
-  else
-    for ( j = 0 ; j < ci->count ; j++ )
-      for ( k = 0 ; k < cj->count ; k++ )
-        interact ci->parts[j] and cj->parts[k].
-  }
-
-void comp_pair_cp ( struct cell *ci , struct cell *cj ) {
-  int k;
-  for ( k = 0 ; k < ci->count ; k++ )
-    interact ci->parts[k] and cj center of mass.
-  }
-    \end{lstlisting}
-\end{minipage}\end{center}
-    \caption{Task functions for the Barnes-Hut tree-code.}
-    \label{fig:BHTasks}
 \end{figure}
@@
 
-This Barnes-Hut tree-code was used to approximate the gravitational
-N-Body problem for 1\,000\,000 particles with random coordinates
-in $[0,1]^3$.
-The parameters $n_\mathsf{max}=100$ and $n_\mathsf{task}=5000$
-were used to generate the tasks, and cell pairs were considered
-well separated if not directly adjacent.
-Using the above scheme generated 161\,613 tasks, of which
-512 self-interaction tasks, 18\,532 particle-particle interaction
-task, 105\,120 particle-cell interaction tasks, and 37\,449
-center of mass tasks.
-A total of 179\,632 dependencies were generated, along with
-142\,696 locks on 37\,449 resources.
-
-For these tests, OpenMP parallelism was used and resource
-re-owning was switched off.
-Resource ownership was attributed by dividing the global
-{\tt parts} array by the number of queues and assigning each cell's
-{\tt res} to the fraction of the {\tt parts} array to which
-the first of its own {\tt parts} belong.
-The interactions  computed 10 times for each number of
-cores, and the average thereof taken for the scaling and
-efficiency results in \fig{BHResults}.
-The timings are for {\tt qsched\_run}, including the cost of
-{\tt qsched\_start}, which does not run in parallel.
-Setting up the scheduler, tasks, and resources took, in all
-cases, an average of 51.3\,ms.
-
-\begin{figure}
-    \centerline{\epsfig{file=figures/BH_scaling.pdf,width=0.9\textwidth}}
-    \caption{Strong scaling and parallel efficiency of the Barnes-Hut tree-code
-        computed over 1\,000\,000 particles.
-        Solving the N-Body problem takes 3.3\,s, achieving 60\% parallel
-        efficiency, over all 64 cores.
-        }
-    \label{fig:BHResults}
-\end{figure}
-
-Unlike the QR decomposition, the results scale well only to
-32 cores, achieving 90\% parallel efficiency, and then
-level off for increasing numbers of cores.
-This, however, is not a problem of the task-based parallel
-algorithm, or of QuickSched, but of the memory bandwidth
-of the underlying hardware.
-\fig{BHTasks} shows the accumulated cost of each task type and of 
-QuickSched over the number of cores.
-At 64 cores, the scheduler overheads account for only 4.7\% of
-the total computational cost, whereas,
-as of 30 cores, the cost of both pair types grow by up to
-50\%.
-This is most probably due to memory bandwidth restrictions, as
-the cost of the self interaction tasks, which do twice as much
-computation per memory access, only grow by up to 15\%.
-
-\begin{figure}
-    \centerline{\epsfig{file=figures/BH_times.pdf,width=0.8\textwidth}}
-    \caption{Accumulated cost of each task type and of the overheads
-        associated with {\tt qsched\_gettask}.
-        As of $\sim 30$ cores, the cost of both pair interaction task
-        types grow by up to 50\%.
-        The cost of the self interactions, which entail twice as much
-        computation per memory access, grow only by at most 15\%.
-        The scheduler overheads make up less than 5\% of the total
-        time.
-        }
-    \label{fig:BHTimes}
-\end{figure}
-
-\section{Discussion and Conclusions}
-
-The task scheduler described in the previous sections, QuickSched,
-differs from existing task-based programming schemes
-in a number of ways.
-The most obvious such difference is the addition of {\em conflicts},
-modeled using exclusive locks on hierarchical resources.
-This extension to the standard dependencies-only model
-of task-based parallelism allows for more complex task relations,
-such as in the Barnes-Hut tree-code described earlier.
-
-Another significant difference is that the tasks, their
-dependencies, and their conflicts must be described
-{\em explicitly} before the parallel computation starts.
-This as opposed to implicit dependencies generated
-by task spawning, e.g.~as in Cilk, or to extracting the
-dependencies from the task parameters, e.g.~in QUARK or OmpSs.
-Explicitly defining dependencies has the advantage that 
-more elaborate dependency structures can be generated.
-Furthermore, knowing the structure of the entire task
-graph from the start of the computation provides valuable
-information when scheduling the tasks, e.g.~using the 
-critical path along the dependencies to compute the
-task weight.
-
-Finally, as opposed to the most other task-based
-programming environments which rely on compiler extensions
-and/or code pre-processors, QuickSched operates as a regular
-C-language library, based on standard parallel functionality
-provided by OpenMP and/or {\tt pthreads}.
-This ensures a maximum amount of portability on existing
-and future architectures.
-The interfaces are also kept as simple
-as possible in order to reduce the burden on the programmer
-when implementing task-based codes.
-
-The QuickSched library itself is remarkably simple, consisting of
-less than 3\,000 lines of code, including comments.
-Both examples, which are distributed with QuickSched,
-require less than 1\,000 lines of code each.
-
-In both examples, QuickSched performs extremely well, even
-on a large number of shared-memory cores.
-This performance is due, on the one hand, to the
-division of labor between the scheduler and the queues,
-and on the other hand due to the simple yet efficient
-algorithms for task selection and resource locking.
-The task weighting based on the length of the critical
-path of the dependencies delivers, in the examples shown,
-good parallel efficiency.
-
-There are several possible improvements to QuickSched which
-have not been addressed in this paper.
-The most obvious of which are the following:
-\begin{itemize}
-    \item {\em Priorities}: The current implementation of
-        QuickSched does not take the resource locks into
-        account when selecting tasks in the queues,
-    \item {\em Work-stealing}: During work-stealing, the
-        queues are probed in a random order although
-        the total relative cost of the tasks in the queue,
-        as well as the length of their critical paths are
-        known,
-    \item {\em Costs}: The size of the resources used by
-        a task are currently not taken into account when
-        assigning it to the queues in {\tt qsched\_enqueue},
-        or when approximating the cost of a task.
-\end{itemize}
-
-QuickSched is distributed under the GNU Lesser General Public Licence
-v\,3.0 and is available for download via
-\url{https://sourceforge.net/projects/quicksched/files/}.
-
-% Acknowledgments
-\begin{acks}
-The author would like to thank Sam Townsend for implementing the
-kernels in the task-based QR decomposition as part of his MSc Thesis
-at Durham University.
-The author also thanks Lydia Heck of the Institute for
-Computational Cosmology at Durham University for providing access
-to, and expertise on, the COSMA cluster used in the performance
-evaluation.
-\end{acks}
-
-% Bibliography
-\bibliographystyle{ACM-Reference-Format-Journals}
-\bibliography{quicksched}
-                             % Sample .bib file with references that match those in
-                             % the 'Specifications Document (V1.5)' as well containing
-                             % 'legacy' bibs and bibs with 'alternate codings'.
-                             % Gerry Murray - March 2012
-
-% History dates
-\received{February 2007}{March 2009}{June 2009}
-
 \end{document}
-% End of v2-acmsmall-sample.tex (March 2012) - Gerry Murray, ACM

### 2014-07-31 examples/test_bh_sorted.c
Initialized some counters to 0 in test_bh() to get rid of gcc warnings

@@ -1562,7 +1562,7 @@ void test_bh(int N, int nr_threads, int runs, char *fileName) {
   FILE *file;
   struct qsched s;
   ticks tic, toc_run, tot_setup = 0, tot_run = 0;
-  int countMultipoles, countPairs, countCoMs;
+  int countMultipoles=0, countPairs=0, countCoMs=0;
 
   /* Runner function. */
   void runner(int type, void * data) {

### 2014-09-21 paper/paper.tex
yet another iteration.

@@
 \cortext[cor]{Corresponding author}
-\address[ecs]{School of Engineering and Computing Sciences, Durham University, South Road, Durham DH1 3LE, United Kingdom}
-\address[icc]{Institute for Computational Cosmology, Durham University, South Road, Durham DH1 3LE, United Kingdom}
-\address[ggl]{Google Switzerland GmbH, Brandshenkestr. 110, 8002 Z\"urich, Switzerland}
+\address[ecs]{School of Engineering and Computing Sciences, Durham University, South Road, Durham DH1 3LE, United Kingdom.}
+\address[icc]{Institute for Computational Cosmology, Durham University, South Road, Durham DH1 3LE, United Kingdom.}
+\address[ggl]{Google Switzerland GmbH, Brandshenkestr. 110, 8002 Z\"urich, Switzerland.}
 
@@
 The tasks and their dependencies can be seen as the nodes and edges
-respectively of a Directed Acyclic Graph (DAG) which can be
+of a Directed Acyclic Graph (DAG) which can be
 traversed in topological order, executing the tasks at the nodes
@@
 \cite{ref:Reinders2010}
-provide task-based parallelism using C++ templates.
-Dependencies are handled either by explicitly waiting
+provide task-based parallelism using C++ templates in which
+dependencies are handled either by explicitly waiting
 for spawned tasks, or by explicitly manipulating
@@
 In the following, such a relationship will be referred to
-as a ``conflict'' between two tasks.
+as a {\em conflict} between two tasks.
 \fig{TaskConflicts} shows a task graph with conflicting tasks
@@
 
-Conflicts can be modeled directly as exclusive locks on a shared resource
-which have to be obtained by a task before it can execute.
-Thus, in \fig{TaskConflicts}, before executing, task $F$ has
-to obtain an exclusive lock on the resource associated with
-the conflict between tasks $F$, $H$, and $I$.
-While task $F$ is being executed, neither $H$ nor $I$ can 
-lock the same resource, and therefore will not execute until
-task $F$ is done and the lock has been released.
-
-In order to be of practical use, and to scale well with modern
-multi-core shared-memory architectures, any task scheduler
-implementing constraints must be:
+This paper presents QuickSched, a framework for task-based
+parallel programming with constraints, which aims to achieve
+the following goals:
 \begin{itemize}
@@
 \end{itemize}
-\noindent This paper presents QuickSched, a framework for task-based
-parallel programming with constraints, which aims to achieve
-all of these goals.
+\noindent 
 Section~2 describes the main design considerations, Section~3 the
@@
 \begin{itemize}
-  \item Implicitly via spawning and waiting, e.g.~as in Cilk, or
-  \item Automatic extraction from data dependencies.
+  \item Implicitly via spawning and waiting, e.g.~as is done in Cilk, or
+  \item Automatic extraction from data dependencies, e.g.~as is done in OmpSs.
 \end{itemize}
@@
 single task, i.e.~if the task DAG is a tree, each task {\em spawns}, or
-creates, its dependent tasks after its completion.
+creates, its dependent tasks after its completion (see \fig{Spawn}).
 Hence for the tasks $A$--$E$ in \fig{Tasks}, task $A$ would spawn
@@
 
-Automatic dependency extraction works by enforcing dependencies
+\begin{figure}
+    \centerline{\epsfig{file=figures/Spawn.pdf,width=0.7\textwidth}}
+    \caption{Two different task graphs and how they can be implemented
+      using spawning and waiting.
+      For the task graph on the left, each task spawns its dependent
+      task or tasks. For the task graph on the right, in which task $C$
+      has multiple dependencies, tasks $A$ and $B$ must be spawned and waited
+      for by the calling thread before task $C$ can be spawned.}
+    \label{fig:Spawn}
+\end{figure}
+
+Automatic dependency extraction, on the other hand,
+works by enforcing dependencies
 between tasks that access the same data.
@@
 This approach usually relies on compiler extensions, e.g.~ {\tt pragma}s
-in C/C++, or a system of runtime wrappers, to describe the task parameters
+in C/C++, or a system of function call wrappers, to describe the task parameters
 and their intent.
@@
 described in the previous section.
-
-Here, QuickSched takes a drastically different approach.
-The programmer creates the task DAG explicitly before execution.
+They may also not be able to understand more complex memory access patterns,
+e.g.~for two tasks modifying the upper and lower triangular parts of a matrix,
+which access the same block of memory, but never actually generate any
+concurrency issues.
+
+Here, QuickSched takes a drastically different approach by requiring
+the programmer to create the complete task graph and its dependencies
+explicitly before execution.
+This approach has two main advantages:
+\begin{itemize}
+  \item It gives the user maximum flexibility with regards to the
+    structure of the task graph generated,
+  \item Knowing the structure of the task graph before execution
+    allows the task scheduler to make more informed decisions
+    as to how the tasks should be prioritized.
+\end{itemize}
+
+\noindent The obvious disadvantage is the burden of producing a correct
+task graph is placed on the user.
+Although some problems such as cyclic dependencies can be detected
+automatically, there is no automatic way to detect whether the
+dependencies actualy reflect the intent of the user.
+Due to this added complexity, we consider QuickSched to be
+a tool not designed for casual parallel programmers, but for 
+those interested in investing more programming effort to achieve
+better performance.
+
+Conflicts between tasks or groups of tasks are not specified directly,
+but are instead modeled as exclusive locks on a shared resource
+which have to be obtained by a task before it can execute.
+Thus, in \fig{TaskConflicts}, before executing, task $F$ has
+to obtain an exclusive lock on the resource associated with
+the conflict between tasks $F$, $H$, and $I$.
+While task $F$ is being executed, neither $H$ nor $I$ can 
+lock the same resource, and therefore will not execute until
+task $F$ is done and the lock has been released.
+
+The partitioning of the computation into tasks is also left entirely
+to the user.
+In theory, any program can be automatically converted to a task-based
+representation since each statement in the program code
+can be considered a single task, with dependencies to the
+statements/tasks that produce the input values.
+This set of basic tasks could, again in theory, be reduced by merging
+tasks that share dependencies and/or resources using a variety
+of graph algorithms.
+The decomposition of a computation into tasks, however, usually
+involves re-thinking the underlying algorithms such that they
+best fit the task-based paradigm, e.g.~as in the examples in the
+following sections, or as in \cite{ref:Gonnet2014}.
+This process requires careful evaluation and is probably best
+not left as an automatic transformation of an otherwise serial code.
+
+Finally, the task granularity is an important issue: if the task
+decomposition is too coarse, then good parallelism
+and load-balancing will be difficult to achieve.
+Converseley, if the tasks are too small, the costs of selecting and
+scheduling tasks, which is usually constant per task, will
+quickly destory any performance gains from parallelism.
+Starting from a per-statement set of tasks, it is therefore
+reasonable to group them by their dependencies and shared resources.
+
+In the examples presented herein, we have chosen our task decomposition
+and granularity such that
+\begin{itemize}
+  \item Each task maximizes the ratio of computation to data required,
+  \item The resources required for each task fit comfortably in the
+    lowest-level caches of the underlying system.
+\end{itemize}
+\noindent The first critera is biased towards bigger tasks, while the
+second limits their size.
+The criteria are thus used to optimize the
+cache efficiency of the computation.
 
@@
 
-The division of labor regarding {\em correctness}
-between the scheduler and
+The division of labor between the scheduler and
 the queue objects is illustrated in \fig{QSched}.
@@
 If the tasks in the queue share the same set of resources,
-in increases the probability of said resources already
+it increases the probability of said resources already
 being present in the thread's cache, thus increasing
-{\em memory efficiency}.
+{\em cache efficiency}.
 The scheduler is in charge of selecting the most appropriate
@@
 that unlock this task.
-Given an array of {\tt N} tasks, the wait-counters can
-be set as follows:
-\begin{center}\begin{minipage}{0.9\textwidth}
-    \begin{lstlisting}
-int j, k;
-for ( k = 0 ; k < N ; k++ )
-  tasks[k].wait = 0;
-for ( k = 0 ; k < N ; k++ )
-  for ( j = 0 ; j < tasks[k].nr_unlocks ; j++ )
-    tasks[k].unlocks[j]->wait += 1;
-    \end{lstlisting}
-\end{minipage}\end{center}
+The wait-counters can
+be set by initializing all the task wait counters to zero and then
+incrementing the wait counter of each unlock-task of each task.
 
@@
 relative cost of the critical path following the
-dependencies of this task, respectively.
-The task weights can be computed by first sorting
-the tasks topologically according to their dependencies, i.e.
-as per \cite{ref:Kahn1962}:
-\begin{center}\begin{minipage}{0.9\textwidth}
-    \begin{lstlisting}
-int top[N], j, k, sorted = 0, maxweight;
-set the task waits as described above.
-for ( k = 0 ; k < N ; k++ )
-  if ( tasks[k].wait == 0 )
-    top[ sorted++ ] = k;
-for ( k = 0 ; k < sorted ; k++ )
-  for ( j = 0 ; j < tasks[ top[k] ].nr_unlocks ; j++ )
-    if ( ( tasks[ top[k] ].unlock[j]->wait -= 1 ) == 0 )
-      top[ sorted++ ] = k;
-if ( k < N )
-  error( "Circular dependencies detected." );
-    \end{lstlisting}
-\end{minipage}\end{center}
-\noindent where the array {\tt top} contains the task indices
-in topological order. 
-The test in line~10 is a convenient check if the tasks and their
-dependencies actually do form an acyclic graph.
-The weights themselves are then computed as follows
-\begin{center}\begin{minipage}{0.9\textwidth}
-    \begin{lstlisting}
-for ( k = N-1 ; k >= 0 ; k-- ) {
-  for ( w = 0 , j = 0 ; j < tasks[ top[k] ].nr_unlocks ; j++ ) {
-    tasks[ top[k] ].unlock[j]->wait += 1;
-    w = max( tasks[ top[k] ].unlock[j]->weight , w );
-    }
-  tasks[ top[k] ].weight = tasks[ top[k] ].cost + w;
-  }
-    \end{lstlisting}
-\end{minipage}\end{center}
-\noindent where the tasks are traversed in reverse
-topological order, computing the recursive weight as the sum of the
-task cost and the maximum weight of the tasks it unlocks (line~6),
-and recomputing the task waits at the same time (line~3).
+dependencies of this task, respectively, i.e.~the task's cost
+plus the maximum dependent task's weight (see \fig{TaskWeight}).
+The task weights are computed by traversing
+the tasks in reverse topological order following their dependencies, i.e.
+as per \cite{ref:Kahn1962}, and computing each task's weight
+up the task DAG.
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/TaskWeight.pdf,height=0.3\textwidth}}
+    \caption{Computation of the task weight.
+      In this task graph, the height of each task corresponds to its
+      computational {\em cost}.
+      The {\em weight} of each task, e.g.~the topmost task,
+      is computed from its cost plus
+      the maximum dependent task's weight.
+      This corresponds to the computatoinal cost of the critical
+      path following the task's dependencies.}
+    \label{fig:TaskWeight}
+\end{figure}
 
@@
 whether this resource is currently in use, i.e.~{\em locked}.
-To avoid race conditions, this value should only be tested
-and set using atomic instructions.
+To avoid race conditions, this value should be tested
+and set using atomic instructions only.
 The {\tt hold} field is a counter indicating how many
@@
     \begin{lstlisting}
-void resource_hold ( struct resource *r ) {
-  if ( atomic_cas( &r->lock , 0 , 1 ) != 0 )
-    return 0;
-  atomic_inc( &r->hold );
+void resource_hold(struct resource *r) {
+  if (atomic_cas(&r->lock, 0, 1) != 0) return 0;
+  atomic_inc(&r->hold);
   r->lock = 0;
@@
 If the resource can be locked, the hold counter is incremented
-atomically (line~4), and the lock is released (line~5),
+atomically (line~3), and the lock is released (line~4),
 returning {\tt 1} or {\tt 0} if the resource could be held
@@
     \begin{lstlisting}
-void resource_lock ( struct resource *r ) {
+void resource_lock(struct resource *r) {
   struct resource *up, *top;
-  if ( r->hold && atomic_cas( &r->lock , 0 , 1 ) != 0 )
+  if (r->hold && atomic_cas(&r->lock, 0, 1) != 0)
     return 0;
-  if ( r->hold ) {
+  if (r->hold) {
     r->lock = 0;
@@
   }
-  for ( up = r->parent ; up != NULL ; up = up->parent ) {
-    if ( !resource_hold( up ) )
-      break;
-  if ( ( top = up ) != NULL ) {
-    for ( up = r->parent ; up != top , up = up->parent )
-      atomic_dec( &up->hold );
+  for (up = r->parent; up != NULL; up = up->parent)
+    if (!resource_hold(up)) break;
+  if ((top = up) != NULL) {
+    for (up = r->parent; up != top, up = up->parent)
+      atomic_dec(&up->hold);
     r->lock = 0;
     return 0;
-    }
-  else
+  } else
     return 1;
@@
 Due to a possible race condition when holding the resource,
-the {\tt hold} counter must checked again once the resource
+the {\tt hold} counter must be checked again once the resource
 has been locked (line~5).
-In lines~9--11 the hold counters of the hierarchical parents
+In lines~9--10 the hold counters of the hierarchical parents
 are incremented using the procedure described earlier.
-If this process fails at any point (line~12), the
-previously set hold counters are decremented (lines~13--14)
-and the lock is released (line~15).
+If this process fails at any point (line~11), the
+previously set hold counters are decremented (line~12)
+and the lock is released (line~13).
 The procedure then returns {\tt 1} or {\tt 0} if the resource
@@
 
-Finally, unlocking a resource is relatively straight-forward:
-\begin{center}\begin{minipage}{0.9\textwidth}
-    \begin{lstlisting}
-void resource_unlock ( struct resource *r ) {
-  struct resource *up;
-  r->lock = 0;
-  for ( up = r->parent ; up != NULL ; up = up->parent ) {
-    atomic_dec( &up->hold );
-  }
-    \end{lstlisting}
-\end{minipage}\end{center}
-\noindent where the resource itself is unlocked (line~3)
-and the hold counters of all of its hierarchical parents
-are decremented (lines~4--5).
+Finally, unlocking a resource is relatively straight-forward,
+i.e.~the lock is set to zero and the hold counters up the
+resource hierarchy are decremented.
 
@@
 Although this would return the best possible task, it
-requires maintaining a sorted list in which inserting
+requires maintaining a sorted array in which inserting
 or removing an entry is in \oh{n} for $n$ elements.
@@
 
-The data structure for the queue is thus defined as follows:
+The data structure for the queue is defined as follows:
 \begin{center}\begin{minipage}{0.9\textwidth}
@@
     \begin{lstlisting}
-void queue_put ( struct queue *q , struct task *t ) {
-  while ( atomic_cas( q->lock , 0 , 1 ) != 0 );
-  q->tasks[ q->count++ ] = t;
-  bubble-up the q->count-1st entry of q->tasks.
+void queue_put(struct queue *q, struct task *t) {
+  while (atomic_cas(q->lock, 0, 1) != 0) {}
+  q->tasks[q->count++] = t;
+  bubble-up the q->count - 1st entry of q->tasks.
   q->lock = 0;
@@
     \begin{lstlisting}
-struct task *queue_get ( struct queue *q ) {
+struct task *queue_get(struct queue *q) {
   struct task *res = NULL;
   int j, k;
-  while ( atomic_cas( q->lock , 0 , 1 ) != 0 );
-  for ( k = 0 ; k < q->count ; k++ ) {
-    for ( j = 0 ; j < q->tasks[k]->nr_locks ; j++ )
-      if ( !resource_lock( q->tasks[k]->lock[j] ) )
-        break;
-    if ( j < q->tasks[k]->nr_locks )
-      for ( j = j-1 ; j >= 0 ; j-- )
-        resource_unlock( q->tasks[k]->lock[j] );
+  while (atomic_cas(q->lock, 0, 1) != 0) {}
+  for (k = 0; k < q->count; k++) {
+    for (j = 0; j < q->tasks[k]->nr_locks; j++)
+      if (!resource_lock(q->tasks[k]->lock[j])) break;
+    if (j < q->tasks[k]->nr_locks)
+      for (j = j - 1; j >= 0; j--)
+        resource_unlock(q->tasks[k]->lock[j]);
     else
@@
   }
-  if ( k < q->count ) {
+  if (k < q->count) {
     res = q->tasks[k];
-    q->tasks[k] = q->tasks[ --q->count ];
+    q->tasks[k] = q->tasks[--q->count];
     trickle-down the kth entry of q->tasks.
@@
 locked for exclusive access (line~4).
-The array of task pointers is then traversed in heap order (line~5),
-locking the resources of each task (lines~6--8).
-If any of these locks fail (line~9), the locks that were obtained
-are released (lines~10--11), otherwise, the traversal is aborted
-(line~13).
-If all the locks on a task could be obtained (line~15), the
-task pointer is replaced by the last pointer in the heap (line~17)
-and the heap order is restored (line~18).
+The array of task pointers is then traversed (line~5),
+locking the resources of each task (lines~6--7).
+If any of these locks fail (line~8), the locks that were obtained
+are released (lines~9--10), otherwise, the traversal is aborted
+(line~12).
+If all the locks on a task could be obtained (line~14), the
+task pointer is replaced by the last pointer in the heap (line~16)
+and the heap order is restored (line~17).
 Finally, the queue lock is released (line~19) and the locked task
@@
     \begin{lstlisting}
-void qsched_run ( qsched *s , void (*fun)( int , void * ) ) {
-  qsched_start( s );
+void qsched_run(qsched *s, void (*fun)(int, void *)) {
+  qsched_start(s);
   #pragma omp parallel
@@
     struct task *t;
-    while ( ( t = qsched_gettask( s , qid ) ) != NULL ) {
-      fun( t->type , t->data );
-      qsched_done( s , t );
+    while ((t = qsched_gettask(s, qid)) != NULL) {
+      fun(t->type, t->data);
+      qsched_done(s, t);
     }
@@
 concurrently.
+A version using {\tt pthreads} \cite{ref:Pthreads1995}
+directly is also available.
 The parallel section consists of a loop (lines~7--10) in
@@
 identifies the tasks that have no dependencies and sends them
-to queues:
-\begin{center}\begin{minipage}{0.9\textwidth}
-    \begin{lstlisting}
-void qsched_start ( qsched *s ) {
-  sort out pointers in tasks.
-  set the task waits and weights.
-  s->waiting = s->nr_tasks;
-  for ( k = 0 ; k < s->nr_tasks && s->tasks[tid[k]].wait == 0 ; k++ )
-    qsched_enqueue( s , &s->tasks[tid[k]] );
-  }
-    \end{lstlisting}
-\end{minipage}\end{center}
-\noindent where line~2 sets the {\tt unlocks}, {\tt locks},
-and {\tt uses} pointers in the tasks.
-The operations in line~3 are described in \sect{tasks}.
-The function {\tt qsched\_enqueue} tries to identify the best
+to queues via the function {\tt qsched\_enqueue} which
+tries to identify the best
 queue for a given task by looking at which queues last used
@@
     \begin{lstlisting}
-void qsched_enqueue ( qsched *s , struct task *t ) {
-  int k, qid, score[ s->nr_queues ];
-  for ( k = 0 ; k < s->nr_queues ; k++ )
+void qsched_enqueue(qsched *s, struct task *t) {
+  int k, qid, score[s->nr_queues];
+  for (k = 0; k < s->nr_queues; k++)
     score[k] = 0;
-  for ( j = 0 ; j < t->nr_locks ; j++ )
-    score[ t->locks[j]->owner ] += 1;
-  for ( j = 0 ; j < t->nr_uses ; j++ )
-    score[ t->uses[j]->owner ] += 1;
-  for ( qid = 0 , k = 1 ; k < s->nr_queues ; k++ )
-    if ( score[k] > score[qid] )
+  for (j = 0; j < t->nr_locks; j++)
+    score[t->locks[j]->owner] += 1;
+  for (j = 0; j < t->nr_uses; j++)
+    score[t->uses[j]->owner] += 1;
+  for (qid = 0, k = 1; k < s->nr_queues; k++)
+    if (score[k] > score[qid])
       qid = k;
-  queue_put( &s->queues[qid] , t );
+  queue_put(&s->queues[qid], t);
 }
@@
     \begin{lstlisting}
-struct task *qsched_gettask ( qsched *s , int qid ) {
+struct task *qsched_gettask(qsched *s, int qid) {
   struct task *res = NULL;
   int k;
-  while ( s->waiting ) {
-    if ( ( res = queue_get( s->queues[qid] ) ) == NULL ) {
+  while (s->waiting) {
+    if ((res = queue_get(s->queues[qid])) == NULL) {
       loop over all other queues in random order with index k
-        if ( ( res = queue_get( s->queues[k] ) ) != NULL )
+        if ((res = queue_get(s->queues[k])) != NULL)
           break;
-      }
-    else
+    } else
       break;
   }
-  if ( res != NULL && s->reown ) {
-    for ( k = 0 ; k < res->nr_locks ; k++ )
+  if (res != NULL && s->reown) {
+    for (k = 0; k < res->nr_locks; k++)
       res->locks[k]->owner = qid;
-    for ( k = 0 ; k < res->nr_uses ; k++ )
+    for (k = 0; k < res->nr_uses; k++)
       res->uses[k]->owner = qid;
@@
 If the queue is empty, or all of the tasks in that queue had
-unresolved conflicts, the scheduler uses {\em work stealing}
-\cite{ref:Blumofe1999}, i.e.~it loops over all other queues
+unresolved conflicts, the scheduler uses a variant of the
+{\em work stealing} described in \cite{ref:Blumofe1999},
+i.e.~it loops over all other queues
 in a random order (line~6) and tries to get a task from them
@@
 If a task could be obtained from any queue and task re-owning
-is switched on (line~13),
+is switched on (line~12),
 the resources it locks and uses are marked as now being owned
-by the preferred queue (lines~14--17).
+by the preferred queue (lines~13--16).
 Finally, the task, or {\tt NULL} if no task could be obtained,
@@
 to unlock the resources and tasks which depend on it.
-This is handled by the function {\tt qsched\_done}:
-\begin{center}\begin{minipage}{0.9\textwidth}
-    \begin{lstlisting}
-void qsched_done ( qsched *s , struct task *t ) {
-  int k;
-  for ( k = 0 ; k < t->nr_locks ; k++ )
-    resource_unlock( t->locks[k] );
-  for ( k = 0 ; k < t->nr_unlocks ; k++ )
-    if ( atomic_dec( &t->unlocks[k]->wait ) == 1 )
-      qsched_enqueue( s , t->unlocks[k] );
-  atomic_dec( &s->waiting );
-  }
-    \end{lstlisting}
-\end{minipage}\end{center}
-\noindent If any of the unlocked tasks' wait counters
-goes to zero (line~6), then the unlocked task is ready to
-run and is immediately dispatched via {\tt qsched\_enqueue},
-as described earlier.
+This is handled by the function {\tt qsched\_done}, which
+calls {\tt qsched\_enqueue} on any tasks for which the
+wait counter is decremented to zero.
 Once all the dependent tasks have been unlocked, the
-{\tt waiting} counter is decremented (line~8).
+{\tt waiting} counter of the scheduler is decremented.
 
@@
 providing benchmarks to assess its efficiency and scalability.
-
 The first test is the tiled QR decomposition originally
@@
 programming infrastructures.
-
 The second example is a Barnes-Hut tree-code, a problem
@@
 This example shows how conflicts, modeled
-via hierarchical resources, can be useful.
+via hierarchical resources, can be useful in modelling and executing
+a problem efficiently.
 
@@
 and generate the plots used in the following.
-All examples were compiled with gcc v.\,4.8.1 using the
-{\tt -O2 -march=native} flags and run on a 
-tests were run on a 64-core AMD Opteron 6376 machine running
-at 2.6\,GHz.
+All examples were compiled with gcc v.\,4.8.2 using the
+{\tt -O2 -march=native} flags and run on
+a 64-core AMD Opteron 6376 machine running
+at 2.67\,GHz.
 
@@
 The former uses the QUARK task scheduler, which was originally
-designed for it, while the latter currently uses the StarPU
-task scheduler.
+designed for this specific task, while the latter currently uses
+the StarPU task scheduler \cite{ref:Agullo2011}.
 
@@
 
-\noindent where the task names are the BLAS-like operation
+\noindent and where the task names are the BLAS-like operation
 performed on the given tiles.
@@
 The QR decomposition was computed for a $2048\times 2048$
-random matrix using tiles of size $64\times 64$ using QuickSched
+random matrix using tiles of size $64\times 64$ floats using QuickSched
 as described above.
-For the DGEQRF and DLARFT kernels, the respective routines
-from the LAPACKE\footnote{\url{http://www.netlib.org/lapack/lapacke.html}}
-library were used.
-The DTSQRF and DSSRFT kernels were implemented by Sam Townsend
-as part of his MSc thesis in Computer Science at Durham University.
 For this matrix, a total of 11440 tasks with 32240 dependencies
@@
 
-For these tests, OpenMP parallelism and resource re-owning
+For these tests, {\tt pthread} parallelism and resource re-owning
 were used with one queue per core.
@@
 \noindent The scaling and efficiency relative to QuickSched are 
-shown in \fig{QRResults} as well.
+shown in \fig{QRResults}.
 The difference in timings is the result of the different
@@
 which shows the assignment of the different tasks to cores for the
-64 core run.
+64-core run.
 The most visible difference between both schedulers is that
@@
 per cell is smaller than some limit $n_\mathsf{max}$.
-
-The particle interactions can also be formulated recursively:
+The particle interactions can then be formulated recursively:
 Given a particle and a set of particles in a cell,
@@
 
-The task-based implementation will consist of four
+The task-based implementation will consist of three
 types of tasks:
@@
         all particles in one cell are interacted with the
-        center of mass of another cell, and
-    \item {\em Center of mass} tasks, which compute
-        the center of mass of a single cell either from
-        the sum of the centers of mass of its sub-cells
-        if it has been split, or directly from the
-        particles otherwise.
+        center of mass of all other cells in the tree.
 \end{itemize}
@@
 The library can be compiled to use OpenMP and/or the
-{\tt pthreads} library \cite{ref:Pthreads1995}.
+{\tt pthreads} library.
 OpenMP is the default, but calling {\tt qsched\_init} with

### 2014-09-23 examples/plot_sorted.py
The sorted interactions now use 4 multipoles ( 1 for each quadrant for increased accuracy ). This only works along the X-axis.

@@ -56,9 +56,9 @@ errx_s = (accx_s - accx_u )/abs(accx_u)
 erry_s = (accy_s - accy_u )/abs(accy_u) 
 errz_s = (accz_s - accz_u )/abs(accz_u) 
 
-e_errx_s = errx_s[abs(errx_s) > 0.001]
-e_erry_s = erry_s[abs(erry_s) > 0.001]
-e_errz_s = errz_s[abs(errz_s) > 0.001]
+e_errx_s = errx_s#[abs(errx_s) > 0.001]
+e_erry_s = erry_s#[abs(erry_s) > 0.001]
+e_errz_s = errz_s#[abs(errz_s) > 0.001]
 
 # Statistics
 meanx_s = mean(errx_s[abs(errx_s) < 0.1])
@@ -74,14 +74,16 @@ stdz_s = std(errz_s[abs(errz_s) < 0.1])
 figure(frameon=True)
 
 subplot(311, title="Acceleration along X")
-plot(id[abs(errx_s) > 0.001], e_errx_s , 'rx')
+#plot(id[abs(errx_s) > 0.001], e_errx_s , 'rx')
+plot(id, e_errx_s , 'rx')
 #text(id[-1], 0.18, "B-H: $%5.3f\\pm%5.3f$\n QuickShed: $%5.3f\\pm%5.3f$"%(meanx_bh, stdx_bh, meanx_new, stdx_new), backgroundcolor="w", va="top", ha="right" )
 ylim(-0.2, 0.2)
 xlim(0, size(id)-1)
 grid()
 
 subplot(312, title="Acceleration along Y")
-plot(id[abs(erry_s) > 0.001], e_erry_s , 'rx')
+#plot(id[abs(erry_s) > 0.001], e_erry_s , 'rx')
+plot(id, e_erry_s , 'rx')
 #text(id[-1], 0.18, "B-H: $%5.3f\\pm%5.3f$\n QuickShed: $%5.3f\\pm%5.3f$"%(meany_bh, stdy_bh, meany_new, stdy_new), backgroundcolor="w", va="top", ha="right" )
 ylim(-0.2, 0.2)
 xlim(0, size(id)-1)
@@ -89,7 +91,8 @@ xlim(0, size(id)-1)
 grid()
 
 subplot(313, title="Acceleration along Z")
-plot(id[abs(errz_s) > 0.001], e_errz_s , 'rx', label="Sorted")
[…]

### 2014-09-23 examples/test_bh_sorted.c

@@ -39,15 +39,16 @@
 #define cell_pool_grow 1000
 #define cell_maxparts 100
 #define task_limit 1e8
-#define const_G 6.6738e-8
+#define const_G 1//6.6738e-8
 #define dist_min 0.5 /* Used fpr legacy walk only */
-#define dist_cutoff_ratio 1.75
+#define dist_cutoff_ratio 1.5
 #define iact_pair_direct iact_pair_direct_sorted
 
-#define ICHECK -1
+#define ICHECK 2
 #define NO_SANITY_CHECKS
 #define NO_COM_AS_TASK
 #define COUNTERS
+#define NO_MANY_MULTIPOLES
 
 /** Data structure for the particles. */
 struct part {
@@ -1025,7 +1026,7 @@ static inline void iact_pair_direct_unsorted(struct cell *ci, struct cell *cj) {
       ++count_direct_unsorted;
 #endif
 
-#if ICHECK >= 0
+#if ICHECK >= 0 && 0
       if (parts_i[i].id == ICHECK)
         printf("[NEW] Interaction with particle id= %d (pair i) a=( %e %e %e )\n", parts_j[j].id, -mi*w*dx[0], -mi*w*dx[1], -mi*w*dx[2]);
 
@@ -1050,7 +1051,7 @@ static inline void iact_pair_direct_unsorted(struct cell *ci, struct cell *cj) {
  */
 static inline void iact_pair_direct_sorted(struct cell *ci, struct cell *cj) {
 
-  int i, j, k;
+  int i, j, k, l;
   int count_i, count_j;
   struct part *parts_i, *parts_j;
   double cjh = cj->h;
@@ -1068,8 +1069,8 @@ static inline void iact_pair_direct_sorted(struct cell *ci, struct cell *cj) {
   /* Get the sorted indices and stuff. */
   struct index *ind_i, *ind_j;
   float corr;
-  double com[3] = {0.0, 0.0, 0.0};
-  float com_mass = 0.0;
+  double com[4][3] = {{0.0, 0.0, 0.0}, {0.0, 0.0, 0.0}, {0.0, 0.0, 0.0}, {0.0, 0.0, 0.0}};
+  float com_mass[4] = {0.0, 0.0, 0.0, 0.0};
   get_axis(&ci, &cj, &ind_i, &ind_j, &corr);
   count_i = ci->count;
   parts_i = ci->parts;
[…]

### 2014-09-27 examples/plot_sorted.py
Now plotting the x coordinate of particles rather than their ID on the x-axis.

@@ -39,6 +39,8 @@ print "Plotting..."
 # Read Quickshed accelerations
 data=loadtxt("interaction_dump.dat")
 id = data[:,0]
+posx = data[:,2]
+
 accx_u=data[:,5]
 accy_u=data[:,6]
 accz_u=data[:,7]
@@ -75,29 +77,29 @@ figure(frameon=True)
 
 subplot(311, title="Acceleration along X")
 #plot(id[abs(errx_s) > 0.001], e_errx_s , 'rx')
-plot(id, e_errx_s , 'rx')
+plot(posx, e_errx_s , 'rx')
 #text(id[-1], 0.18, "B-H: $%5.3f\\pm%5.3f$\n QuickShed: $%5.3f\\pm%5.3f$"%(meanx_bh, stdx_bh, meanx_new, stdx_new), backgroundcolor="w", va="top", ha="right" )
 ylim(-0.2, 0.2)
-xlim(0, size(id)-1)
+#xlim(0, size(id)-1)
 grid()
 
 subplot(312, title="Acceleration along Y")
 #plot(id[abs(erry_s) > 0.001], e_erry_s , 'rx')
-plot(id, e_erry_s , 'rx')
+plot(posx, e_erry_s , 'rx')
 #text(id[-1], 0.18, "B-H: $%5.3f\\pm%5.3f$\n QuickShed: $%5.3f\\pm%5.3f$"%(meany_bh, stdy_bh, meany_new, stdy_new), backgroundcolor="w", va="top", ha="right" )
 ylim(-0.2, 0.2)
-xlim(0, size(id)-1)
+#xlim(0, size(id)-1)
 
 grid()
 
 subplot(313, title="Acceleration along Z")
 #plot(id[abs(errz_s) > 0.001], e_errz_s , 'rx', label="Sorted")
-plot(id, e_errz_s , 'rx', label="Sorted")
+plot(posx, e_errz_s , 'rx', label="Sorted")
 #text(id[-1], 0.18, "B-H: $%5.3f\\pm%5.3f$\n QuickShed: $%5.3f\\pm%5.3f$"%(meanz_bh, stdz_bh, meanz_new, stdz_new), backgroundcolor="w", va="top", ha="right" )
 legend(loc="upper right")
 
 ylim(-0.2, 0.2)
-xlim(0, size(id)-1)
+#xlim(0, size(id)-1)
 grid()
 
 savefig("accelerations.png")

### 2014-09-28 examples/plot_sorted.py
The sorted interactions test case can now be run for each of the three possible orientations.

@@ -36,113 +36,117 @@ import os
 
 print "Plotting..."
 
-# Read Quickshed accelerations
-data=loadtxt("interaction_dump.dat")
-id = data[:,0]
-posx = data[:,2]
+names = ["side", "edge", "corner"]
 
-accx_u=data[:,5]
-accy_u=data[:,6]
-accz_u=data[:,7]
+for orientation in range( 3 ):
 
-accx_s=data[:,8]
-accy_s=data[:,9]
-accz_s=data[:,10]
-
-
-
-
-# Build error ------------------------------------------------
-
-errx_s = (accx_s - accx_u )/abs(accx_u) 
-erry_s = (accy_s - accy_u )/abs(accy_u) 
-errz_s = (accz_s - accz_u )/abs(accz_u) 
-
-e_errx_s = errx_s#[abs(errx_s) > 0.001]
-e_erry_s = erry_s#[abs(erry_s) > 0.001]
-e_errz_s = errz_s#[abs(errz_s) > 0.001]
-
-# Statistics
-meanx_s = mean(errx_s[abs(errx_s) < 0.1])
-stdx_s = std(errx_s[abs(errx_s) < 0.1])
-meany_s = mean(erry_s[abs(erry_s) < 0.1])
-stdy_s = std(erry_s[abs(erry_s) < 0.1])
-meanz_s = mean(errz_s[abs(errz_s) < 0.1])
-stdz_s = std(errz_s[abs(errz_s) < 0.1])
+    # Read Quickshed accelerations
+    data=loadtxt( "interaction_dump_%d.dat"%orientation )
+    id = data[:,0]
+    pos = data[:,2]
+    
+    accx_u=data[:,6]
+    accy_u=data[:,7]
+    accz_u=data[:,8]
+    
+    accx_s=data[:,9]
+    accy_s=data[:,10]
+    accz_s=data[:,11]
+    
 
 
 
-# Plot -------------------------------------------------------
-figure(frameon=True)
-
-subplot(311, title="Acceleration along X")
-#plot(id[abs(errx_s) > 0.001], e_errx_s , 'rx')
-plot(posx, e_errx_s , 'rx')
-#text(id[-1], 0.18, "B-H: $%5.3f\\pm%5.3f$\n QuickShed: $%5.3f\\pm%5.3f$"%(meanx_bh, stdx_bh, meanx_new, stdx_new), backgroundcolor="w", va="top", ha="right" )
-ylim(-0.2, 0.2)
[…]

### 2014-09-28 examples/test_bh_sorted.c

@@ -44,7 +44,7 @@
 #define dist_cutoff_ratio 1.5
 #define iact_pair_direct iact_pair_direct_sorted
 
-#define ICHECK 2
+#define ICHECK -1
 #define NO_SANITY_CHECKS
 #define NO_COM_AS_TASK
 #define COUNTERS
@@ -2018,8 +2018,10 @@ void test_bh(int N, int nr_threads, int runs, char *fileName) {
 /**
  * @brief Creates two neighbouring cells wiht N_parts per celland makes them interact using both the 
  * sorted and unsortde interactions. Outputs then the tow sets of accelerations for accuracy tests.
+ * @param N_parts Number of particles in each cell
+ * @param orientation Orientation of the cells ( 0 == side, 1 == edge, 2 == corner )
  */
-void test_direct_neighbour( int N_parts ) {
+void test_direct_neighbour( int N_parts , int orientation ) {
 
   int k;
   struct part *parts;
@@ -2032,19 +2034,34 @@ void test_direct_neighbour( int N_parts ) {
   /* Create random set of particles in both cells */
   for (k = 0; k < 2*N_parts; k++) {
       parts[k].id = k;
-      parts[k].x[0] = ((double)rand()) / RAND_MAX + (k >= N_parts ? 1 : 0. );
+
+      parts[k].x[0] = ((double)rand()) / RAND_MAX;
+      if ( k >= N_parts )
+	parts[k].x[0] += 1.;
+
       parts[k].x[1] = ((double)rand()) / RAND_MAX;
+      if ( orientation > 0 && k >= N_parts )
+      	parts[k].x[1] += 1.;
+
       parts[k].x[2] = ((double)rand()) / RAND_MAX;
-      parts[k].mass = 1.;//((double)rand()) / RAND_MAX;
+      if ( orientation > 1 && k >= N_parts )
+      	parts[k].x[2] += 1.;
+
+      parts[k].mass = ((double)rand()) / RAND_MAX;
       parts[k].a[0] = 0.0;
       parts[k].a[1] = 0.0;
       parts[k].a[2] = 0.0;
     }
 
   /* Get the cell geometry right */
-  left.loc[0] = 0.; left.loc[1] = 0.; left.loc[2] = 0.;
+  left.loc[0] = 0.; 
+  left.loc[1] = 0.; 
+  left.loc[2] = 0.;
   left.h = 1.;
[…]

### 2014-09-28 examples/test_bh_sorted.c
modified get_axis to return two orthogonal vectors to the axis, use them to compute the binning for the multipoles in the sorted interactions. tested using matthieu's unit test and seems to work.

@@ -39,7 +39,7 @@
 #define cell_pool_grow 1000
 #define cell_maxparts 100
 #define task_limit 1e8
-#define const_G 1//6.6738e-8
+#define const_G 1    // 6.6738e-8
 #define dist_min 0.5 /* Used fpr legacy walk only */
 #define dist_cutoff_ratio 1.5
 #define iact_pair_direct iact_pair_direct_sorted
@@ -48,7 +48,7 @@
 #define NO_SANITY_CHECKS
 #define NO_COM_AS_TASK
 #define COUNTERS
-#define NO_MANY_MULTIPOLES
+#define MANY_MULTIPOLES
 
 /** Data structure for the particles. */
 struct part {
@@ -60,7 +60,7 @@ struct part {
   // };
   float mass;
   int id;
-}; // __attribute__((aligned(64)));
+};  // __attribute__((aligned(64)));
 
 /** Data structure for the sorted particle positions. */
 struct index {
@@ -71,7 +71,7 @@ struct index {
 struct multipole {
   double com[3];
   float mass;
-}; 
+};
 
 /** Data structure for the BH tree cell. */
 struct cell {
@@ -105,7 +105,7 @@ struct cell {
 enum task_type {
   task_type_self = 0,
   task_type_pair,
-  task_type_self_pc, 
+  task_type_self_pc,
   task_type_pair_direct,
   task_type_com,
   task_type_count
@@ -135,6 +135,39 @@ const float axis_shift[13 * 3] = {
   5.773502691896258e-01, -5.773502691896258e-01, -5.773502691896258e-01, 0.0,
   7.071067811865475e-01, 7.071067811865475e-01, 0.0, 1.0, 0.0, 0.0,
   7.071067811865475e-01, -7.071067811865475e-01, 0.0, 0.0, 1.0,
+};
+const float axis_orth_first[13 * 3] = {
+  7.071067811865475e-01, -7.071067811865475e-01, 0.0, 0.0, 0.0, 1.0,
+  7.071067811865475e-01, -7.071067811865475e-01, 0.0, 0, 1.0, 0, 0.0, 1.0, 0.0,
+  0.0, 1.0, 0, 7.071067811865475e-01, 0.0, -7.071067811865475e-01, 0, 0.0, 1.0,
+  7.071067811865475e-01, 7.071067811865475e-01, 0.0, 1.0, 0, 0, 1.0, 0.0, 0.0,
[…]

### 2014-09-29 paper/paper.tex
fixed description of BH solver, need to add correct numbers and figures.

@@
     \item If called with a single unsplit cell (line~13),
-        create a self-interaction task on that cell (line~14),
-    \item If called with two cells that are sufficiently well
-        separated (line~21), create two particle-cell pair
-        interactions (lines~23 and~28) over both cells in
-        opposite orders, which depend on the center of mass
-        task of each cell,
-    \item If called with two cells that are not well
-        separated and both cells are split (line~33),
+        create a self-interaction task as well as a particle-cell
+        task on that cell (line~14),
+    \item If called with two non-neighbouring cells (line~21),
+        do nothing, as these interactions
+        will be computed by the particle-cell task,
+    \item If called with two neighbouring cells and both cells
+        are split (line~33),
         recurse over all pairs of sub-cells spanning
         both cells (line~37), and
-    \item If called with two cells that are not well separated
-        and either of the cells are not split, create
+    \item If called with two neighbouring cells
+        and one of the cells are not split, create
         a particle-particle pair task over both cells.
@@
 the cells on which it operates (lines~16, 25, 30, and 42--43).
-
 In order to prevent generating
@@
 of threads each (lines~7 and~34).
-The tasks themselves are then left to recurse over the sub-trees,
-which is why in these cases, the tasks are made to depend on the
-center of mass tasks (lines~17--18 and~41--47)
-which may be used in the ensuing interactions.
+
+The particle-particle pair interaction tasks are implemented
+by computing the interactions between all particle pairs spanning
+both cells in a double for-loop.
+The particle-cell interactions for each leaf node are computed by
+traversing the tree recursively starting from the root node:
+\begin{itemize}
+  \item If called with a node that is a hierarchical parent of
+    the leaf node, or with a node that is a direct neighbour of
+    a hierarchical parent of the leaf node, recurse over the
+    node's sub-cells,
+  \item Otherwise, compute the interaction between the leaf node's
+    center of mass and all the particles in the leaf node.
+\end{itemize}
+
+This task decomposition differs from the traditional tree-walk
+in the Barnes-Hut algorithm in that the particle-cell interactions
+are grouped per leaf, with each leaf doing its own tree walk.
+This approach was chosen to maximize the memory locality
+of the particle-cell calculation, as the particles in the leaf,
+which are traversed for each particle-cell interaction, are
+contiguous in memory, and are thus more likely to remain in the
+lowest-level cache.
 
 This Barnes-Hut tree-code was used to approximate the gravitational
-N-Body problem for 1\,000\,000 particles with random coordinates
+N-Body problem for 1\,000\,000 particles with uniformly random coordinates
 in $[0,1]^3$.
 The parameters $n_\mathsf{max}=100$ and $n_\mathsf{task}=5000$
-were used to generate the tasks, and cell pairs were considered
-well separated if not directly adjacent.
+were used to generate the tasks.
 Using the above scheme generated 97\,553 tasks, of which
 512 self-interaction tasks, ??? particle-particle interaction
-task, ??? particle-cell interaction tasks, and 37\,449
-center of mass tasks.
+task, and ??? particle-cell interaction tasks.
 A total of 141\,840 dependencies were generated, along with
@@
 
-For these tests, OpenMP parallelism was used and resource
+For these tests, {\tt pthread}s parallelism was used and resource
 re-owning was switched off.
@@
     \begin{lstlisting}[basicstyle=\scriptsize\tt]
-void comp_com(struct cell *c) {
-  int j, k;
-  c->com[0] = 0.0; c->com[1] = 0.0; c->com[2] = 0.0;
-  c->mass = 0.0;
+void comp_self(struct cell *c) {
   if (c->split)
-    for (k = 0; k < 8; k++) {
-      struct cell *cp = c->progeny[k];
-      for (j = 0; j < 3; j++)
-        c->com[j] += cp->com[j] * cp->mass;
-      c->mass += cp->mass;
+    for (int j = 0; j < 8; j++) {
+      comp_self(c->progeny[j]);
+      for (int k = j + 1; k < 8; k++)
+        comp_pair(c->progeny[j], c->progeny[k]);
     }
   else
-    for (k = 0; k < 8; k++) {
-      struct part *p = &c->parts[k];
-      for (j = 0; j < 3; j++)
-        c->com[j] += p->x[j] * p->mass;
-      c->mass += p->mass;
-    }
-  c->com[0] /= c->mass; c->com[1] /= c->mass; c->com[2] /= c->mass;
-}
-
-void comp_self(struct cell *c) {
-  int j, k;
-  if (c->split)
-    for (j = 0; j < 8; j++) {
-      comp_self(c->progeny[j]);
-      for (k = j + 1; k < 8; k++)
-        comp_pair(c->progeny[j], c->progeny[k]);
-    }
-  else
-    for (j = 0; j < c->count; j++)
-      for (k = j + 1; k < c->count; k++)
+    for (int j = 0; j < c->count; j++)
+      for (int k = j + 1; k < c->count; k++)
         interact c->parts[j] and c->parts[k].
@@
 void comp_pair(struct cell *ci, struct cell *cj) {
-  int j, k;
-  if (ci and cj well separated) {
-    comp_pair_pc(ci, cj);
-    comp_pair_pc(cj, ci);
-  } else if (ci->split && cj->split)
-    for (j = 0; j < 8; j++)
-      for (k = 0; k < 8; k++)
+  if (ci and cj are not neighbours)
+    return;
+  if (ci->split && cj->split) {
+    for (int j = 0; j < 8; j++)
+      for (int k = 0; k < 8; k++)
         comp_pair(ci->progeny[j], cj->progeny[k]);
-  else
-    for (j = 0; j < ci->count; j++)
-      for (k = 0; k < cj->count; k++)
+  } else {
+    for (int j = 0; j < ci->count; j++)
+      for (int k = 0; k < cj->count; k++)
         interact ci->parts[j] and cj->parts[k].
   }
-
-void comp_pair_cp(struct cell *ci, struct cell *cj) {
-  int k;
-  for (k = 0; k < ci->count; k++)
-    interact ci->parts[k] and cj center of mass.
+}
+
+void comp_pair_cp(struct cell *leaf, struct cell *c) {
+  if (c is a parent of leaf ||
+      c is a neighbour of a parent of leaf) {
+    for (int k = 0; k < 8; k++)
+     comp_pair_cp(leaf, c->progeny[k]);
+  } else if (leaf and c are not direct neighbours) {
+    for (int k = 0; k < leaf->count; k++)
+      interact leaf->parts[k] and c center of mass.
+  }
 }
@@
     \begin{lstlisting}[basicstyle=\scriptsize\tt]
-enum { tSELF , tPAIR_PP , tPAIR_PC , tCOM };
+enum { tSELF , tPAIR_PP , tPAIR_PC };
 void make_tasks(struct qsched *s, struct cell *ci, struct cell *cj) {
@@
       tid = qsched_addtask(s, tSELF, qsched_flags_none, &ci,
-                           sizeof(struct cell *), ci->count * ci->count);
+                           sizeof(struct cell *),
+                           ci->count * ci->count);
       qsched_addlock(s, tid, ci->res);
-      if (ci->split) qsched_addunlock(s, ci->com, tid);
-    }
-  } else if (ci and cj are well separated) {
-    data[0] = ci; data[1] = cj;
-    tid = qsched_addtask(s, tPAIR_PC, qsched_flags_none, data,
-                         sizeof(struct cell *) * 2, ci->count);
+      tid = qsched_addtask(s, tPAIR_PC, qsched_flags_none, &ci,
+                           sizeof(struct cell *), ci->count);
       qsched_addlock(s, tid, ci->res);
-    qsched_addunlock(s, cj->com, tid);
-    data[0] = cj; data[1] = ci;
-    tid = qsched_addtask(s, tPAIR_PC, qsched_flags_none, data,
-                         sizeof(struct cell *) * 2, cj->count);
-    qsched_addlock(s, tid, cj->res);
-    qsched_addunlock(s, ci->com, tid);
-  } else if (ci->split && cj->split && ci->count * cj->count > n_task * n_task)
+    }
+  } else if (ci->split && cj->split &&
+             ci->count * cj->count > n_task * n_task)
     for (j = 0; j < 8; j++)
-      for (k = 0; k < 8; k++) make_tasks(s, ci->progeny[j], cj->progeny[k]);
+      for (k = 0; k < 8; k++)
+        make_tasks(s, ci->progeny[j], cj->progeny[k]);
   else {
@@
     tid = qsched_addtask(s, tPAIR_PP, qsched_flags_none, data,
-                         sizeof(struct cell *) * 2, ci->count * cj->count);
+                         sizeof(struct cell *) * 2,
+                         ci->count * cj->count);
     qsched_addlock(s, tid, ci->res);
@@
     case tPAIR_PC:
-      comp_pair_pc( cells[0] , cells[1] );
+      comp_pair_pc( cells[0] , root );
       break;

### 2014-09-30 examples/plot_sorted.py
Modified the unit test such that all 13 directions are tested. The code computes the interactions correctly in all directions.

@@ -22,7 +22,7 @@ params = {'axes.labelsize': 16,
           'figure.subplot.wspace'  : 0.26  ,  # the amount of width reserved for blank space between subplots                              
           'figure.subplot.hspace'  : 0.22  ,  # the amount of height reserved for white space between subplots                             
 
-          'lines.markersize' : 6,
+          'lines.markersize' : 2,
           'lines.linewidth' : 2,
 
 #          'axes.formatter.limits' : (-2, 1),                                                                                              
@@ -36,9 +36,26 @@ import os
 
 print "Plotting..."
 
-names = ["side", "edge", "corner"]
+axis = [
+    1.0,  1.0,  1.0, 
+    1.0,  1.0,  0.0, 
+    1.0,  1.0, -1.0, 
+    1.0,  0.0,  1.0, 
+    1.0,  0.0,  0.0, 
+    1.0,  0.0, -1.0, 
+    1.0, -1.0,  1.0, 
+    1.0, -1.0,  0.0, 
+    1.0, -1.0, -1.0, 
+    0.0,  1.0,  1.0, 
+    0.0,  1.0,  0.0, 
+    0.0,  1.0, -1.0, 
+    0.0,  0.0,  1.0  
+  ]
 
-for orientation in range( 3 ):
+
+#names = ["side", "edge", "corner"]
+
+for orientation in range( 13 ):
 
     # Read Quickshed accelerations
     data=loadtxt( "interaction_dump_%d.dat"%orientation )
@@ -80,34 +97,32 @@ for orientation in range( 3 ):
     figure(frameon=True)
     
     subplot(311, title="Acceleration along X")
-    #plot(id[abs(errx_s) > 0.001], e_errx_s , 'rx')
-    plot(pos, e_errx_s , 'rx')
-    #text(id[-1], 0.18, "B-H: $%5.3f\\pm%5.3f$\n QuickShed: $%5.3f\\pm%5.3f$"%(meanx_bh, stdx_bh, meanx_new, stdx_new), backgroundcolor="w", va="top", ha="right" )
+    #plot(id[abs(errx_s) > 0.001], e_errx_s , 'ro')
+    plot(pos, e_errx_s , 'ro')
[…]

### 2014-09-30 examples/test_bh_sorted.c

@@ -169,6 +169,23 @@ const int axis_num_orth[13] = {
   /* (  0 , -1 ,  1 ) */ 1,
   /* (  0 ,  0 , -1 ) */ 2
 };
+/* const int axis_num_orth[13] = { */
+/*   /\* ( -1 , -1 , -1 ) *\/ 2, */
+/*   /\* ( -1 , -1 ,  0 ) *\/ 2, */
+/*   /\* ( -1 , -1 ,  1 ) *\/ 2, */
+/*   /\* ( -1 ,  0 , -1 ) *\/ 2, */
+/*   /\* ( -1 ,  0 ,  0 ) *\/ 2, */
+/*   /\* ( -1 ,  0 ,  1 ) *\/ 2, */
+/*   /\* ( -1 ,  1 , -1 ) *\/ 2, */
+/*   /\* ( -1 ,  1 ,  0 ) *\/ 2, */
+/*   /\* ( -1 ,  1 ,  1 ) *\/ 2, */
+/*   /\* (  0 , -1 , -1 ) *\/ 2, */
+/*   /\* (  0 , -1 ,  0 ) *\/ 2, */
+/*   /\* (  0 , -1 ,  1 ) *\/ 2, */
+/*   /\* (  0 ,  0 , -1 ) *\/ 2 */
+/* }; */
+
+
 const char axis_flip[27] = {1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0,
                             0, 0, 0, 0, 0, 0, 0, 0, 0, 0};
 
@@ -465,8 +482,11 @@ void get_axis(struct cell **ci, struct cell **cj, struct index **ind_i,
   orth1[3] = -d1;
   orth2[3] = -d2;
 
-  /* message("dx=[%.3e,%.3e,%.3e], axis=[%.3e,%.3e,%.3e], corr=%.3e.",
-    dx[0], dx[1], dx[2], axis[0], axis[1], axis[2], *corr); */
+  message("dx=[%.3e,%.3e,%.3e], axis=[%.3e,%.3e,%.3e]",
+	  dx[0], dx[1], dx[2], axis[0], axis[1], axis[2]);
+  message("N_planes= %d", *num_orth_planes);
+  message("o1=[%.3e,%.3e,%.3e] %.3e", orth1[0], orth1[1], orth1[2], orth1[3]);
[…]

### 2014-09-30 examples/plot_sorted.py
The sorted interactions test is now performed on all 26 directions and the plotting script also generates absolute error plots.

@@ -49,13 +49,26 @@ axis = [
     0.0,  1.0,  1.0, 
     0.0,  1.0,  0.0, 
     0.0,  1.0, -1.0, 
-    0.0,  0.0,  1.0  
+    0.0,  0.0,  1.0,
+    -1.0, -1.0, -1.0, 
+    -1.0, -1.0,  0.0, 
+    -1.0, -1.0,  1.0, 
+    -1.0,  0.0, -1.0, 
+    -1.0,  0.0,  0.0, 
+    -1.0,  0.0,  1.0, 
+    -1.0,  1.0, -1.0, 
+    -1.0,  1.0,  0.0, 
+    -1.0,  1.0,  1.0, 
+     0.0, -1.0, -1.0, 
+     0.0, -1.0,  0.0, 
+     0.0, -1.0,  1.0, 
+     0.0,  0.0, -1.0  
   ]
 
 
 #names = ["side", "edge", "corner"]
 
-for orientation in range( 13 ):
+for orientation in range( 26 ):
 
     # Read Quickshed accelerations
     data=loadtxt( "interaction_dump_%d.dat"%orientation )
@@ -121,7 +134,42 @@ for orientation in range( 13 ):
     #xlim(0, size(id)-1)
     grid()
 
-    savefig("accelerations_%d.png"%orientation)
+    savefig("accelerations_relative_%d.png"%orientation)
+    close()
+
+
+    # Plot -------------------------------------------------------
+    figure(frameon=True)
+    
+    subplot(311, title="Acceleration along X")
+    #plot(id[abs(errx_s) > 0.001], e_errx_s , 'ro')
+    plot(pos, accx_u , 'bx')
+    plot(pos, accx_s , 'ro')
+    text( 0., 0.1, "axis=( %d %d %d )"%(axis[orientation*3 + 0], axis[orientation*3 + 1], axis[orientation*3 + 2]) , ha='center', backgroundcolor='w', fontsize=14)
+    ylim(-700, 700)
+    grid()
+    
+    subplot(312, title="Acceleration along Y")
+    #plot(id[abs(erry_s) > 0.001], e_erry_s , 'ro')
+    plot(pos, accy_u , 'bx')
+    plot(pos, accy_s , 'ro')
+    text( 0., 0.1, "axis=( %d %d %d )"%(axis[orientation*3 + 0], axis[orientation*3 + 1], axis[orientation*3 + 2]) , ha='center', backgroundcolor='w', fontsize=14)
+    ylim(-700, 700) 
[…]

### 2014-09-30 examples/test_bh_sorted.c

@@ -2064,7 +2064,7 @@ void test_bh(int N, int nr_threads, int runs, char *fileName) {
  * sorted and unsortde interactions. Outputs then the tow sets of accelerations
  * for accuracy tests.
  * @param N_parts Number of particles in each cell
- * @param orientation Orientation of the cells ( 0 <= orientation < 13 )
+ * @param orientation Orientation of the cells ( 0 <= orientation < 26 )
  */
 void test_direct_neighbour(int N_parts, int orientation) {
 
@@ -2073,23 +2073,36 @@ void test_direct_neighbour(int N_parts, int orientation) {
   struct cell left, right;
 
   /* All 13 configurations */
-  const float cell_shift[13 * 3] = {
-    1.0,  1.0,  1.0, /* 0 */
-    1.0,  1.0,  0.0, /* 1 */
-    1.0,  1.0, -1.0, /* 2 */
-    1.0,  0.0,  1.0, /* 3 */
-    1.0,  0.0,  0.0, /* 4 */
-    1.0,  0.0, -1.0, /* 5 */
-    1.0, -1.0,  1.0, /* 6 */
-    1.0, -1.0,  0.0, /* 7 */
-    1.0, -1.0, -1.0, /* 8 */
-    0.0,  1.0,  1.0, /* 9 */
-    0.0,  1.0,  0.0, /* 10 */
-    0.0,  1.0, -1.0, /* 11 */
-    0.0,  0.0,  1.0  /* 12 */
+  const float cell_shift[26 * 3] = {
+     1.0,  1.0,  1.0, /* 0 */
+     1.0,  1.0,  0.0, /* 1 */
+     1.0,  1.0, -1.0, /* 2 */
+     1.0,  0.0,  1.0, /* 3 */
+     1.0,  0.0,  0.0, /* 4 */
+     1.0,  0.0, -1.0, /* 5 */
+     1.0, -1.0,  1.0, /* 6 */
+     1.0, -1.0,  0.0, /* 7 */
+     1.0, -1.0, -1.0, /* 8 */
+     0.0,  1.0,  1.0, /* 9 */
+     0.0,  1.0,  0.0, /* 10 */
+     0.0,  1.0, -1.0, /* 11 */
+     0.0,  0.0,  1.0, /* 12 */
+    -1.0, -1.0, -1.0, /* 13 */
+    -1.0, -1.0,  0.0, /* 14 */
[…]

### 2014-10-02 examples/plot_sorted.py
The error measurement now normalises the relative error using the norm of the acceleration and no the individual components.

@@ -88,9 +88,9 @@ for orientation in range( 26 ):
 
     # Build error ------------------------------------------------
     
-    errx_s = (accx_s - accx_u )/abs(accx_u) 
-    erry_s = (accy_s - accy_u )/abs(accy_u) 
-    errz_s = (accz_s - accz_u )/abs(accz_u) 
+    errx_s = (accx_s - accx_u )/sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
+    erry_s = (accy_s - accy_u )/sqrt(accy_u**2 + accy_u**2 + accz_u**2) 
+    errz_s = (accz_s - accz_u )/sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
     
     e_errx_s = errx_s#[abs(errx_s) > 0.001]
     e_erry_s = erry_s#[abs(erry_s) > 0.001]
@@ -177,14 +177,14 @@ for orientation in range( 26 ):
     # Plot -------------------------------------------------------
     bins = linspace(-3, 3, 10000)
     
-    e_errx_s = errx_s[(abs(errx_s) > 0.001) & (abs(errx_s) < 1.)]
-    e_erry_s = erry_s[(abs(erry_s) > 0.001) & (abs(erry_s) < 1.)]
-    e_errz_s = errz_s[(abs(errz_s) > 0.001) & (abs(errz_s) < 1.)]
+    e_errx_s = errx_s[(abs(errx_s) >= 0.000) & (abs(errx_s) < 1.)]
+    e_erry_s = erry_s[(abs(erry_s) >= 0.000) & (abs(erry_s) < 1.)]
+    e_errz_s = errz_s[(abs(errz_s) >= 0.000) & (abs(errz_s) < 1.)]
     
     
     
     figure(frameon=True)
-    subplot(311, title="Acceleration along X")#, yscale='log')
+    subplot(311, title="Acceleration along X")
     hist(e_errx_s, bins=bins, normed=1, histtype='step', rwidth=0.01, color='r', label="Sorted")
     legend(loc="upper right")
     xlim(-0.12, 0.12)

### 2014-10-02 examples/plot_sorted_all.py
Added a new unit test to th BH test case where a cell insteracting with all of its 27 neighbours.

@@ -1,0 +1,185 @@
+#!/usr/bin/env python2                                                                                                                     
+# -*- coding: utf-8 -*-                                                                                                                    
+import matplotlib
+matplotlib.use("Agg")
+from pylab import *
+import numpy
+# tex stuff                                                                                                                                
+#rc('font',**{'family':'serif','serif':['Palatino']})                                                                                      
+params = {'axes.labelsize': 16,
+          'axes.titlesize': 20,
+          'text.fontsize': 14,
+          'legend.fontsize': 14,
+          'xtick.labelsize': 14,
+          'ytick.labelsize': 14,
+          'text.usetex': True,
+
+          'figure.figsize' : (12,9),
+          'figure.subplot.left'    : 0.07,  # the left side of the subplots of the figure                                                  
+          'figure.subplot.right'   : 0.98  ,  # the right side of the subplots of the figure                                               
+          'figure.subplot.bottom'  : 0.09  ,  # the bottom of the subplots of the figure                                                   
+          'figure.subplot.top'     : 0.9  ,  # the top of the subplots of the figure                                                       
+          'figure.subplot.wspace'  : 0.26  ,  # the amount of width reserved for blank space between subplots                              
+          'figure.subplot.hspace'  : 0.22  ,  # the amount of height reserved for white space between subplots                             
+
+          'lines.markersize' : 2,
+          'lines.linewidth' : 2,
+
[…]

### 2014-10-02 examples/test_bh_sorted.c

@@ -40,7 +40,7 @@
 #define cell_maxparts 100
 #define task_limit 1e8
 #define const_G 1    // 6.6738e-8
-#define dist_min 0.5 /* Used fpr legacy walk only */
+#define dist_min 0.5 /* Used for legacy walk only */
 #define dist_cutoff_ratio 1.5
 #define iact_pair_direct iact_pair_direct_sorted
 
@@ -482,11 +482,11 @@ void get_axis(struct cell **ci, struct cell **cj, struct index **ind_i,
   orth1[3] = -d1;
   orth2[3] = -d2;
 
-  message("dx=[%.3e,%.3e,%.3e], axis=[%.3e,%.3e,%.3e]",
-	  dx[0], dx[1], dx[2], axis[0], axis[1], axis[2]);
-  message("N_planes= %d", *num_orth_planes);
-  message("o1=[%.3e,%.3e,%.3e] %.3e", orth1[0], orth1[1], orth1[2], orth1[3]);
-  message("o2=[%.3e,%.3e,%.3e] %.3e", orth2[0], orth2[1], orth2[2], orth2[3]);
+  /* message("dx=[%.3e,%.3e,%.3e], axis=[%.3e,%.3e,%.3e]", */
+  /* 	  dx[0], dx[1], dx[2], axis[0], axis[1], axis[2]); */
+  /* message("N_planes= %d", *num_orth_planes); */
+  /* message("o1=[%.3e,%.3e,%.3e] %.3e", orth1[0], orth1[1], orth1[2], orth1[3]); */
+  /* message("o2=[%.3e,%.3e,%.3e] %.3e", orth2[0], orth2[1], orth2[2], orth2[3]); */
 
   /* Make sure the sorts are ok. */
   /* for (int k = 1; k < (*ci)->count; k++)
@@ -2058,6 +2058,39 @@ void test_bh(int N, int nr_threads, int runs, char *fileName) {
   free(parts);
 }
 
+
+/* All 26 configurations for the cell tests */
+const float cell_shift[27 * 3] = {
+  1.0,  1.0,  1.0, /* 0 */
+  1.0,  1.0,  0.0, /* 1 */
+  1.0,  1.0, -1.0, /* 2 */
+  1.0,  0.0,  1.0, /* 3 */
+  1.0,  0.0,  0.0, /* 4 */
[…]

### 2014-11-13 paper/paper.tex
more corrections and proofreading, fixed some of the figures.

@@
 before task $A$ has completed.
-The tasks and their dependencies can be seen as the nodes and edges
-of a Directed Acyclic Graph (DAG) which can be
+The tasks and their dependencies can be seen as the nodes and edges,
+respectively, of a Directed Acyclic Graph (DAG) which can be
 traversed in topological order, executing the tasks at the nodes
@@
 as a {\em conflict} between two tasks.
-\fig{TaskConflicts} shows a task graph with conflicting tasks
+\fig{TaskConflicts} shows a task graph extended by conflicting tasks
 joined by thick dashed lines.
@@
 ordering on conflicting tasks.
-This unnecessary restriction on the order
-in which tasks can be scheduled can severely limit the
+This artificial restriction on the order
+in which tasks can be scheduled can, however, severely limit the
 parallelizability of a computation, especially in the presence
@@
 \begin{itemize}
-    \item {\em Correct}: All constraints, i.e.~dependencies and
+    \item {\em Correctnes}: All constraints, i.e.~dependencies and
         conflicts, must be correctly enforced,
-    \item {\em Fast}: The overheads associated with task management
+    \item {\em Speed}: The overheads associated with task management
         should be as small as possible,
-    \item {\em Memory/cache-efficient}: Tasks accessing similar
+    \item {\em Memory/cache efficiency}: Tasks accessing similar
         sets of data should be preferentially executed on the
         same core to preserve memory/cache locality as far as possible, and
-    \item {\em Parallel-efficient}: Tasks should be executed in an order
+    \item {\em Parallel efficiency}: Tasks should be executed in an order
         that sufficient work is available for all computational
@@
 
-\begin{itemize}
-  \item Two flavours of dependency generation: Through spawning as in Cilk,
-    or automatically as per QUARK, OmpSS, etc...
-  \item Limitation of spawning with complex task hierarchies.
-  \item Limitation of dependency guessing is sometimes over-protective dependencies.
-  \item We let the user define dependencies explictly.
-  \item Con: Some effort needed to get dependencies right,
-  \item Pro: All dependencies known at the start of a computation, better
-    scheduling decisions can be made (see QR).
-  \item A word on task granularity...
-\end{itemize}
-
 From a programmer's perspective, there are two main paradigms for generating
@@
 \begin{itemize}
-  \item Implicitly via spawning and waiting, e.g.~as is done in Cilk, or
-  \item Automatic extraction from data dependencies, e.g.~as is done in OmpSs.
+  \item Implicitly via spawning and waiting, e.g.~as is done in Cilk
+    \cite{ref:Blumofe1995}, or
+  \item Automatic extraction from data dependencies, e.g.~as is done in OmpSs
+    \cite{ref:Duran2011}.
 \end{itemize}
@@
 single task, i.e.~if the task DAG is a tree, each task {\em spawns}, or
-creates, its dependent tasks after its completion (see \fig{Spawn}).
+creates, its dependent tasks after its completion (see \fig{Spawn}a).
 Hence for the tasks $A$--$E$ in \fig{Tasks}, task $A$ would spawn
@@
 and first spawns tasks $D$ and $F$, and waits for both their completion
-before doing its own computations.
+before doing its own computations (see \fig{Spawn}b).
 
 Although simple to use, this implicit dependency management
-limits the types of DAGs that can be represented, i.e.~for
+limits the types of DAGs that can be represented, e.g.~for
 all the tasks in \fig{Tasks}, using such a spawning and waiting model
@@
 created.
-This approach usually relies on compiler extensions, e.g.~ {\tt pragma}s
+This approach usually relies on compiler extensions, e.g.~{\tt pragma}s
 in C/C++, or a system of function call wrappers, to describe the task parameters
@@
 This approach allows programmers to specify rich dependency hierarchies
-with very little effort, they still only allow for one type of relationship,
+with very little effort, i.e.~without having to explicitly think about
+dependencies at all, yet they still only allow for one type of relationship,
 i.e.~dependencies, and lack the ability to deal with conflicts as
@@
     structure of the task graph generated,
-  \item Knowing the structure of the task graph before execution
+  \item Knowing the complete structure of the task graph before execution
     allows the task scheduler to make more informed decisions
@@
 \noindent The obvious disadvantage is the burden of producing a correct
-task graph is placed on the user.
+task graph is placed on the programmer.
 Although some problems such as cyclic dependencies can be detected
 automatically, there is no automatic way to detect whether the
-dependencies actualy reflect the intent of the user.
+dependencies actualy reflect the intent of the programmer.
 Due to this added complexity, we consider QuickSched to be
 a tool not designed for casual parallel programmers, but for 
-those interested in investing more programming effort to achieve
+those interested in investing a bit more programming effort to achieve
 better performance.
 
-Conflicts between tasks or groups of tasks are not specified directly,
+As opposed to the dependencies,
+conflicts between tasks or groups of tasks are not specified directly,
 but are instead modeled as exclusive locks on a shared resource
@@
 
-The partitioning of the computation into tasks is also left entirely
-to the user.
+As with all other task-based libraries, the partitioning of the
+computation into tasks is also left entirely to the programmer.
 In theory, any program can be automatically converted to a task-based
@@
 best fit the task-based paradigm, e.g.~as in the examples in the
-following sections, or as in \cite{ref:Gonnet2014}.
-This process requires careful evaluation and is probably best
+following sections, or as in \cite{ref:Gonnet2014,ref:Buttari2009,ref:Ltaief2012}.
+This process requires careful evaluation of the underlying
+computation, and is probably best
 not left as an automatic transformation of an otherwise serial code.
@@
 and {\em queue}.
-
 The task and resource objects are used
@@
 Given a set of tasks with similar resources for which all
-dependencies are resolved, the queue then decides which
+dependencies are resolved, it is up to the queue to decide which
 tasks to prioritize.
-This decision is made based on the length of the critical
-path of the dependencies of each task.
 
@@
 \begin{figure}
-    \centerline{\epsfig{file=figures/TaskWeight.pdf,height=0.3\textwidth}}
+    \centerline{\epsfig{file=figures/TaskWeight.pdf,height=0.4\textwidth}}
     \caption{Computation of the task weight.
@@
 If this process fails at any point (line~11), the
-previously set hold counters are decremented (line~12)
-and the lock is released (line~13).
+previously set hold counters are decremented (line~13)
+and the lock is released (line~14).
 The procedure then returns {\tt 1} or {\tt 0} if the resource
@@
 The main job of the task queues is, given a set of ready tasks,
-to find the task with maximum weight whose resources can all
+to find the task with maximum weight, i.e.~the task along the
+longest critical path, whose resources can all
 be locked, and to do so as efficiently as possible.
@@
 in the first position.
-Maintaining this heap structure thus requires \oh{\log n}
+Maintaining this heap structure requires \oh{\log n}
 operations for both insertion and deletion, i.e. for the
@@
 
-The array of tasks is then traversed as if it were sorted,
+Unfortunately, there is no way of efficiently traversing all
+the elements in the heap in decreasing order.
+The array of tasks is therefore traversed as if it were sorted,
 returning the first task that can be locked.
@@
 $\lfloor n/k\rfloor -1$ other tasks.
+Although this is not a particularly good lower bound, it turns
+out to be quite sufficient in practice.
 
@@
 
+Note that this approach of sequentially locking multiple resources
+is prone to the so-called ``dining philosophers'' problem, i.e.~if
+two tasks attempt, simultaneously, to lock the resources $A$ and $B$;
+and $B$ and $A$, respectively, via separate queues, their respective calls
+to {\tt queue\_get} will potentially fail perpetually.
+This type of deadlock, however, is easily avoided by sorting the
+resources in each task according to some global creiteria, e.g.~the
+resource ID or the address in memory of the resource.
+
 \subsection{Scheduler}
@@
 A version using {\tt pthreads} \cite{ref:Pthreads1995}
-directly is also available.
+directly\footnote{In most environments, OpenMP is implemented
+on top of {\tt pthreads}, e.g. gcc's libgomp.} is also available.
 The parallel section consists of a loop (lines~7--10) in
@@
 void qsched_enqueue(qsched *s, struct task *t) {
-  int k, qid, score[s->nr_queues];
-  for (k = 0; k < s->nr_queues; k++)
+  int best = 0, score[s->nr_queues];
+  for (int k = 0; k < s->nr_queues; k++)
     score[k] = 0;
-  for (j = 0; j < t->nr_locks; j++)
-    score[t->locks[j]->owner] += 1;
-  for (j = 0; j < t->nr_uses; j++)
-    score[t->uses[j]->owner] += 1;
-  for (qid = 0, k = 1; k < s->nr_queues; k++)
-    if (score[k] > score[qid])
-      qid = k;
-  queue_put(&s->queues[qid], t);
+  for (int k = 0; k < t->nr_locks; k++) {
+    int qid = t->locks[k]->owner;
+    if (++score[qid] > score[best]) best = qid;
+  }
+  for (int k = 0; k < t->nr_uses; k++) {
+    int qid = t->uses[k]->owner;
+    if (++score[qid] > score[best]) best = qid;
+  }
+  queue_put(&s->queues[best], t);
 }
@@
 task resources ``owned'', or last used, by each queue.
-In lines~9--11 the queue with the highest such score is 
-chosen on which the task is then put (line~12).
+The task is then sent to the queue with the highest such score
+(line~13).
 
@@
 
-This section presents two test cases showing
+This section presents two test cases that show
 how QuickSched can be used in real-world applications, and
-providing benchmarks to assess its efficiency and scalability.
+provides benchmarks to assess its efficiency and scalability.
 The first test is the tiled QR decomposition originally
@@
 {\tt -O2 -march=native} flags and run on
-a 64-core AMD Opteron 6376 machine running
-at 2.67\,GHz.
+a 64-core AMD Opteron 6376 machine at 2.67\,GHz.
 
@@
 near the end of the computation.
+
+Since in QuickSched the entire task structure is known explicitly
+in advance, the scheduler ``knows'' that the DGEQRF tasks all
+lie on the longest critical path and therefore executes them as
+soon as possible.
+OmpSs, does not exploit this knowledge, resulting in the less efficient
+scheduling seen in \fig{QRTasks}.
 

### 2014-11-18 examples/plot_sorted_all.py
Correction of the bug affecting the plotting scripts in the remaining scripts

@@ -63,11 +63,11 @@ accz_bh=data[:,13]
 # Build error ------------------------------------------------
     
 errx_s = (accx_s - accx_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
-erry_s = (accy_s - accy_u ) / sqrt(accy_u**2 + accy_u**2 + accz_u**2) 
+erry_s = (accy_s - accy_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
 errz_s = (accz_s - accz_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
 
 errx_bh = (accx_bh - accx_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
-erry_bh = (accy_bh - accy_u ) / sqrt(accy_u**2 + accy_u**2 + accz_u**2) 
+erry_bh = (accy_bh - accy_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
 errz_bh = (accz_bh - accz_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
 
 e_errx_s = errx_s#[abs(errx_s) > 0.001]

### 2014-11-20 src/queue.c
destroy threads on free.

@@ -47,7 +47,7 @@
  
 int queue_get ( struct queue *q , struct qsched *s , int insist ) {
 
-    int k, j, temp, tid, *inds, w, count;
+    int k, j, temp, tid, *inds, count;
     struct task *tasks = s->tasks;
 
     /* Should we even try? */
@@ -88,7 +88,7 @@ int queue_get ( struct queue *q , struct qsched *s , int insist ) {
         inds[k] = inds[ count ];
     
         /* Fix the heap. */
-        w = tasks[ inds[k] ].weight;
+        long long int w = tasks[ inds[k] ].weight;
         if ( k > 0 && w > tasks[ inds[(k-1)/2] ].weight )
             while ( k > 0 ) {
                 j = (k - 1) / 2;
@@ -144,7 +144,7 @@ void queue_put ( struct queue *q , struct qsched *s , int tid ) {
 
     int ind, j, temp;
     struct task *tasks = s->tasks;
-    int *inds, *inds_new, w;
+    int *inds, *inds_new;
     
     /* Lock this queue. */
     if ( lock_lock( &q->lock ) != 0 )
@@ -180,7 +180,7 @@ void queue_put ( struct queue *q , struct qsched *s , int tid ) {
     inds[ind] = tid;
     
     /* Bubble up the new entry. */
-    w = tasks[ inds[ind] ].weight;
+    long long int w = tasks[ inds[ind] ].weight;
     while ( ind > 0 ) {
         j = (ind - 1)/2;
         if ( tasks[ inds[j] ].weight < w ) {

### 2014-11-20 paper/paper.tex
yet another iteration, fixed the BH_times figure, formatted the code correctly.

@@
 
-The particle data is sorted hierarchically, following the
-octree structure.
+In our implementation, the particle data is sorted hierarchically,
+following the octree structure.
 Unlike in many codes, where the leaves store an array of
-pointers to the underlying particles, the cells, at all
+pointers to the underlying particles, which are not necessarily
+contiguous in memory, the cells, at all
 levels, store only a pointer to the first of their own particles,
@@
 
-The task-based implementation will consist of three
+The task-based implementation consists of three
 types of tasks:
 \begin{itemize}
-    \item {\em Self}-interactions in which all particles
+    \item {\em Self}-interactions, in which all particles
         in a single cell interact with all other particles in the
         same cell,
-    \item {\em Particle-particle} pair interactions in which
+    \item {\em Particle-particle} pair interactions, in which
         all particles in a cell interact with all
         particles in another cell,
-    \item {\em Particle-cell} pair interactions in which
+    \item {\em Particle-cell} pair interactions, in which
         all particles in one cell are interacted with the
@@
 and {\tt NULL} as its two cell parameters.
-The function recurses as follows:
+The function recurses as follows (line numbers refer to \fig{MakeTasks}:
 \begin{itemize}
-    \item If called with a single (line~6), split (line~7) cell,
-        recurse over all the cell's sub-cells (line~9), and all
-        pairs of the cell's sub-cells (line~11),
+    \item If called with a single, split cell (lines~6--7),
+        recurse over all the cell's sub-cells, and all
+        pairs of the cell's sub-cells (lines~8--11),
     \item If called with a single unsplit cell (line~13),
-        create a self-interaction task as well as a particle-cell
-        task on that cell (line~14),
-    \item If called with two non-neighbouring cells (line~21),
-        do nothing, as these interactions
-        will be computed by the particle-cell task,
+        create a self-interaction task (line~14) as well as a particle-cell
+        task on that cell (line~18),
     \item If called with two neighbouring cells and both cells
-        are split (line~33),
+        are split (line~22),
         recurse over all pairs of sub-cells spanning
-        both cells (line~37), and
+        both cells (lines~24--26), and
     \item If called with two neighbouring cells
         and one of the cells are not split, create
-        a particle-particle pair task over both cells.
+        a particle-particle pair task over both cells (line~29),
+    \item If called with two non-neighbouring cells,
+        do nothing, as these interactions
+        will be computed by the particle-cell task.
 \end{itemize}
 \noindent where every interaction task additionally locks
-the cells on which it operates (lines~16, 25, 30, and 42--43).
+the cells on which it operates (lines~17, 20, and 32--33).
 In order to prevent generating
@@
 if the cells contain more than a minimum number $n_\mathsf{task}$
-of threads each (lines~7 and~34).
-
-The particle-particle pair interaction tasks are implemented
+of threads each (lines~7 and~23).
+
+As shown in \fig{BHTasks}, the particle-self and particle-particle pair
+interaction tasks are implemented
 by computing the interactions between all particle pairs spanning
-both cells in a double for-loop.
+both cells in a double for-loop (lines~9--11 and~22-24 therein).
+Some extra logic (lines~2--7 and~15--20) is added to deal with
+split cells that did not contain enough particles to warrant the
+generation of finer tasks.
 The particle-cell interactions for each leaf node are computed by
-traversing the tree recursively starting from the root node:
+traversing the tree recursively starting from the root node and:
 \begin{itemize}
@@
     a hierarchical parent of the leaf node, recurse over the
-    node's sub-cells,
+    node's sub-cells (lines~29--32),
   \item Otherwise, compute the interaction between the leaf node's
-    center of mass and all the particles in the leaf node.
+    center of mass and all the particles in the leaf node (lines~33--35).
 \end{itemize}
@@
 in the Barnes-Hut algorithm in that the particle-cell interactions
-are grouped per leaf, with each leaf doing its own tree walk.
+are grouped per leaf, with each leaf doing its own tree walk,
+as opposed to doing a tree walk for each individual particle.
 This approach was chosen to maximize the memory locality
@@
 contiguous in memory, and are thus more likely to remain in the
-lowest-level cache.
+lowest-level cache during the entire tree-walk.
 
@@
 Using the above scheme generated 97\,553 tasks, of which
-512 self-interaction tasks, ??? particle-particle interaction
-task, and ??? particle-cell interaction tasks.
-A total of 141\,840 dependencies were generated, along with
-104\,392 locks on 37\,449 resources.
+512 self-interaction tasks, 5\,068 particle-particle interaction
+task, and 32\,768 particle-cell interaction tasks.
+A total of 43\,416 locks on 37\,449 resources were generated.
 
@@
         computed over 1\,000\,000 particles.
-        Solving the N-Body problem takes 3.3\,s, achieving 60\% parallel
+        Solving the N-Body problem takes 323\,ms, achieving 75\% parallel
         efficiency, over all 64 cores.
@@
     \label{fig:BHResults}
+\end{figure}
+
+\begin{figure}
+    \centerline{\epsfig{file=figures/tasks_bh_dynamic_64.pdf,width=0.9\textwidth}}
+    \caption{Task scheduling of the Barnes-Hut tree-code on 64 cores.
+      The red tasks correspond to particle self-interactions, the green
+      tasks to the particle-particle pair interactions, and the blue
+      tasks to the particle-cell interactions.}
+    \label{fig:BHTasksPlot}
 \end{figure}
@@
 of the underlying hardware.
-\fig{BHTasks} shows the accumulated cost of each task type and of 
+\fig{BHTimes} shows the accumulated cost of each task type and of 
 QuickSched over the number of cores.
-At 64 cores, the scheduler overheads account for only 4.7\% of
+At 64 cores, the scheduler overheads account for only $\sim 1$\% of
 the total computational cost, whereas,
-as of 30 cores, the cost of both pair types grow by up to
-50\%.
+as of 32 cores, the cost of both pair types grow by up to
+40\%.
 This is most probably due to memory bandwidth restrictions, as
-the cost of the self interaction tasks, which do twice as much
-computation per memory access, only grow by up to 15\%.
+the cost of the particle-cell interaction tasks, which do significantly more
+computation per memory access, only grow by up to 10\%.
 
@@
     \caption{Accumulated cost of each task type and of the overheads
-        associated with {\tt qsched\_gettask}.
-        As of $\sim 30$ cores, the cost of both pair interaction task
-        types grow by up to 50\%.
-        The cost of the self interactions, which entail twice as much
-        computation per memory access, grow only by at most 15\%.
-        The scheduler overheads make up less than 5\% of the total
-        time.
-        }
+        associated with {\tt qsched\_gettask}, summed over all cores.
+        As of 32 cores, the cost of both pair interaction task
+        types grow by up to 40\%.
+        The cost of the particle-cell interactions, which entail significantly more
+        computation per memory access, grow only by at most 10\%.
+        The scheduler overheads, i.e.~{\tt qsched\_gettask},
+        make up less than 1\% of the total time.}
     \label{fig:BHTimes}
@@
 require less than 1\,000 lines of code each.
+For a more complex, large-scale example of a task-based computation
+based on the same algorithms, we refer to \cite{ref:Gonnet2014},
+for which the scheduler was originally designed, and from
+which QuickSched was back-ported.
 
@@
         QuickSched does not take the resource locks into
-        account when selecting tasks in the queues,
+        account when selecting tasks in the queues, e.g.~it
+        may be advantageous, in some settings, to avoid tasks
+        which are involved in too many potential conflicts
+        and would therefore restrict the maximum degree of
+        parallelism when scheduled.
     \item {\em Work-stealing}: During work-stealing, the
@@
 \section*{Acknowledgments}
-The author would like to thank Sam Townsend for implementing the
-kernels in the task-based QR decomposition as part of his MSc Thesis
-at Durham University.
-The author also thanks Lydia Heck of the Institute for
+The authors would like to thank Lydia Heck of the Institute for
 Computational Cosmology at Durham University for providing access
@@
 
-The algorithms, as described in the previous section, have
-all been implemented as part of the Open-Source C-language
-QuickSched library.\footnote{\url{http://sourceforge.net/projects/quicksched/}}
-This section describes the interface functions and how they
-are called, and the following section contains
-examples of how QuickSched can be used.
+This section describes the QuickSched interface functions and how they
+are called.
 
@@
 \begin{itemize}
-    \item {\tt void qsched\_init( struct qsched *s , int nr\_queues , int flags )} \\
+    \item {\tt void qsched\_init(struct qsched *s, int nr\_queues, int flags)} \\
         Initializes a {\tt qsched} object with the given number of queues.
@@
         \vspace{1mm}
-    \item {\tt void qsched\_free( struct qsched *s )} \\
+    \item {\tt void qsched\_free(struct qsched *s)} \\
         Releases all the memory and other resources allocated by the
@@
         \vspace{1mm}
-    \item {\tt void qsched\_reset( struct qsched *s )} \\
+    \item {\tt void qsched\_reset(struct qsched *s)} \\
         Clears the tasks and resources in the given {\tt qsched},
@@
         \vspace{1mm}
-    \item {\tt qsched\_task\_t qsched\_addtask( struct qsched *s , int type , unsigned int flags , void *data , int data\_size , int cost )} \\
+    \item {\tt qsched\_task\_t qsched\_addtask(struct qsched *s, int type, unsigned int flags, void *data, int data\_size, int cost)} \\
         Creates a new task within the given {\tt qsched} and returns
@@
         \vspace{1mm}
-    \item {\tt qsched\_res\_t qsched\_addres( struct qsched *s , int owner , qsched\_res\_t parent )} \\
+    \item {\tt qsched\_res\_t qsched\_addres(struct qsched *s, int owner, qsched\_res\_t parent)} \\
         Creates a new resource within the given {\tt qsched} and returns
@@
         \vspace{1mm}
-    \item {\tt void qsched\_addlock( struct qsched *s , qsched\_task\_t t , qsched\_res\_t res )} \\
+    \item {\tt void qsched\_addlock(struct qsched *s, qsched\_task\_t t, qsched\_res\_t res)} \\
         Append the resource {\tt res} to the task {\tt t}'s list of
@@
         \vspace{1mm}
-    \item {\tt void qsched\_adduse( struct qsched *s , qsched\_task\_t t , qsched\_res\_t res )} \\
+    \item {\tt void qsched\_adduse(struct qsched *s, qsched\_task\_t t, qsched\_res\_t res)} \\
         Similar to {\tt qsched\_addlock}, yet the resource is only used and
@@
         \vspace{1mm}
-    \item {\tt void qsched\_addunlock( struct qsched *s , qsched\_task\_t ta , qsched\_task\_t tb )} \\
+    \item {\tt void qsched\_addunlock(struct qsched *s, qsched\_task\_t ta, qsched\_task\_t tb)} \\
         Appends the task {\tt tb} to the list of tasks that the task {\tt ta}
@@
         \vspace{1mm}
-    \item {\tt void qsched\_run( struct qsched *s , int nr\_threads , qsched\_funtype fun )} \\
+    \item {\tt void qsched\_run(struct qsched *s, int nr\_threads, qsched\_funtype fun)} \\
         Executes the tasks in the given {\tt qsched} using {\tt nr\_threads}
@@
 This may be a problem if other parts of the user application
-are running concurrently in the background.
+are running concurrently in the background outside of QuickSched.
 Calling {\tt qsched\_init} with the {\tt qsched\_flag\_yield}
@@
     \begin{lstlisting}
-void exec_fun ( int type , void *data ) {
+void exec_fun(int type, void *data) {
   int *idata = (int *)data;
   int i = idata[0], j = idata[1], k = idata[2];
-  switch ( type ) {
+  switch (type) {
     case tDGEQRF:
-      DGEQRF( A[i,j] , ... );
+      DGEQRF(A[i, j], ...);
       break;
     case tDLARFT:
-      DLARFT( A[i,j] , A[i,i] , ... );
+      DLARFT(A[i, j], A[i, i], ...);
       break;
     case tDTSQRF:
-      DTSQRF( A[i,j] , A[j,j]... );
+      DTSQRF(A[i, j], A[j, j], ...);
       break;
     case tDSSRFT:
-      DSSRFT( A[i,j] , A[i,k] , A[k,j] , ... );
+      DSSRFT(A[i, j], A[i, k], A[k, j], ...);
       break;
     default:
-      error( "Unknown task type." );
+      error("Unknown task type.");
   }
@@
 void comp_self(struct cell *c) {
-  if (c->split)
+  if (c->split) {
     for (int j = 0; j < 8; j++) {
@@
     }
-  else
+  } else {
     for (int j = 0; j < c->count; j++)
@@
         interact c->parts[j] and c->parts[k].
+  }
 }
@@
     \begin{lstlisting}[basicstyle=\scriptsize\tt]
-enum { tSELF , tPAIR_PP , tPAIR_PC };
+enum { tSELF, tPAIR_PP, tPAIR_PC };
 void make_tasks(struct qsched *s, struct cell *ci, struct cell *cj) {
@@
   if (cj == NULL) {
-    if (ci->split && ci->count > n_task)
+    if (ci->split && ci->count > n_task) {
       for (j = 0; j < 8; j++) {
@@
       }
-    else {
+    } else {
       tid = qsched_addtask(s, tSELF, qsched_flags_none, &ci,
@@
   } else if (ci->split && cj->split &&
-             ci->count * cj->count > n_task * n_task)
+             ci->count * cj->count > n_task * n_task) {
     for (j = 0; j < 8; j++)
@@
         make_tasks(s, ci->progeny[j], cj->progeny[k]);
-  else {
+  } else {
     data[0] = ci; data[1] = cj;
@@
     qsched_addlock(s, tid, cj->res);
-    if (ci->split && cj->split) {
-      qsched_addunlock(s, ci->com, tid);
-      qsched_addunlock(s, cj->com, tid);
-    }
   }
@@
     \begin{lstlisting}
-void exec_fun ( int type , void *data ) {
+void exec_fun(int type, void *data) {
   struct cell **cells = (struct cell **)data;
-  switch ( type ) {
+  switch (type) {
     case tSELF:
-      comp_self( cells[0] );
+      comp_self(cells[0]);
       break;
     case tPAIR_PP:
-      comp_pair( cells[0] , cells[1] );
+      comp_pair(cells[0], cells[1]);
       break;
     case tPAIR_PC:
-      comp_pair_pc( cells[0] , root );
+      comp_pair_pc(cells[0], root);
       break;
     case tCOM:
-      comp_com( cells[0] );
+      comp_com(cells[0]);
       break;
     default:
-      error( "Unknown task type." );
+      error("Unknown task type.");
   }

### 2014-11-23 paper/paper.tex
added paragraph on Gadget-2.

@@
 
+For comparison, the same computations were run using the popular
+astrophysics simulation software Gadget-2 \cite{ref:Springel2005},
+using a traditional Barnes-Hut implementation based on octrees
+and distributed-memory parallelism based on domain decompositions
+and MPI \cite{ref:Snir1998}.
+To achieve the same accuracy, an opening angle of 0.5 was used.
+On a single core, the task-based tree traversal is already 1.9$\times$
+faster than Gadget-2, due to the cache efficiency of the task-based
+computations, which, by design, maximize the amount of computation
+per memory access.
+At 59 cores, where Gadget-2 performs best, the task-based tree traversal is
+2.51$\times$ faster, and at the full 64 cores it is 4$\times$ faster,
+due to the better strong scaling of the task-based approach as opposed
+to the MPI-based parallelism in Gadget-2.
+
 \begin{figure}
@@
         efficiency, over all 64 cores.
+        For comparison, timings are shown for the same computation using
+        the popular astrophysics code Gadget-2.
+        The scaling for Gadget-2 (left) is shown relative to the performance of
+        QuickSched, whereas the parallel efficiency (right) is computed relative
+        to Gadget-2 on a single core.
         }

### 2014-11-25 paper/paper.tex
final corrections.

@@
 \begin{itemize}
-    \item {\em Correctnes}: All constraints, i.e.~dependencies and
+    \item {\em Correctness}: All constraints, i.e.~dependencies and
         conflicts, must be correctly enforced,
@@
 \begin{figure}
-    \centerline{\epsfig{file=figures/Spawn.pdf,width=0.7\textwidth}}
+    \centerline{\epsfig{file=figures/Spawn.pdf,width=0.9\textwidth}}
     \caption{Two different task graphs and how they can be implemented
@@
 scheduling tasks, which is usually constant per task, will
-quickly destory any performance gains from parallelism.
+quickly destroy any performance gains from parallelism.
 Starting from a per-statement set of tasks, it is therefore
@@
 \begin{figure}
-    \centerline{\epsfig{file=figures/QSched.pdf,width=0.7\textwidth}}
+    \centerline{\epsfig{file=figures/QSched.pdf,width=0.8\textwidth}}
     \caption{Schematic of the QuickSched task scheduler.
@@
 \begin{figure}
-    \centerline{\epsfig{file=figures/Resources.pdf,width=0.6\textwidth}}
+    \centerline{\epsfig{file=figures/Resources.pdf,width=0.7\textwidth}}
     \caption{A hierarchy of cells (left) and the hierarchy of
@@
 This type of deadlock, however, is easily avoided by sorting the
-resources in each task according to some global creiteria, e.g.~the
+resources in each task according to some global criteria, e.g.~the
 resource ID or the address in memory of the resource.
@@
 \begin{figure}
-    \centerline{\epsfig{file=figures/QR.pdf,width=0.8\textwidth}}
+    \centerline{\epsfig{file=figures/QR.pdf,width=0.9\textwidth}}
     \caption{Task-based QR decomposition of a matrix consisting
@@
 task additionally modifies the lower triangular part of the $(j,j)$th tile.
+Although the tile-based QR decomposition requires only dependencies,
+i.e.~no additional conflicts are needed to avoid concurrent access to
+the matrix tiles, we still model each tile as a separate resource
+in QuickSched such that the scheduler can preferrentially assign
+tasks using the same tiles to the same thread.
 
@@
 \begin{quote}
-    \tt --disable-yield --schedule=socket --cores-per-socket=16 --num-sockets=4
+    \tt --disable-yield --schedule=socket --cores-per-socket=16 \\--num-sockets=4
 \end{quote}
@@
 \begin{figure}
-    \centerline{\epsfig{file=figures/QR_scaling.pdf,width=0.9\textwidth}}
+    \centerline{\epsfig{file=figures/QR_scaling.pdf,width=\textwidth}}
     \caption{Strong scaling and parallel efficiency of the tiled QR decomposition
@@
 \begin{figure}
-    \centerline{\epsfig{file=figures/tasks_qr.pdf,width=0.9\textwidth}}
-    \centerline{\epsfig{file=figures/tasks_qr_ompss.pdf,width=0.9\textwidth}}
+    \centerline{\epsfig{file=figures/tasks_qr.pdf,width=\textwidth}}
+    \centerline{\epsfig{file=figures/tasks_qr_ompss.pdf,width=\textwidth}}
     \caption{Task scheduling in QuickSched (above) and OmpSs (below)
@@
 
-The Barnes-Hut tree-code is an algorithm to approximate the
+The Barnes-Hut tree-code \cite{ref:Barnes1986}
+is an algorithm to approximate the
 solution of an $N$-body problem, i.e.~computing all the
@@
 \begin{figure}
-    \centerline{\epsfig{file=figures/BH_scaling.pdf,width=0.9\textwidth}}
+    \centerline{\epsfig{file=figures/BH_scaling.pdf,width=\textwidth}}
     \caption{Strong scaling and parallel efficiency of the Barnes-Hut tree-code
@@
 \begin{figure}
-    \centerline{\epsfig{file=figures/tasks_bh_dynamic_64.pdf,width=0.9\textwidth}}
+    \centerline{\epsfig{file=figures/tasks_bh_dynamic_64.pdf,width=\textwidth}}
     \caption{Task scheduling of the Barnes-Hut tree-code on 64 cores.
@@
 40\%.
-This is most probably due to memory bandwidth restrictions, as
+This is due to memory bandwidth restrictions, as
 the cost of the particle-cell interaction tasks, which do significantly more

### 2014-11-26 examples/plot_sorted.py
Added the multipole expansion equations to the repository.

@@ -114,14 +114,14 @@ for orientation in range( 26 ):
     #plot(id[abs(errx_s) > 0.001], e_errx_s , 'ro')
     plot(pos, e_errx_s , 'ro')
     text( 0., 0.1, "axis=( %d %d %d )"%(axis[orientation*3 + 0], axis[orientation*3 + 1], axis[orientation*3 + 2]) , ha='center', backgroundcolor='w', fontsize=14)
-    ylim(-0.02, 0.02)
+    ylim(-0.05, 0.05)
     grid()
     
     subplot(312, title="Acceleration along Y")
     #plot(id[abs(erry_s) > 0.001], e_erry_s , 'ro')
     plot(pos, e_erry_s , 'ro')
     text( 0., 0.1, "axis=( %d %d %d )"%(axis[orientation*3 + 0], axis[orientation*3 + 1], axis[orientation*3 + 2]) , ha='center', backgroundcolor='w', fontsize=14)
-    ylim(-0.02, 0.02)  
+    ylim(-0.05, 0.05)  
     grid()
     
     subplot(313, title="Acceleration along Z")
@@ -131,11 +131,11 @@ for orientation in range( 26 ):
 
     legend(loc="upper right")
     
-    ylim(-0.02, 0.02)
+    ylim(-0.05, 0.05)
     #xlim(0, size(id)-1)
     grid()
 
-    savefig("accelerations_relative_%d.png"%orientation)
+    savefig("matthieu_accelerations_relative_%d.png"%orientation)
     close()
 
 

### 2014-11-26 examples/test_bh_sorted.c

@@ -37,7 +37,7 @@
 
 /* Some local constants. */
 #define cell_pool_grow 1000
-#define cell_maxparts 100
+#define cell_maxparts 1
 #define task_limit 1e8
 #define const_G 1    // 6.6738e-8
 #define dist_min 0.5 /* Used for legacy walk only */
@@ -162,6 +162,8 @@ static inline void dipole_iact(const struct dipole *d, const float *x,
     offset = 0.0f;
   }
 
+  //offset = 0.f;
+
   /* Compute the first dipole. */
   for (r2 = 0.0f, k = 0; k < 3; k++) {
     dx[k] = x[k] - (d->x[k] * inv_mass + offset * d->axis[k]);
@@ -245,6 +247,8 @@ static inline void tripole_iact(const struct tripole *d, const float *x,
     offset2 = 0.0f;
   }
 
+  //offset1 = offset2 = 0.;
+
   /* Compute the first dipole. */
   for (r2 = 0.0f, k = 0; k < 3; k++) {
     dx[k] = x[k] - (d->x[k] * inv_mass + offset1 * d->axis1[k]);
@@ -302,12 +306,14 @@ const float axis_shift[13 * 3] = {
   5.773502691896258e-01, 5.773502691896258e-01, -5.773502691896258e-01,
   7.071067811865475e-01, 0.0, 7.071067811865475e-01,
   1.0, 0.0, 0.0,
-  7.071067811865475e-01, 0.0,
-  -7.071067811865475e-01, 5.773502691896258e-01, -5.773502691896258e-01,
-  5.773502691896258e-01, 7.071067811865475e-01, -7.071067811865475e-01, 0.0,
-  5.773502691896258e-01, -5.773502691896258e-01, -5.773502691896258e-01, 0.0,
-  7.071067811865475e-01, 7.071067811865475e-01, 0.0, 1.0, 0.0, 0.0,
-  7.071067811865475e-01, -7.071067811865475e-01, 0.0, 0.0, 1.0,
[…]

### 2014-11-26 examples/plot_sorted.py
Reverted the previous commit that wrongly modified two code files.

@@ -214,3 +214,4 @@ for orientation in range( 26 ):
 
 
 
+ 

### 2014-11-26 examples/test_bh_sorted.c

@@ -3240,3 +3240,4 @@ int main(int argc, char *argv[]) {
 
   return 0;
 }
+ 

### 2014-12-10 examples/plot_sorted.py
Preliminary version of the sorted interactions using quadrupoles.

@@ -74,7 +74,8 @@ for orientation in range( 26 ):
     data=loadtxt( "interaction_dump_%d.dat"%orientation )
     id = data[:,0]
     pos = data[:,2]
-    
+    pos -= mean(pos)
+        
     accx_u=data[:,6]
     accy_u=data[:,7]
     accz_u=data[:,8]
@@ -114,6 +115,7 @@ for orientation in range( 26 ):
     #plot(id[abs(errx_s) > 0.001], e_errx_s , 'ro')
     plot(pos, e_errx_s , 'ro')
     text( 0., 0.1, "axis=( %d %d %d )"%(axis[orientation*3 + 0], axis[orientation*3 + 1], axis[orientation*3 + 2]) , ha='center', backgroundcolor='w', fontsize=14)
+    xlim(-1.2*max(abs(pos)), 1.2*max(abs(pos)))
     ylim(-0.05, 0.05)
     grid()
     
@@ -121,6 +123,7 @@ for orientation in range( 26 ):
     #plot(id[abs(erry_s) > 0.001], e_erry_s , 'ro')
     plot(pos, e_erry_s , 'ro')
     text( 0., 0.1, "axis=( %d %d %d )"%(axis[orientation*3 + 0], axis[orientation*3 + 1], axis[orientation*3 + 2]) , ha='center', backgroundcolor='w', fontsize=14)
+    xlim(-1.2*max(abs(pos)), 1.2*max(abs(pos)))
     ylim(-0.05, 0.05)  
     grid()
     
@@ -130,76 +133,76 @@ for orientation in range( 26 ):
     text( 0., 0.1, "axis=( %d %d %d )"%(axis[orientation*3 + 0], axis[orientation*3 + 1], axis[orientation*3 + 2]) , ha='center', backgroundcolor='w', fontsize=14)
 
     legend(loc="upper right")
-    
+    xlim(-1.2*max(abs(pos)), 1.2*max(abs(pos)))
     ylim(-0.05, 0.05)
     #xlim(0, size(id)-1)
     grid()
 
-    savefig("matthieu_accelerations_relative_%d.png"%orientation)
+    savefig("1quadrupole_accelerations_relative_%d.png"%orientation)
     close()
 
 
-    # Plot -------------------------------------------------------
-    figure(frameon=True)
+    # # Plot -------------------------------------------------------
+    # figure(frameon=True)
     
[…]

### 2014-12-10 examples/test_bh_sorted.c

@@ -42,13 +42,13 @@
 #define const_G 1    // 6.6738e-8
 #define dist_min 0.5 /* Used for legacy walk only */
 #define dist_cutoff_ratio 1.5
-#define iact_pair_direct iact_pair_direct_sorted_tripole
+#define iact_pair_direct iact_pair_direct_sorted_multipole
 
 #define ICHECK -1
 #define NO_SANITY_CHECKS
 #define NO_COM_AS_TASK
 #define COUNTERS
-#define MANY_MULTIPOLES
+//#define MANY_MULTIPOLES
 
 /** Data structure for the particles. */
 struct part {
@@ -68,7 +68,7 @@ struct index {
   float d;
 };
 
-struct multipole {
+struct monopole {
   double com[3];
   float mass;
 };
@@ -90,10 +90,10 @@ struct cell {
   union {
 
     /* Information for the legacy walk */
-    struct multipole legacy;
+    struct monopole legacy;
 
     /* Information for the QuickShed walk */
-    struct multipole new;
+    struct monopole new;
   };
 
   int res, com_tid;
@@ -183,108 +183,128 @@ static inline void dipole_iact(const struct dipole *d, const float *x,
   for (k = 0; k < 3; k++) a[k] -= w * dx[k];
 }
 
-/** Tripole stuff. Moment-matching multipole along a given direction. */
-struct tripole {
-  float axis1[3], axis2[3];
+
+
+
+/** Multipole stuff. */
+struct multipole {
   float x[3];
   float mass;
-  float da, da2, db, db2;
-  int count;
+  float Q[3][3];
 };
 
-static inline void tripole_init(struct tripole *d, const float *axis1, const float *axis2) {
+
+static inline void multipole_init(struct multipole *d) {
   for (int k = 0; k < 3; k++) {
-    d->axis1[k] = axis1[k];
-    d->axis2[k] = axis2[k];
     d->x[k] = 0.0;
   }
+  for (int i = 0; i < 3; i++) {
+    for (int j = 0; j < 3; j++) {
+      d->Q[i][j] = 0.0;
+    }
+  }
+
   d->mass = 0.0;
-  d->da = 0.0; d->db = 0.0;
-  d->da2 = 0.0; d->db2 = 0.0;
-  d->count = 0;
 }
 
[…]

### 2014-12-11 examples/plot_sorted.py
Correct version of the quadrupole accelerations.

@@ -69,6 +69,11 @@ axis = [
 #names = ["side", "edge", "corner"]
 
 for orientation in range( 26 ):
+# for jjj in range(2):
+#     if jjj == 0:
+#         orientation = 0
+#     if jjj == 1:
+#         orientation = 8
 
     # Read Quickshed accelerations
     data=loadtxt( "interaction_dump_%d.dat"%orientation )
@@ -114,6 +119,11 @@ for orientation in range( 26 ):
     subplot(311, title="Acceleration along X")
     #plot(id[abs(errx_s) > 0.001], e_errx_s , 'ro')
     plot(pos, e_errx_s , 'ro')
+    # for j in range(size(pos)):
+    #     if ( pos[j-1] != pos[j] ):
+    #         text(pos[j], e_errx_s[j]-0.005, "%d"%id[j], fontsize=10)
+    #     else:
+    #         text(pos[j], e_errx_s[j]-0.015, "%d"%id[j], fontsize=10)
     text( 0., 0.1, "axis=( %d %d %d )"%(axis[orientation*3 + 0], axis[orientation*3 + 1], axis[orientation*3 + 2]) , ha='center', backgroundcolor='w', fontsize=14)
     xlim(-1.2*max(abs(pos)), 1.2*max(abs(pos)))
     ylim(-0.05, 0.05)
@@ -122,6 +132,11 @@ for orientation in range( 26 ):
     subplot(312, title="Acceleration along Y")
     #plot(id[abs(erry_s) > 0.001], e_erry_s , 'ro')
     plot(pos, e_erry_s , 'ro')
+    # for j in range(size(pos)):
+    #     if ( pos[j-1] != pos[j] ):
+    #         text(pos[j], e_errx_s[j]-0.005, "%d"%id[j], fontsize=10)
+    #     else:
+    #         text(pos[j], e_errx_s[j]-0.015, "%d"%id[j], fontsize=10)
     text( 0., 0.1, "axis=( %d %d %d )"%(axis[orientation*3 + 0], axis[orientation*3 + 1], axis[orientation*3 + 2]) , ha='center', backgroundcolor='w', fontsize=14)
     xlim(-1.2*max(abs(pos)), 1.2*max(abs(pos)))
     ylim(-0.05, 0.05)  
@@ -130,6 +145,11 @@ for orientation in range( 26 ):
     subplot(313, title="Acceleration along Z")
[…]

### 2014-12-11 examples/test_bh_sorted.c

@@ -43,7 +43,6 @@
 #define dist_min 0.5 /* Used for legacy walk only */
 #define dist_cutoff_ratio 1.5
 #define iact_pair_direct iact_pair_direct_sorted_multipole
-
 #define ICHECK -1
 #define NO_SANITY_CHECKS
 #define NO_COM_AS_TASK
@@ -232,6 +231,8 @@ static inline void multipole_add_matrix(struct multipole *d, const float *x, flo
   for ( k = 0; k < 3; ++k) 
     dx[k] = x[k] - d->x[k] / d->mass;
   
+  //  message("x_i: %f %f %f    dx: %f %f %f", x[0], x[1], x[2], dx[0], dx[1], dx[2]);
+
   r2 = dx[0] * dx[0] + dx[1] * dx[1] + dx[2] * dx[2];
 
   /* Now fill in the matrix */
@@ -282,23 +283,37 @@ static inline void multipole_iact(struct multipole *d, const float *x, float* a)
   
   /* Compute the acceleration from the monopole */
   w = const_G * ir * ir * ir * d->mass;
-  for (k = 0; k < 3; k++) a[k] -= w * dx[k];
+  float a_m[3] = {0., 0., 0.};
+  for (k = 0; k < 3; k++) a_m[k] -= w * dx[k];
 
+  
   //return;
 
   /* Compute the acceleration from the quadrupole */
-  w = const_G * ir * ir * ir * ir * ir * ir * ir * 0.5f;
-  for (i=0 ; i<3 ; ++i) {
-    for (j=0 ; j<3 ; ++j) {
-
-      a[0] += w * dx[i]*dx[j]*(2.*dx[j] - 5.*dx[0]) * d->Q[i][j];
-      a[1] += w * dx[i]*dx[j]*(2.*dx[j] - 5.*dx[1]) * d->Q[i][j];
-      a[2] += w * dx[i]*dx[j]*(2.*dx[j] - 5.*dx[2]) * d->Q[i][j];
-	
+  float a_q[3] = {0., 0., 0.};
+  float w1 = 0.5f * const_G * ir * ir * ir * ir * ir; 
+  float w2 = -2.5f * const_G * ir * ir * ir * ir * ir * ir * ir;
+  float sum = 0.;
+  for (i = 0; i < 3; i++) {
[…]

### 2014-12-11 examples/plot_sorted_all.py
Cleaned up a bit the code

@@ -61,14 +61,14 @@ accz_bh=data[:,13]
     
 
 # Build error ------------------------------------------------
-    
-errx_s = (accx_s - accx_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
-erry_s = (accy_s - accy_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
-errz_s = (accz_s - accz_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
+inv_acc_tot = 1.0 / sqrt(accx_u**2 + accy_u**2 + accz_u**2)     
+errx_s = (accx_s - accx_u ) * inv_acc_tot
+erry_s = (accy_s - accy_u ) * inv_acc_tot
+errz_s = (accz_s - accz_u ) * inv_acc_tot
 
-errx_bh = (accx_bh - accx_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
-erry_bh = (accy_bh - accy_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
-errz_bh = (accz_bh - accz_u ) / sqrt(accx_u**2 + accy_u**2 + accz_u**2) 
+errx_bh = (accx_bh - accx_u ) * inv_acc_tot
+erry_bh = (accy_bh - accy_u ) * inv_acc_tot
+errz_bh = (accz_bh - accz_u ) * inv_acc_tot
 
 e_errx_s = errx_s#[abs(errx_s) > 0.001]
 e_erry_s = erry_s#[abs(erry_s) > 0.001]
@@ -94,10 +94,10 @@ figure(frameon=True)
 
 subplot(311, title="Acceleration along X")
 #plot(id[abs(errx_s) > 0.001], e_errx_s , 'ro')
-plot(pos_x, e_errx_s , 'ro')
 plot(pos_x, e_errx_bh , 'bx', label='B-H')
+plot(pos_x, e_errx_s , 'ro')
 #text( 0., 0.1, "axis=( %d %d %d )"%(axis[orientation*3 + 0], axis[orientation*3 + 1], axis[orientation*3 + 2]) , ha='center', backgroundcolor='w', fontsize=14)
-ylim(-0.2, 0.2)
+ylim(-0.1, 0.1)
 grid()
 
 subplot(312, title="Acceleration along Y")
@@ -105,7 +105,7 @@ subplot(312, title="Acceleration along Y")
[…]

### 2014-12-11 examples/test_bh_sorted.c

@@ -231,8 +231,6 @@ static inline void multipole_add_matrix(struct multipole *d, const float *x, flo
   for ( k = 0; k < 3; ++k) 
     dx[k] = x[k] - d->x[k] / d->mass;
   
-  //  message("x_i: %f %f %f    dx: %f %f %f", x[0], x[1], x[2], dx[0], dx[1], dx[2]);
-
   r2 = dx[0] * dx[0] + dx[1] * dx[1] + dx[2] * dx[2];
 
   /* Now fill in the matrix */
@@ -273,7 +271,7 @@ static inline void multipole_iact(struct multipole *d, const float *x, float* a)
 
   float inv_mass = 1. / d->mass;
 
-  /* Compute the distance. */
+  /* Compute the distance to the CoM. */
   for (r2 = 0.0f, k = 0; k < 3; k++) {
     dx[k] = x[k] - d->x[k] * inv_mass;
     r2 += dx[k] * dx[k];
@@ -286,8 +284,6 @@ static inline void multipole_iact(struct multipole *d, const float *x, float* a)
   float a_m[3] = {0., 0., 0.};
   for (k = 0; k < 3; k++) a_m[k] -= w * dx[k];
 
-  
-  //return;
 
   /* Compute the acceleration from the quadrupole */
   float a_q[3] = {0., 0., 0.};
@@ -303,7 +299,8 @@ static inline void multipole_iact(struct multipole *d, const float *x, float* a)
   a_q[1] = w2 * sum * dx[1] + w1 * (2.*dx[0]*d->Q[1][0] + 2.*dx[1]*d->Q[1][1] + 2.*dx[2]*d->Q[1][2]);
   a_q[2] = w2 * sum * dx[2] + w1 * (2.*dx[0]*d->Q[2][0] + 2.*dx[1]*d->Q[2][1] + 2.*dx[2]*d->Q[2][2]);
 
-
+  
+  /* Get the total acceleration */
   float a_t[3] = {0., 0., 0.};
   for (k = 0; k < 3; k++) a_t[k] = a_m[k] + a_q[k];
 
@@ -1861,8 +1858,6 @@ static inline void iact_pair_direct_sorted_multipole(struct cell *ci,
 
 #endif
 
-  //message("start");
-
   /* Get the sorted indices and stuff. */
   struct index *ind_i, *ind_j;
   struct multipole dip[4];
@@ -1872,9 +1867,6 @@ static inline void iact_pair_direct_sorted_multipole(struct cell *ci,
[…]

### 2014-12-11 examples/plot_sorted_all.py
Improved the plotting script

@@ -167,28 +167,46 @@ close()
 # Plot -------------------------------------------------------
 bins = linspace(-3, 3, 10000)
 
-e_errx_s = errx_s[(abs(errx_s) >= 0.000) & (abs(errx_s) < 1.)]
-e_erry_s = erry_s[(abs(erry_s) >= 0.000) & (abs(erry_s) < 1.)]
-e_errz_s = errz_s[(abs(errz_s) >= 0.000) & (abs(errz_s) < 1.)]
+e_errx_s1 = errx_s[ pos_x < 0.5 ]
+e_erry_s1 = erry_s[ pos_y < 0.5 ]
+e_errz_s1 = errz_s[ pos_z < 0.5 ]
+
+e_errx_s2 = errx_s[ pos_x > 0.5 ]
+e_erry_s2 = erry_s[ pos_y > 0.5 ]
+e_errz_s2 = errz_s[ pos_z > 0.5 ]
+
+e_errx_bh1 = errx_bh[ pos_x < 0.5 ]
+e_erry_bh1 = erry_bh[ pos_y < 0.5 ]
+e_errz_bh1 = errz_bh[ pos_z < 0.5 ]
+
+e_errx_bh2 = errx_bh[ pos_x > 0.5 ]
+e_erry_bh2 = erry_bh[ pos_y > 0.5 ]
+e_errz_bh2 = errz_bh[ pos_z > 0.5 ]
 
 
 
 figure(frameon=True)
 subplot(311, title="Acceleration along X")
-hist(e_errx_s, bins=bins, normed=1, histtype='step', rwidth=0.01, color='r', label="Sorted")
-hist(e_errx_bh, bins=bins, normed=1, histtype='step', rwidth=0.01, color='g', label="B-H")
+hist(e_errx_s1, bins=bins, normed=1, histtype='step', rwidth=0.01, color='r', label="Sorted")
+hist(e_errx_s2, bins=bins, normed=1, histtype='step', rwidth=0.01, color='r', ls='dashed')
+hist(e_errx_bh1, bins=bins, normed=1, histtype='step', rwidth=0.01, color='b', label="B-H")
+hist(e_errx_bh2, bins=bins, normed=1, histtype='step', rwidth=0.01, color='b', ls='dashed')
 legend(loc="upper right")
-xlim(-0.15, 0.15)
+xlim(-0.05, 0.05)
 
 subplot(312, title="Acceleration along Y")
-hist(e_erry_s, bins=bins, normed=1, histtype='step', rwidth=0.01, color='r')
[…]

### 2014-12-11 examples/test_bh_sorted.c
Simplified the multipole calculation (still N^2 operations)

@@ -187,38 +187,26 @@ static inline void dipole_iact(const struct dipole *d, const float *x,
 
 /** Multipole stuff. */
 struct multipole {
-  float x[3];
   float mass;
-  float Q[3][3];
+  float mu[3]; /* Centre of Mass */
+  float Q_00, Q_11, Q_22, Q_01, Q_12, Q_02;
 };
 
-
-static inline void multipole_init(struct multipole *d) {
-  for (int k = 0; k < 3; k++) {
-    d->x[k] = 0.0;
-  }
-  for (int i = 0; i < 3; i++) {
-    for (int j = 0; j < 3; j++) {
-      d->Q[i][j] = 0.0;
-    }
-  }
-
-  d->mass = 0.0;
-}
-
 static inline void multipole_reset(struct multipole *d) {
-  for (int k = 0; k < 3; k++) d->x[k] = 0.0;
-  for (int i = 0; i < 3; i++) {
-    for (int j = 0; j < 3; j++) {
-      d->Q[i][j] = 0.0;
-    }
-  }
+  for (int k = 0; k < 3; k++) d->mu[k] = 0.0;
+
+  d->Q_00 = 0.;
+  d->Q_11 = 0.;
+  d->Q_22 = 0.;
+  d->Q_01 = 0.;
+  d->Q_02 = 0.;
+  d->Q_12 = 0.;
 
   d->mass = 0.0;
 }
 
 static inline void multipole_add_com(struct multipole *d, const float *x, float mass) {
-  for (int k = 0; k < 3; k++) d->x[k] += x[k] * mass;
+  for (int k = 0; k < 3; k++) d->mu[k] += x[k] * mass;
   d->mass += mass;
 }
 
@@ -229,23 +217,19 @@ static inline void multipole_add_matrix(struct multipole *d, const float *x, flo
   
   /* Distance from CoM */
   for ( k = 0; k < 3; ++k) 
-    dx[k] = x[k] - d->x[k] / d->mass;
+    dx[k] = x[k] - d->mu[k] / d->mass;
   
   r2 = dx[0] * dx[0] + dx[1] * dx[1] + dx[2] * dx[2];
 
   /* Now fill in the matrix */
-  d->Q[0][0] += ((3. * mass * dx[0] * dx[0]) - (mass * r2)); 
-  d->Q[1][1] += ((3. * mass * dx[1] * dx[1]) - (mass * r2)); 
-  d->Q[2][2] += ((3. * mass * dx[2] * dx[2]) - (mass * r2)); 
[…]

### 2014-12-12 examples/test_bh_sorted.c
Updated version of the multipole calculation where the matrix is computed on the fly. Algorithm is now O(N).

@@ -189,99 +189,97 @@ static inline void dipole_iact(const struct dipole *d, const float *x,
 struct multipole {
   float mass;
   float mu[3]; /* Centre of Mass */
-  float Q_00, Q_11, Q_22, Q_01, Q_12, Q_02;
+  float sigma_00, sigma_11, sigma_22, sigma_01, sigma_02, sigma_12;
 };
 
 static inline void multipole_reset(struct multipole *d) {
-  for (int k = 0; k < 3; k++) d->mu[k] = 0.0;
+  d->mu[0] = 0.0;
+  d->mu[1] = 0.0;
+  d->mu[2] = 0.0;
 
-  d->Q_00 = 0.;
-  d->Q_11 = 0.;
-  d->Q_22 = 0.;
-  d->Q_01 = 0.;
-  d->Q_02 = 0.;
-  d->Q_12 = 0.;
+  d->sigma_00 = 0.0;
+  d->sigma_11 = 0.0;
+  d->sigma_22 = 0.0;
+  d->sigma_01 = 0.0;
+  d->sigma_02 = 0.0;
+  d->sigma_12 = 0.0;
 
   d->mass = 0.0;
 }
 
-static inline void multipole_add_com(struct multipole *d, const float *x, float mass) {
-  for (int k = 0; k < 3; k++) d->mu[k] += x[k] * mass;
+static inline void multipole_add(struct multipole *d, const float *x, float mass) {
+  d->mu[0] += x[0] * mass;
+  d->mu[1] += x[1] * mass;
+  d->mu[2] += x[2] * mass;
+  
+  d->sigma_00 += mass * x[0] * x[0];
+  d->sigma_11 += mass * x[1] * x[1];
+  d->sigma_22 += mass * x[2] * x[2];
+  d->sigma_01 += mass * x[0] * x[1];
+  d->sigma_02 += mass * x[0] * x[2];
+  d->sigma_12 += mass * x[1] * x[2];
+
   d->mass += mass;
 }
 
-static inline void multipole_add_matrix(struct multipole *d, const float *x, float mass) {
-
-  float r2, dx[3];
-  int k;
-  
-  /* Distance from CoM */
-  for ( k = 0; k < 3; ++k) 
-    dx[k] = x[k] - d->mu[k] / d->mass;
-  
-  r2 = dx[0] * dx[0] + dx[1] * dx[1] + dx[2] * dx[2];
-
-  /* Now fill in the matrix */
[…]

### 2014-12-12 examples/test_bh_sorted.c
Simplified a bit the quadrupole calculation

@@ -226,36 +226,34 @@ static inline void multipole_add(struct multipole *d, const float *x, float mass
 static inline void multipole_iact(struct multipole *d, const float *x, float* a) {
 
   int k;
-  float r2,ir,w, dx[3], mu[3];
-  float Q_00, Q_11, Q_22, Q_01, Q_02, Q_12;
+  float r2,ir,w, dx[3];
 
   /* early abort */
   if (d->mass == 0.) return;
 
   float inv_mass = 1. / d->mass;
   
-  /* Construct CoM */
-  for ( k = 0 ; k < 3 ; ++k ) mu[k] = d->mu[k] * inv_mass;
-
   /* Temporary quantity */
-  dx[0] = d->sigma_00 - d->mass*mu[0]*mu[0];  /* Abusing a so far unused */
-  dx[1] = d->sigma_11 - d->mass*mu[1]*mu[1];  /* variable temporarily... */
-  dx[2] = d->sigma_22 - d->mass*mu[2]*mu[2];
+  dx[0] = d->sigma_00 - inv_mass*d->mu[0]*d->mu[0];  /* Abusing a so far unused */
+  dx[1] = d->sigma_11 - inv_mass*d->mu[1]*d->mu[1];  /* variable temporarily... */
+  dx[2] = d->sigma_22 - inv_mass*d->mu[2]*d->mu[2];
 
   /* Construct quadrupole matrix */
-  Q_00 = 2.f*dx[0] - dx[1] - dx[2]; 
-  Q_11 = 2.f*dx[1] - dx[0] - dx[2]; 
-  Q_22 = 2.f*dx[2] - dx[0] - dx[1]; 
-  Q_01 = d->sigma_01 - d->mass*mu[0]*mu[1];
-  Q_02 = d->sigma_02 - d->mass*mu[0]*mu[2];
-  Q_12 = d->sigma_12 - d->mass*mu[1]*mu[2];
-  Q_01 *= 3.f;
-  Q_02 *= 3.f;
-  Q_12 *= 3.f;
+  float Q_00 = 2.0f*dx[0] - dx[1] - dx[2]; 
+  float Q_11 = 2.0f*dx[1] - dx[0] - dx[2]; 
+  float Q_22 = Q_00 + Q_11; /* Q_ij is traceless */
+  Q_22 *= -1.0f;
+
[…]

### 2014-12-12 examples/test_bh_sorted.c
Code clean up

@@ -46,8 +46,8 @@
 #define ICHECK -1
 #define NO_SANITY_CHECKS
 #define NO_COM_AS_TASK
-#define COUNTERS
-//#define MANY_MULTIPOLES
+//#define COUNTERS
+
 
 /** Data structure for the particles. */
 struct part {
@@ -109,78 +109,6 @@ enum task_type {
   task_type_com,
   task_type_count
 };
-
-/** Dipole stuff. Moment-matching multipole along a given direction. */
-struct dipole {
-  float axis[3];
-  float x[3];
-  float mass;
-  float da, da2;
-};
-
-static inline void dipole_init(struct dipole *d, const float *axis) {
-  for (int k = 0; k < 3; k++) {
-    d->axis[k] = axis[k];
-    d->x[k] = 0.0;
-  }
-  d->mass = 0.0;
-  d->da = 0.0;
-  d->da2 = 0.0;
-}
-
-static inline void dipole_reset(struct dipole *d) {
-  for (int k = 0; k < 3; k++) d->x[k] = 0.0;
-  d->mass = 0.0;
-  d->da = 0.0;
-  d->da2 = 0.0;
-}
-
-static inline void dipole_add(struct dipole *d, const float *x, float mass,
-                              float da) {
-  for (int k = 0; k < 3; k++) d->x[k] += x[k] * mass;
-  d->mass += mass;
-  d->da += da * mass;
-  d->da2 += da * da * mass;
-}
-
-static inline void dipole_iact(const struct dipole *d, const float *x,
-                               float *a) {
-  // Bail early if no mass.
-  if (d->mass == 0.0) return;
-
-  float inv_mass = 1.0f / d->mass;
-  float dx[3], r2, ir, w;
-  int k;
-
-  /* Get the offset along the dipole axis. This looks weird, but negative
-     offsets can happen due to rounding error. */
-  float offset = (d->da2 - d->da * d->da * inv_mass) * inv_mass;
-  if (offset > 0) {
-    offset = sqrtf(offset);
-  } else {
-    offset = 0.0f;
-  }
-
-  //offset = 0.f;
-
-  /* Compute the first dipole. */
[…]

### 2014-12-14 examples/test_bh.c
add single-precision version of the interaction functions.

@@ -619,13 +619,13 @@ static inline void iact_self_pc(struct cell *c, struct cell *leaf) {
  * @param ci The #cell containing the particles.
  * @param cj The #cell containing the other particles
  */
-static inline void iact_pair_direct(struct cell *ci, struct cell *cj) {
+static inline void iact_pair_direct_dp(struct cell *ci, struct cell *cj) {
 
   int i, j, k;
   int count_i = ci->count, count_j = cj->count;
-  struct part *parts_i = ci->parts, *parts_j = cj->parts;
   double xi[3];
   float dx[3], ai[3], mi, mj, r2, w, ir;
+  struct part *parts_i = ci->parts, *parts_j = cj->parts;
 
 #ifdef SANITY_CHECKS
 
@@ -688,6 +688,112 @@ static inline void iact_pair_direct(struct cell *ci, struct cell *cj) {
   } /* loop over all particles. */
 }
 
+static inline void iact_pair_direct(struct cell *ci, struct cell *cj) {
+
+  int i, j, k;
+  int count_i = ci->count, count_j = cj->count;
+  float xi[3];
+  float dx[3], ai[3], mi, mj, r2, w, ir;
+  double com[3];
+  struct part_local {
+    float x[3];
+    float a[3];
+    float mass;
+  } parts_i[count_i], parts_j[count_j] __attribute__((aligned(64)));;
+
+
+#ifdef SANITY_CHECKS
+
+  /* Bad stuff will happen if cell sizes are different */
+  if (ci->h != cj->h)
+    error("Non matching cell sizes !! h_i=%f h_j=%f\n", ci->h, cj->h);
+
+#endif
+
+  /* Find the center point of the interaction. */
+  for (k = 0; k < 3; k++) {
+    com[k] = 0.5 * (ci->loc[k] + cj->loc[k] + ci->h);
+  }
+
+  /* Init the local copies of the particles. */
+  for (k = 0; k < count_i; k++) {
+    for (j = 0; j < 3; j++) {
+      parts_i[k].x[j] = ci->parts[k].x[j] - com[j];
+      parts_i[k].a[j] = 0.0f;
+    }
+    parts_i[k].mass = ci->parts[k].mass;
+  }
+  for (k = 0; k < count_j; k++) {
+    for (j = 0; j < 3; j++) {
+      parts_j[k].x[j] = cj->parts[k].x[j] - com[j];
+      parts_j[k].a[j] = 0.0f;
+    }
+    parts_j[k].mass = ci->parts[k].mass;
[…]

### 2015-01-13 examples/test_fmm_sorted.c
Initial work on the FMM code. No non-direct interactions are computed any more.

@@ -104,7 +104,7 @@ struct cell {
 enum task_type {
   task_type_self = 0,
   task_type_pair,
-  task_type_self_pc,
+  //task_type_self_pc,
   task_type_pair_direct,
   task_type_com,
   task_type_count
@@ -724,15 +724,8 @@ void cell_split(struct cell *c, struct qsched *s) {
 #endif
 
     /* Otherwise, we're at a leaf, so create the cell's particle-cell task. */
-  } else {
-    struct cell *data[2] = {root, c};
-    int tid = qsched_addtask(s, task_type_self_pc, task_flag_none, data,
-                             2 * sizeof(struct cell *), 1);
-    qsched_addlock(s, tid, c->res);
-#ifdef COM_AS_TASK
-    qsched_addunlock(s, root->com_tid, tid);
-#endif
-  } /* does the cell need to be split? */
+  }
+
 
 /* Compute the cell's center of mass. */
 #ifndef COM_AS_TASK
@@ -916,129 +909,129 @@ static inline int are_neighbours(struct cell *ci, struct cell *cj) {
   return (dx[0] <= min_dist) && (dx[1] <= min_dist) && (dx[2] <= min_dist);
 }
 
-/**
- * @brief Compute the interactions between all particles in a cell leaf
- *        and the center of mass of all the cells in a part of the tree
-* described by ci and cj
- *
- * @param ci The #cell containing the particle
- * @param cj The #cell containing the center of mass.
- */
-static inline void iact_pair_pc(struct cell *ci, struct cell *cj,
-                                struct cell *leaf) {
+/* /\** */
+/*  * @brief Compute the interactions between all particles in a cell leaf */
+/*  *        and the center of mass of all the cells in a part of the tree */
+/* * described by ci and cj */
+/*  * */
+/*  * @param ci The #cell containing the particle */
+/*  * @param cj The #cell containing the center of mass. */
+/*  *\/ */
+/* static inline void iact_pair_pc(struct cell *ci, struct cell *cj, */
+/*                                 struct cell *leaf) { */
 
-  struct cell *cp, *cps;
+/*   struct cell *cp, *cps; */
 
-#ifdef SANITY_CHECKS
+/* #ifdef SANITY_CHECKS */
 
-  /* Early abort? */
-  if (ci->count == 0 || cj->count == 0) error("Empty cell !");
+/*   /\* Early abort? *\/ */
[…]

### 2015-01-14 examples/test_bh_sorted.c
merged-in changes made in the svn last night, i.e. added the single-precision code from test_bh.c and cleaned up the multipole interaction computation a bit.

@@ -25,6 +25,7 @@
 #include <stdio.h>
 #include <stdlib.h>
 #include <string.h>
+#include <strings.h>
 #include <unistd.h>
 #include <math.h>
 #include <float.h>
@@ -37,7 +38,7 @@
 
 /* Some local constants. */
 #define cell_pool_grow 1000
-#define cell_maxparts 1
+#define cell_maxparts 100
 #define task_limit 1e8
 #define const_G 1    // 6.6738e-8
 #define dist_min 0.5 /* Used for legacy walk only */
@@ -60,6 +61,13 @@ struct part {
   float mass;
   int id;
 };  // __attribute__((aligned(64)));
+
+struct part_local {
+  float x[3];
+  float a[3];
+  float mass;
+  float d;
+} __attribute__((aligned(32)));
 
 /** Data structure for the sorted particle positions. */
 struct index {
@@ -121,31 +129,21 @@ struct multipole {
 };
 
 static inline void multipole_reset(struct multipole *d) {
-  d->mu[0] = 0.0;
-  d->mu[1] = 0.0;
-  d->mu[2] = 0.0;
-
-  d->sigma_00 = 0.0;
-  d->sigma_11 = 0.0;
-  d->sigma_22 = 0.0;
-  d->sigma_01 = 0.0;
-  d->sigma_02 = 0.0;
-  d->sigma_12 = 0.0;
-
-  d->mass = 0.0;
+  bzero(d, sizeof(struct multipole));
 }
 
 static inline void multipole_add(struct multipole *d, const float *x, float mass) {
-  d->mu[0] += x[0] * mass;
-  d->mu[1] += x[1] * mass;
-  d->mu[2] += x[2] * mass;
+  float xm[3] = { x[0] * mass, x[1] * mass, x[2] * mass };
+  d->mu[0] += xm[0];
+  d->mu[1] += xm[1];
+  d->mu[2] += xm[2];
   
-  d->sigma_00 += mass * x[0] * x[0];
-  d->sigma_11 += mass * x[1] * x[1];
-  d->sigma_22 += mass * x[2] * x[2];
-  d->sigma_01 += mass * x[0] * x[1];
-  d->sigma_02 += mass * x[0] * x[2];
-  d->sigma_12 += mass * x[1] * x[2];
+  d->sigma_00 += xm[0] * x[0];
+  d->sigma_11 += xm[1] * x[1];
[…]

### 2015-06-09 paper/paper.tex
started porting to PeerJ template.

@@
-\documentclass[preprint]{elsarticle}
+\documentclass[fleqn,10pt]{wlpeerj}
 
@@
 
-% Document starts
-\begin{document}
-
 % Title portion
 \title{QuickSched: Task-based parallelism with dependencies and conflicts}
-\author[ecs,ggl]{Pedro Gonnet\corref{cor}} \ead{pedro.gonnet@durham.ac.uk}
-\author[ecs]{Aidan B. G. Chalk} \ead{aidan.chalk@durham.ac.uk}
-\author[icc]{Matthieu Schaller} \ead{matthieu.schaller@durham.ac.uk}
-\cortext[cor]{Corresponding author}
-\address[ecs]{School of Engineering and Computing Sciences, Durham University, South Road, Durham DH1 3LE, United Kingdom.}
-\address[icc]{Institute for Computational Cosmology, Durham University, South Road, Durham DH1 3LE, United Kingdom.}
-\address[ggl]{Google Switzerland GmbH, Brandshenkestr. 110, 8002 Z\"urich, Switzerland.}
+\author[1,3]{Pedro Gonnet}
+\author[1]{Aidan B. G. Chalk}
+\author[2]{Matthieu Schaller}
+\affil[1]{School of Engineering and Computing Sciences, Durham University, United Kingdom.}
+\affil[2]{Institute for Computational Cosmology, Durham University, United Kingdom.}
+\affil[3]{Google Switzerland GmbH, Z\"urich, Switzerland.}
+
+\keywords{task-based parallelism, shared-memory parallelism, high performance computing, parallel computing}
 
@@
 
-\begin{keyword}
-    task-based parallelism \sep shared-memory parallelism \sep high performance computing \sep parallel computing
-\end{keyword}
-
+% Document starts
+\begin{document}
+
+\flushbottom
 \maketitle
+\thispagestyle{empty}
 
@@
 
-This computational model is trivial to parallelize.
-Given a set of inter-dependent tasks and a set of computational
+Once modelled in such a way, the computation is somewhat trivial
+to parallelize:
+given a set of inter-dependent tasks and a set of computational
 threads, each thread repeatedly selects a task with no
@@
 new tasks, or until all tasks in the DAG have been executed.
+Note that although the parallel execution
+itself is trivial, it does not always guaranteed to be efficient.
+Several factors may limit the maximum degree of parallelism, e.g.~the
+structure of the task dependency DAG itself, or the order in which
+available tasks are executed.
+
 \fig{Tasks} shows such a DAG for a set of tasks.
@@
 One of the first implementations of a task-based parallel programming
-systems is Cilk \cite{ref:Blumofe1995}, an extension to the C
+systems is Cilk \citep{ref:Blumofe1995}, an extension to the C
 programming language which allows function calls to be ``spawned''
@@
 
-In SMP superscalar \cite{ref:Perez2008}, StarPU \cite{ref:Augonnet2011},
-QUARK \cite{ref:Yarkhan2011}, and KAAPI \cite{ref:Gautier2007}
+In SMP superscalar \citep{ref:Perez2008}, StarPU \citep{ref:Augonnet2011},
+QUARK \citep{ref:Yarkhan2011}, and KAAPI \citep{ref:Gautier2007}
 the programmer specifies
@@
 Intel's Threading Building Blocks (TBB)
-\cite{ref:Reinders2010}
+\citep{ref:Reinders2010}
 provide task-based parallelism using C++ templates in which
@@
 for spawning tasks, similar to Cilk, as of version 3.0
-\cite{ref:OpenMP2008}.
-OmpSs \cite{ref:Duran2011} extends this scheme with automatic
+\citep{ref:OpenMP2008}.
+OmpSs \citep{ref:Duran2011} extends this scheme with automatic
 dependency generation as in SMP superscalar, of which it
@@
 In order to avoid concurrent access to that resource, it is
-imperative that the execution of both tasks do not overlap,
+imperative that the execution of both tasks does not overlap,
 yet the order in which the tasks are executed is irrelevant.
@@
   \item Implicitly via spawning and waiting, e.g.~as is done in Cilk
-    \cite{ref:Blumofe1995}, or
+    \citep{ref:Blumofe1995}, or
   \item Automatic extraction from data dependencies, e.g.~as is done in OmpSs
-    \cite{ref:Duran2011}.
+    \citep{ref:Duran2011}.
 \end{itemize}
@@
 second limits their size.
-The criteria are thus used to optimize the
-cache efficiency of the computation.
+The parameters controlling the size of the tasks in the examples, 
+i.e.~the tile size in the QR decomposition and the limits $n_\mathsf{max}$
+and $n_\mathsf{task}$ were determined empirically and only optimized
+to the closest power of two or rough power of ten, respectively.
 
@@
 
-The QuickSched task scheduler consist of four main
+The QuickSched task scheduler consists of four main
 objects types: {\em task}, {\em resource}, {\em scheduler},
@@
 plus the maximum dependent task's weight (see \fig{TaskWeight}).
+The task cost can be either a rough estimate provided by the user,
+or the actual cost of the same task last time it was executed.
 The task weights are computed by traversing
-the tasks in reverse topological order following their dependencies, i.e.
-as per \cite{ref:Kahn1962}, and computing each task's weight
-up the task DAG.
+the tasks DAG in reverse topological order following their dependencies,
+e.g.~as per \cite{ref:Kahn1962}, and computing each task's weight, e.g.
+\begin{equation*}
+  \mbox{weight}_i = \mbox{cost}_i + \max_{j \in \mbox{\small unlocks}_i}\left\{\mbox{weight}_j\right\}.
+\end{equation*}
+\noindent where $\mbox{weight}_i$ and $\mbox{cost}_i$ are the
+weight and cost of the $i$th task, respectively, and 
+$\mbox{unlocks}_i$ is the set of tasks that the $i$th task
+unlocks.
 
@@
 resource ID or the address in memory of the resource.
+
+Note also that although protecting the entire queue with a mutex
+is not particularly scalable, and several authors, e.g.~REFS,
+have presented concurrent data structures that avoid this type
+of locking.
+However, since we normally use one queue per computational thread,
+contention will only happens due to work-stealing, i.e.~when
+another idle computational thread tries to poach tasks.
+Since this happens only rarely, we opt for the simpler locking approach.
 
@@
 fills the queues (line~1).
-For simplicity, OpenMP \cite{ref:Dagum1998}, which is available
+For simplicity, OpenMP \citep{ref:Dagum1998}, which is available
 for most compilers, is used to create a parallel section
@@
 concurrently.
-A version using {\tt pthreads} \cite{ref:Pthreads1995}
+A version using {\tt pthreads} \citep{ref:Pthreads1995}
 directly\footnote{In most environments, OpenMP is implemented
@@
 If a task could be obtained from any queue and task re-owning
-is switched on (line~12),
+is enabled (line~12),
 the resources it locks and uses are marked as now being owned
@@
 described in \cite{ref:Buttari2009}, which has been used as a benchmark
-by other authors \cite{ref:Agullo2009b,ref:Badia2009,ref:Bosilca2012}.
+by other authors \citep{ref:Agullo2009b,ref:Badia2009,ref:Bosilca2012}.
 This example only requires dependencies and is presented 
@@
 These algorithms are now part of the PLASMA and MAGMA
-libraries for parallel linear algebra \cite{ref:Agullo2009}.
+libraries for parallel linear algebra \citep{ref:Agullo2009}.
 The former uses the QUARK task scheduler, which was originally
 designed for this specific task, while the latter currently uses
-the StarPU task scheduler \cite{ref:Agullo2011}.
+the StarPU task scheduler \citep{ref:Agullo2011}.
 
@@
 task additionally modifies the lower triangular part of the $(j,j)$th tile.
+
 Although the tile-based QR decomposition requires only dependencies,
@@
 tasks using the same tiles to the same thread.
+The resources/tiles are initially assigned to the queues in column-major
+order, i.e.~the first $\lfloor n_\mathsf{tiles}/n_\mathsf{queues}\rfloor$
+are assigned to the first queue, and so on.
 
@@
 as described above.
-For this matrix, a total of 11440 tasks with 32240 dependencies
+The task costs were initialized to the asymptotic cost of the underlying
+operations.
+For this matrix, a total of 11\,440 tasks with 32\,240 dependencies
 were generated.
@@
 Setting up the scheduler, tasks, and resources took, in all
-cases, an average of 7.2\,ms.
+cases, an average of 7.2\,ms, i.e.~at most 3\% of the total
+computational cost.
 
@@
 \end{quote}
-\noindent The scaling and efficiency relative to QuickSched are 
+\noindent Several different schedulers and parameterizations
+were discussed with the authors of OmpSs and tested, with
+the above settings produced the best results.
+
+The scaling and efficiency relative to QuickSched are 
 shown in \fig{QRResults}.
@@
 
-The Barnes-Hut tree-code \cite{ref:Barnes1986}
+The Barnes-Hut tree-code \citep{ref:Barnes1986}
 is an algorithm to approximate the
@@
 The cost of sorting the particles, with a recursive
-partitioning similar to QuickSort \cite{ref:Hoare1962},
+partitioning similar to QuickSort \citep{ref:Hoare1962},
 is in \oh{N\log N}.
@@
 A total of 43\,416 locks on 37\,449 resources were generated.
+Setting up the tasks took, on average XXX\,ms, i.e.~at most
+XXX\% of the total computation time.
+Storing the tasks, resources, and dependencies required XXX\,MB,
+which is only XX\% of the XXX\,MB required to store the particle
+data.
 
@@
 For comparison, the same computations were run using the popular
-astrophysics simulation software Gadget-2 \cite{ref:Springel2005},
+astrophysics simulation software Gadget-2 \citep{ref:Springel2005},
 using a traditional Barnes-Hut implementation based on octrees
 and distributed-memory parallelism based on domain decompositions
-and MPI \cite{ref:Snir1998}.
+and MPI \citep{ref:Snir1998}.
 To achieve the same accuracy, an opening angle of 0.5 was used.
@@
 % Bibliography
-\bibliographystyle{elsarticle-num}
+% \bibliographystyle{elsarticle-num}
 \bibliography{quicksched}
@@
 The following loops mirror the task generation described in
-Algorithm~2 of \cite{ref:Buttari2009}.
+Algorithm~2 of \citep{ref:Buttari2009}.
 For each level {\tt k} (line~10), a DGEQRF task is created

### 2015-10-20 paper/paper.tex
first round of corrections, sections 1--3. will update validation and results next.

@@
 hierarchical resources.
-The scheduler is shown to perform and scale well on a 64-core parallel
+The scheduler itself prioritizes tasks along the critical path
+of execution and is shown to perform and scale well on a 64-core parallel
 shared-memory machine for two example problems: A tiled QR
@@
 
-\fig{Tasks} shows such a DAG for a set of tasks.
-The arrows indicate the direction of the dependency, i.e.~an
+\fig{Tasks} shows such a DAG for a set of tasks with
+arrows indicating the direction of the dependencies, i.e.~an
 arrow from task $A$ to task $B$ indicates that task $B$ depends
@@
 
+Several libraries provide some mechanism to model such
+conflicts, either directly or indirectly.
+In the QUARK scheduler, conflicts can be modeled by explicitly
+marking dependencies as concurrent.
+KAAPI and OmpSS, on the other hand, allow marking access to
+certain variables as reductions, yet only for basic operations,
+e.g.~summation or maximum/minimum.
+
 This paper presents QuickSched, a framework for task-based
@@
         same core to preserve memory/cache locality as far as possible, and
-    \item {\em Parallel efficiency}: Tasks should be executed in an order
+    \item {\em Parallel efficiency}: The order in which the tasks
+        are executed should be chosen such
         that sufficient work is available for all computational
@@
   \item Implicitly via spawning and waiting, e.g.~as is done in Cilk
-    \citep{ref:Blumofe1995}, or
-  \item Automatic extraction from data dependencies, e.g.~as is done in OmpSs
-    \citep{ref:Duran2011}.
+    and OpenMP~3.0, or
+  \item Automatic extraction from data dependencies, e.g.~as is done in
+    StarPU, QUARK, and OmpSs.
 \end{itemize}
@@
 to the closest power of two or rough power of ten, respectively.
+Further tuning these parameters could very likely lead to further
+performance gains, but such an effort would go beyond the scope,
+and point, of this paper.
 
@@
 
-What the task does is determined by the {\tt type}
+{\em What} the task does is determined by the {\tt type}
 field, e.g.~which can be mapped to any particular function,
@@
 the tasks DAG in reverse topological order following their dependencies,
-e.g.~as per \cite{ref:Kahn1962}, and computing each task's weight, e.g.
+e.g.~as per \cite{ref:Kahn1962} in $\mathcal O(n)$ for $n$ tasks,
+and computing each task's weight, e.g.
 \begin{equation*}
@@
 Unfortunately, there is no way of efficiently traversing all
-the elements in the heap in decreasing order.
+the elements in such a heap in decreasing order.
 The array of tasks is therefore traversed as if it were sorted,
@@
 \end{minipage}\end{center}
-\noindent where as with the queue insertion, the queue is first
+\noindent where, as with the queue insertion, the queue is first
 locked for exclusive access (line~4).
@@
 and the heap order is restored (line~17).
-Finally, the queue lock is released (line~19) and the locked task
-or, if no lockable task could be found, {\tt NULL} is returned.
+Finally, the queue lock is released (line~19) and the locked task,
+or {\tt NULL} if no lockable task could be found, is returned.
 
@@
 
-Note also that although protecting the entire queue with a mutex
-is not particularly scalable, and several authors, e.g.~REFS,
+Note also that protecting the entire queue with a mutex
+is not particularly scalable, and several authors, e.g.~\cite{ref:Sundell2003},
 have presented concurrent data structures that avoid this type
@@
 Since this happens only rarely, we opt for the simpler locking approach.
+This decision is backed by the results in Section~5.
 
@@
 directly\footnote{In most environments, OpenMP is implemented
-on top of {\tt pthreads}, e.g. gcc's libgomp.} is also available.
+on top of {\tt pthreads}, e.g. the {\tt gcc} compiler's libgomp.}
+is also available.
 The parallel section consists of a loop (lines~7--10) in

### 2015-10-24 paper/paper.tex
second round of corrections. still need to check the appendix and re-do the numbers.

@@
 were discussed with the authors of OmpSs and tested, with
-the above settings produced the best results.
+the above settings producing the best results.
 
@@
 soon as possible.
-OmpSs, does not exploit this knowledge, resulting in the less efficient
+OmpSs does not exploit this knowledge, resulting in the less efficient
 scheduling seen in \fig{QRTasks}.
@@
 pairwise interactions between a set of $N$ particles,
-in \oh{N\log N} operations, as opposed to the \oh{N^2}
+in \oh{N\log N} operations, as opposed to in \oh{N^2} for the
 naive direct computation.
@@
     \item If called with two neighbouring cells
-        and one of the cells are not split, create
+        and at least one of the cells is not split, create
         a particle-particle pair task over both cells (line~29),
@@
 \end{itemize}
-\noindent where every interaction task additionally locks
+\noindent Every interaction task additionally locks
 the cells on which it operates (lines~17, 20, and 32--33).
@@
 if the cells contain more than a minimum number $n_\mathsf{task}$
-of threads each (lines~7 and~23).
+of particles each (lines~7 and~23).
 
@@
         Solving the N-Body problem takes 323\,ms, achieving 75\% parallel
-        efficiency, over all 64 cores.
+        efficiency over all 64 cores.
         For comparison, timings are shown for the same computation using
@@
 40\%.
-This is due to memory bandwidth restrictions, as
-the cost of the particle-cell interaction tasks, which do significantly more
-computation per memory access, only grow by up to 10\%.
+This is due to the cache hierarchy of the AMD Opteron 6376 in which
+pairs of cores share a comon 2\,MB L2 cache.
+When using half the cores or less, each core has its L@ cache to
+itself, whereas beyond 32 cores they are shared, resulting in more
+frequent cache misses.
+This cen be seen when comparing the costs of the particle-particle
+interaction and particle-cell interaction tasks: while the former grow by
+roughly 30\%, the latter grow by only 10\% as they do much more
+computation per memory access.
 
@@
         As of 32 cores, the cost of both pair interaction task
-        types grow by up to 40\%.
+        types grow by up to 30\%.
         The cost of the particle-cell interactions, which entail significantly more
@@
 \section*{Acknowledgments}
-The authors would like to thank Lydia Heck of the Institute for
-Computational Cosmology at Durham University for providing access
-to, and expertise on, the COSMA cluster used in the performance
-evaluation.
-This work was supported by a Durham University Seedcorn Grant.
+The authors would like to thank Tom Theuns and Richard Bowers of the
+Institute for Computational Cosmology at Durham University for the
+helpful discussions.
+This work was supported by a Durham University Seedcorn Grant
+number 21.12.080130 from
+which the hardware used in the experiments was purchased.
 
